### 2015-03-24 paper.tex
initial commit

\documentclass[a4paper]{article}
\usepackage{defaultstyle}
\usepackage{localstyle}

\newcommand{\comment}[1]{\textsf{#1}}

\title{Bringing AppPAL to Android}
%\subtitle{Capturing and Checking App Installation Policies}
\author{Joseph Hallett \and David Aspinall}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
  The abstract goes\ldots
  \begin{itemize}
    \item \comment{Sentence 1: State the problem}
    \item \comment{Sentence 2: Why is it an interesting problem}
    \item \comment{Sentence 3: What does my solution achieve}
    \item \comment{Sentence 4: What follows from my solution}
  \end{itemize}
\end{abstract}

\section{Introduction \comment{1 page}}

\section{The Problem \comment{1 page}}

\section{My idea \comment{2 pages}}

\section{The details \comment{5 pages}}

\section{Related work \comment{1--2 pages}}

\section{Conclusions and further work \comment{$\frac{1}{2}$ page}}

\bibliographystyle{plainnat}
\bibliography{paper}
\end{document}

### 2015-04-01 paper.tex
wrote most of details

@@
-\documentclass[a4paper]{article}
+% vim: set nowrap tw=0:
+\documentclass[a4paper]{scrartcl}
 \usepackage{defaultstyle}
@@
 
-\newcommand{\comment}[1]{\textsf{#1}}
+\newcommand{\comment}[1]{}%\textsf{#1}}
 
-\title{Bringing AppPAL to Android}
-%\subtitle{Capturing and Checking App Installation Policies}
+\title{AppPAL for Android}
+\subtitle{Capturing and Checking App Installation Policies}
 \author{Joseph Hallett \and David Aspinall}
@@
 \begin{abstract}
-  The abstract goes\ldots
-  \begin{itemize}
-    \item \comment{Sentence 1: State the problem}
-    \item \comment{Sentence 2: Why is it an interesting problem}
-    \item \comment{Sentence 3: What does my solution achieve}
-    \item \comment{Sentence 4: What follows from my solution}
-  \end{itemize}
+  % \begin{itemize}
+  %   \item \comment{Sentence 1: State the problem}
+  When a user wants to pick an app on Android they are forced to judge apps based only on the information presented to them by the store.
+  %   \item \comment{Sentence 2: Why is it an interesting problem}
+  Employers must rely on employees using company devices properly to enforce security policies about apps at work.
+  Whilst some users spend time to pick apps with care others don't and become frustrated at apps that steal personal information.
+  %   \item \comment{Sentence 3: What does my solution achieve}
+  We present AppPAL: a policy language that can enforce app installation policies using statements from trusted third parties and delegation relationships.
+  %   \item \comment{Sentence 4: What follows from my solution}
+  This gives us a rigorous and flexible framework for enforcing and comparing and app policies and the trust relationships surrounding them.
+  % \end{itemize}
 \end{abstract}
@@
 \section{Introduction \comment{1 page}}
+\label{sec:introduction}
 
-\section{The Problem \comment{1 page}}
+%\comment{The problem is\ldots}
+How do you pick which apps to install on your phone?  
+App stores give some information about their apps; such as permissions and review scores.
+Ultimately the decision which apps to use must be made by the user, however.
+
+Not all apps are suitable. 
+There is a large amount of \ac{pus} available for Android~\citep{Truong:2014bi,Svajcer:2013tp}. 
+Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
+Their company may want to restrict which apps their employees can use.
+The \ac{ciso} or IT department may set a policy to prevent information leaks.
+Some users worry apps will misuse their personal data.
+They avoid apps which can access their location, or contact data.
+They apply their own personal security policy when downloading apps.
+
+Currently these policies are enforced by hand.
+This is error-prone and can be inconsistent.
+A better solution is to use a logic of authorization.
+The policy can be written in the logic and enforced automatically.
+
+%\comment{My contribution is\ldots}
+We present AppPAL, an authorization logic for reasoning about Android apps.
+The language is an instantiation of SecPAL~\citep{Becker:2006vh} tailored to making statements about apps.
+It allows us to reason about apps using statements from third parties.
+We can express trust relationships in these speakers and use constraints to perform additional checks at evaluation time.
+This allows us to enforce deeper and more complex policies than existing policy tools like Kirin~\citep{Enck:2009ko}.
+
+\begin{itemize}
+  %\item \comment{Specifically I have}
+  \item Described several scenarios where an authorization logic could replace informal policies for companies and users (\autoref{sec:problem}.) 
+    \comment{Provide forward refs saying where we do this and about what}
+  \item Shown how app installation policies can be expressed in AppPAL
+    including how a corporate \ac{byod} policy could be written, and how a user's personal security policy could be formalized (\autoref{sec:idea}.)
+  \item Implemented the AppPAL language.  
+    Our implementation runs on Android and the JVM (\autoref{sec:details}.)
+\end{itemize}
+
+\section{The Problem\comment{1 page}}
+\label{sec:problem}
+
+Suppose \emph{Alice} works for \emph{Bob}.
+Bob allows Alice to use her personal phone as a work phone, but he has some specific concerns.
+\begin{itemize}
+  \item Alice shouldn't have any apps that can track her movements.
+    The testing labs are at a secret location and it mustn't be leaked.
+  \item Apps should come from a reputable source, such as the Play store.
+  \item Bob uses an \ac{av} program by McAfee. 
+    It should check all apps before they're installed.
+\end{itemize}
+
+To ensure this policy is fully met Alice could promise to follow it; but this is error-prone---what if she makes a mistake or misses something?
+Bob's policy could be implemented using existing tools.
+\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
+  could configure Alice's device to disallow apps from outside the Play Store.
+Various tools~\citep{Backes:2012vm,Jeon:2012ki,Hornyack:2011wq} can control the permissions or data an app can get.
+These could be used to used to ensure no location data is ever obtained.
+Alternately other tools~\citep{Enck:2009ko,Fritz:2013vi,Gordon:2015et} could check that the locations are ever leaked to the web.
+Various anti-virus programs are available for Android---one could be installed on Alice's phone.
+
+Whilst we \emph{could} implement Bob's policy using existing tools, it isn't a great solution.
+It isn't flexible, Alice needs to alter and remove the software on her phone if Bob changes his policy or she changes jobs.
+It isn't clear what an app must do to be installed, or what checks must have been done if it already has been put on the phone.
+The link between Alice, the user; Bob, the policy setter; and the tools Bob trusts to implement his policy isn't immediately apparent.
+
+This last point is particularly pertinent with respect to \textsc{CVE-2015-2077}.
+In 2015 Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
+Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
+Unfortunately they had configured it to ignore Superfish\footnote{The malware was an ad framework designed to show users products similar to those they viewed on the web.
+Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.}.
+
+From a user's perspective there was a delegation of trust to Lenovo to detect malware.
+Lenovo then delegated further to McAfee to supply the antivirus; but left an exception that they would allow Superfish.
+A user might wonder where the breach of trust occurred and who is to blame?  
+Is it the \ac{av} failing to spot the malware? 
+Has someone else configured something incorrectly?
+How should they fix the problem?
+
+If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
+\begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
+"user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
+"lenovo" says "mcafee" can-say inf File isSafe.
+"lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
+\end{lstlisting}
+The fault lies with \autoref{lst:lenovo_unsafe}.  
+Lenovo has caused the problem (they \emph{said} it was safe).
+The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
+
+It is important to stress that we are not trying to solve the problem of detecting malware.
+There are many tools, and much research, that already does this.
+Our problem is different. 
+Given there are some apps you want and others you don't:
+  how can you express your preferences in such a way that they can be enforced automatically.
+Further more how can we show the trust relationships used to make these decisions clearly and precisely.
 
 \section{My idea \comment{2 pages}}
+\label{sec:idea}
+
+Returning to Alice and Bob's dilemma in \autoref{sec:problem}, we propose a solution.
+Instead of using several different tools to enforce Bob's policy disjointedly, we use an authorization logic.
+In \autoref{lst:corporate} we give an AppPAL policy implementing Bob's app concerns on Alice's phone.
+A pictorial representation, is given in \autoref{fig:bobs_policy}; color is used to split statements by each speaker.
+
+\begin{figure}
+  \includegraphics{figures/bob-policy.eps}
+  \caption{Pictoral representation of Alice's app install policy.}
+\label{fig:bobs_policy}
+\end{figure}
+
+AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app installation policies. 
+SecPAL is a logic of authorization for access control decisions in distributed systems.
+It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
+SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\citep{Becker:2009ula} and data-sharing~\citep{Aziz:2011vt}.
+We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.
+
+\begin{lstlisting}[label={lst:corporate}, 
+                   caption={An AppPAL policy implementing Bob's security requirements},
+                   float,
+                   numbers=left,
+                   escapeinside={@}{@}]
+"alice" says "bob" can-say inf App isInstallable. @\label{lst:corporate_1}@
+"bob" says App isInstallable @\label{lst:corporate_2}@
+  if "no-tracking-policy" isMetBy(App),
+     "reputable-source-policy" isMetBy(App),
+     "anti-virus-policy" isMetBy(App).
+"bob" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
+  if App isBuyable.
+"bob" says "google-play" can-say 0 App isBuyable. @\label{lst:corporate_4}@
+"bob" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
+  if App isAnApp
+  where
+    mcAfeeVirusCheck(App) = False.
+"bob" says "no-location-permissions" can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
+"bob" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
+  if App isAnApp
+  where
+    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
+    hasPermission(App, "ACCESS_COURSE_LOCATION") = False.
+\end{lstlisting}
+
+In \autoref{lst:corporate_1} Alice gives Bob the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
+She allows him to delegate the decision further if he choses (\code{can-say inf}).
+Next in \autoref{lst:corporate_2} Bob specifies his concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
+If Bob can be convinced all these policies are met then he will say the \code{App isInstallable}.
+In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Bob specifies that an app meets the \code{reputable-source-policy} if the \code{App isBuyable};
+  with \code{"google-play"} specified as the decider of what is buyable or not.
+This time Google is not allowed to delegate the decision further (\code{can-say 0}).
+In other words Google is not allowed to specify Amazon as a supplier of apps as well.
+Google must say what is buyable directly for Bob to accept it.
+Bob specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
+Here we use a constraint.
+When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
+Only if this returns \code{False} will the policy be met.
+To specify the \code{"no-tracking-policy"} Bob says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
+Bob specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.
+
+Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
+To meet Bob's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isInstallable} predicate.
+Specifically it needs:
+\begin{description}
+  \item[\texttt{"bob" says "com.facebook.katana" isAnApp.}] 
+    A simple typing statement that can be generated for all apps as they are encountered.
+  \item[\texttt{"google-play" says "com.facebook.katana" isBuyable.}] 
+    Required to convince Bob the app came from a reputable source. 
+    It should be able to obtain this statement from the Play store as the app is available there.
+  \item[\texttt{"bob" says "anti-virus-policy" isMetBy("com.facebook.katana").}]
+    She can obtain this by running the \ac{av} program on her app.
+  \item[\texttt{"bob" says "no-locations-permissions" isMetBy("com.facebook.katana").}]
+    Needed to show the App meets Bob's no-tracking-policy.
+    Bob will say this if after examining the app the location permissions are missing.
+\end{description}
+These last two statements require the checker to do some extra checks to satisfy the constraints.
+To get the third statement it must run the \ac{av} program on her app and check the result.
+The results from the \ac{av} program may change with time as it's signatures are updated.  
+Consequently the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
+For the forth statement the checker needs to check the permissions of the app.
+It does this by looking in the \texttt{MANIFEST.xml} inside the app itself.
+
+In this scenario we have imagined Alice wanting to check the apps as she installs them.
+Alternatively we could imagine Bob wanting a personalised app store where all apps sold meet his policy.
+With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which meet the policy.
 
 \section{The details \comment{5 pages}}
+\label{sec:details}
+
+AppPAL is implemented as a library for Android. 
+A standalone Java version also exists for testing.  
+The parser is implemented using ANTLR4.
+Code, tests, and prebuilt APKs are available from Github\footnote{\url{https://github.com/bogwonch/AppPAL}}.
+
+\subsection{The Language}
+
+The structure of an AppPAL assertion is shown in \autoref{fig:assertion}.
+
+\begin{figure}
+  \centering
+  \begin{equation*}\small
+    \begin{array}{r l}
+      \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} & 
+      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isInstallable}}^{\text{predicate}}}^{\text{fact}} \\
+      & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
+      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = True}}^{\text{constraint}}. 
+    \end{array}
+  \end{equation*}
+  \caption{Structure of an AppPAL assertion.}
+\label{fig:assertion}
+\end{figure}
+
+In the \citeauthor{Becker:2006vh}'s paper~\citep{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
+With AppPAL we make explicit our predicates and how they relate to Android.  
+Some of these predicates require arguments.
+We add a light typing system to help ensure these predicates have the correct arguments.
+
+Specifically we define the following predicates:
+\begin{description}
+  \item[\texttt{App isAnApp}, \texttt{Policy isAPolicy}\ldots] 
+    Simple typing relations.  
+    The app \text{App} is an App, and so on.
+
+  \item[\texttt{App isInstallable}]
+    Used to indicate that an App meets the install policy for the device.
+    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
+
+  \item[\texttt{Policy isMetBy(App)}]
+    Policies let you split app behaviour into sub-policies.
+    For example in \autoref{sec:idea} we showed how Bob's installation policy could be written using three \texttt{isMetBy} statements.
+    Splitting policies allows greater control of how each one is checked.
+    We can delegate checking a policy to an expert using the \code{can-say} statement.
+    We can specify more detailed checks using the \code{can-act-as} statements.
+\end{description}
+
+We also add the following constraint functions which check simple properties of the app:
+In every case the check can be done by examining the app itself.
+\begin{description}
+  \item[\texttt{hasPermission(App, String)} : Bool]
+    Checks the permissions requested in the app's manifest to see if the string is included in them.
+
+  \item[\texttt{hasPackage(App, String)} : Bool]
+    Checks the app to see if a package called \emph{String} is included within it.
+    Examples might include aggressive advertisement frameworks a user wishes to ban.
+
+  \item[\texttt{hasNativeCode(App)} : Bool]
+    Checks to see whether the app contains native code.
+    Native (i.e. non-Dalvik) code is potentially more dangerous than the sandboxed Dalvik versions.
+    It is typically used for performance sensitive code (i.e. games).
+
+  \item[\texttt{isSignedBy(App, Speaker)} : Bool]
+    Checks the \texttt{CERT.RSA} used to sign the App and compares it with the certificate used to identify a speaker.
+
+  \item[\texttt{hasVersion(App)} : Int]
+    Gets the version number of an app.
+\end{description}
+
+We also add some constraints that do more complex checks:
+\begin{description}
+  \item[\texttt{googlePlayRating(App)} : Int]
+    Used to get the review score (out of 5) from the Google Play store.
+    This could be extended to other reviewers.
+
+  \item[tool\texttt{Check(App, Config)} : Bool]
+    Uses the \emph{tool} to run a check on an app.
+    This allows static analysis tools to run inferences on apps on demand.
+    Static analysis can be expensive for devices to perform.
+    An alternative would be to query a database of results~\citep{Noel:2014vt}.
+
+  \item[tool\texttt{CheckEvidence(App, Config, Evidence)} : Bool]
+    Digital evidence and proof~carrying~code~\citep{Necula:1996tr,Stark:2009uc} allow you to check an analysis result without repeating it.
+    In a mobile phone this can be preferable to doing the inference as draining batteries to perform checks is undesirable.
+    Tools, such as \emph{evicheck}~\citep{Noel:2014vt}, can produce digital evidence for Android apps.
+    This constraint function allows us to incorporate these tools into AppPAL policies.
+\end{description}
+
+\subsection{Evaluation}
+
+To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
+We do not use the DatalogC~\citep{Li:2003ix}~based translation and evaluation algorithm suggested by \citeauthor{Becker:2006vh}.
+Rather we implement the rules directly in Java. 
+Brief pseudocode is given in \autoref{fig:pseudocode}
+Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
+We also store intermediate results to avoid re-computation.
+On a mobile device memory is at a premium.
+We would like to keep the context as small as possible.
+To do this we allow some assertions to be imported at evaluation-time.
+
+\begin{figure}
+\[\footnotesize
+% AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase 
+% ----------------------------------------------------------- (can-act-as)
+%                 AC, D ⊧ A says B verbphrase
+\infer[\text{\tiny can-act-as}]{
+  AC, D \models \text{\tt A says B verbphrase}}{%
+  AC, D \models \text{\tt A says B can-act-as C} & %
+  AC, D \models \text{\tt A says C verbphrase}} 
+\]
+\[\footnotesize
+% AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
+% ----------------------------------------------------- (can-say)
+%                   AC, ∞ ⊧ A says fact
+\infer[\text{\tiny can-say}]{
+  AC, \infty \models \text{\tt A says fact}}{%
+  AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
+  AC, D \models \text{\tt B says fact}} 
+\]
+\[\footnotesize
+% (A says fact if fact₁, …, factₖ, c) ∈ AC
+% AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
+% ----------------------------------------------------------- (cond)
+%                    AC, D ⊧ A says factθ
+\infer[\text{\tiny cond}]{
+  AC, D \models \text{\tt A says fact$\theta$}}{
+  \begin{array}{c}
+    (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
+    \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
+  \end{array} &
+  \models \mathtt{c}\theta &
+  \text{\sl vars}(\text{\tt fact$\theta$})}
+\]
+\caption{AppPAL's evaluation rules.}
+\label{fig:rules}
+\end{figure}
+
+This gives us greater control of the evaluation and how the assertion context is created.
+For example, when checking the \code{isAnApp} predicate; 
+  we can fetch the assertion that the subject is an app based on the app in question.
+Similarly when we use a statement from \emph{Bob} that \emph{Google-Play can-say} whether an app is purchasable;
+  it is sensible to go fetch from the store whether the app is salable and make Google say it then and there.
+
+In SecPAL there is no notion of type.
+All entities are either constants (effectively strings) or variables.
+There is no differentiation made between a \emph{voiced} constant (one who speaks),
+and a \emph{subject} (one that is the subject of a verb-phrase, or predicate argument).
+In AppPAL the subjects are usually apps or policies.
+Apps do not utter facts.
+Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
+Similarly if a constant never appears as a subject of a statement, as most speakers don't,
+  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
+We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
+This allows us to cut the time spent searching for proofs.
+
+\begin{figure}
+\begin{minipage}[b]{0.5\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
+# ac, d |-? q (using previous-results table rt)
+def evaluate(ac, rt, q, d)
+  return rt[q, d] if rt.contains q, d
+  p = cond(ac, rt, q, d)
+  return (Proven, rt.update q, d, p) if p.isValid 
+  p = canSay_CanActAs(ac, rt, q, d)
+  return (Proven, rt.update q, d, p) if p.isValid 
+  return (Failure, rt.update q, d, Failure)
+
+# Try can-say and can-act-as
+def canSay_CanActAs(ac, rt, q, d)
+  ac.constants.each do |c| 
+    if c.is_a :subject
+      p = canActAs ac, rt, q, d
+      return Proven if p.isValid
+    elsif c.is_a :speaker
+      p = canSay ac, rt, q d
+      return Proven if p.isValid
+  return Failure
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[b]{0.5\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
+# Try cond rule
+def cond(ac, rt, q, d)
+  ac.add q.fetch if q.isFetchable
+  ac.assertions.each do |a|
+    if (u = q.unify a.consequent) &&
+       (a = u.substitution a).variables == none
+      return checkConditions ac, rt, a, d
+  return Failure
+
+# Evaluate the antecedents
+def checkConditions(ac, rt, a, d)
+  getVarSubstitutions(a, ac.constants).each do |s|
+    sa = s.substitute a
+    if sa.antecedents.all 
+        { |a| evaluate(ac, rt, a, d).isValid }
+      p = evaluateC sa.constraint
+      return Proven if p.isValid
+  return Failure
+\end{lstlisting}
+\end{minipage}
+\caption{Partial-pseudocode for AppPAL evaluation.}
+\label{fig:pseudocode}
+\end{figure}
+
+\subsection{Policy Idioms}
+\label{ssec:idioms}
 

### 2015-04-16 paper.tex
first draft

@@
 
-\title{AppPAL for Android}
-\subtitle{Capturing and Checking App Installation Policies}
+\title{AppPAL for Android\\
+Capturing and Checking App Installation Policies}
 \author{Joseph Hallett \and David Aspinall}
@@
   %   \item \comment{Sentence 1: State the problem}
-  When a user wants to pick an app on Android they are forced to judge apps based only on the information presented to them by the store.
+  When a user buys an app on Android they can only judge available apps based only on the information presented to them by the store.
   %   \item \comment{Sentence 2: Why is it an interesting problem}
   Employers must rely on employees using company devices properly to enforce security policies about apps at work.
-  Whilst some users spend time to pick apps with care others don't and become frustrated at apps that steal personal information.
+  Some users spend time to pick apps with care others don't.
+  This can cause frustration when they realise what data the apps have access to.
   %   \item \comment{Sentence 3: What does my solution achieve}
-  We present AppPAL: a policy language that can enforce app installation policies using statements from trusted third parties and delegation relationships.
+  We present AppPAL: a policy language that can enforce app installation policies.
+  Policies use statements from third parties and delegation relationships.
   %   \item \comment{Sentence 4: What follows from my solution}
-  This gives us a rigorous and flexible framework for enforcing and comparing and app policies and the trust relationships surrounding them.
+  This gives us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
   % \end{itemize}
@@
 App stores give some information about their apps; such as permissions and review scores.
-Ultimately the decision which apps to use must be made by the user, however.
+Android apps show a confusing list of permissions (\autoref{fig:facebook}).
+Whilst improving, users don't all understand how this relates to their device~\citep{Felt:2012hm,Thompson:2013eb}.
+Ultimately the decision which apps to use must be made by the user.
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.3\linewidth]{figures/facebook.png}
+  \caption{Security information presented when a user downloads the Facebook app.}
+\label{fig:facebook}
+\end{figure}
+
 %\comment{My contribution is\ldots}
@@
 
+Specifically we have:
 \begin{itemize}
-  %\item \comment{Specifically I have}
-  \item Described several scenarios where an authorization logic could replace informal policies for companies and users (\autoref{sec:problem}.) 
-    \comment{Provide forward refs saying where we do this and about what}
-  \item Shown how app installation policies can be expressed in AppPAL
-    including how a corporate \ac{byod} policy could be written, and how a user's personal security policy could be formalized (\autoref{sec:idea}.)
+  \item 
+    Described a scenario where a user has a policy they want to enforce for their employees~(\autoref{sec:problem}.) 
+    We show how the trust relationships in a system can be difficult to see;
+      and how AppPAL makes the trust between users explicit.
+  \item 
+    Shown how the user's policy could be implemented using AppPAL~(\autoref{sec:idea}); 
+      and describe what AppPAL would check for an app to meet this policy.
+
   \item Implemented the AppPAL language.  
-    Our implementation runs on Android and the JVM (\autoref{sec:details}.)
+    Our implementation runs on Android (\autoref{sec:details}.)
+    We show how we have tailored the language to Android (\autoref{ssec:language}),
+      how we evaluate it (\autoref{ssec:eval}),
+      and some idiomatic policies (\autoref{ssec:idioms}.)
 \end{itemize}
 
-\section{The Problem\comment{1 page}}
+\section{The problem\comment{1 page}}
 \label{sec:problem}
@@
   \includegraphics{figures/bob-policy.eps}
-  \caption{Pictoral representation of Alice's app install policy.}
+  \caption{Pictorial representation of Alice's app install policy.}
 \label{fig:bobs_policy}
@@
 In \autoref{lst:corporate_1} Alice gives Bob the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
-She allows him to delegate the decision further if he choses (\code{can-say inf}).
+She allows him to delegate the decision further if he chooses (\code{can-say inf}).
 Next in \autoref{lst:corporate_2} Bob specifies his concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
@@
 \subsection{The Language}
+\label{ssec:language}
 
@@
     We can specify more detailed checks using the \code{can-act-as} statements.
+
+  \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
+    A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
 \end{description}
@@
     Gets the version number of an app.
+
+  \item[\texttt{released(App)} : Date]
+    Checks when the app was last updated and released.
 \end{description}
@@
 
-\subsection{Evaluation}
+subsection{Evaluation}
 
@@
 Rather we implement the rules directly in Java. 
-Brief pseudocode is given in \autoref{fig:pseudocode}
+Brief pseudo-code is given in \autoref{fig:pseudocode}
 Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
@@
 Similarly when we use a statement from \emph{Bob} that \emph{Google-Play can-say} whether an app is purchasable;
-  it is sensible to go fetch from the store whether the app is salable and make Google say it then and there.
+  it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
 
@@
 We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
-This allows us to cut the time spent searching for proofs.
 
@@
 
-\subsection{Policy Idioms}
+\subsection{Policy Examples}
 \label{ssec:idioms}
+
+In this section we give examples of policies and queries and show how they can be implemented in AppPAL.
+
+\begin{description}
+  \item[App white-listing by an IT department]
+    Inside Bob's company employees in the IT department may white-list certain apps as installable.
+    To do this Bob delegates to them to decide if something is installable.
+    \begin{lstlisting}
+"bob" says "it-department" can-say 0 App isInstallable.
+    \end{lstlisting}
+    Charlie and Diveena work in the IT department.
+    When they have checked an app they say it is installable.
+    \begin{lstlisting}
+"bob" says "charlie" can-act-as "it-department".
+"bob" says "diveena" can-act-as "it-department".
+
+"charlie" says "com.facebook.katana" isInstallable.
+"diveena" says "org.thoughtcrime.securesms" isInstallable.
+    \end{lstlisting}
+
+  \item[Overpriviliged applications]
+    Alice is particularly worried about apps stealing her data.
+    She knows that certain apps need certain permissions.
+    For example a photography app needs access to the camera.
+    She checks each app's permissions carefully for things that seem unusual.
+
+    Alice's policy can be implemented by checking the permissions of the app, using constraints.
+    \begin{lstlisting}
+"alice" says App isInstallable
+   if "permissionsPolicy" isMetBy(App).
+"alice" says "permissionsPolicy" isMetBy(App)
+   if App isAnApp
+   where
+      category(App, "Photography"),
+      hasPermission(App, "ACCESS_COARSE_LOCATION") = False,
+      hasPermission(App, "CAMERA") = True.
+    \end{lstlisting}
+
+    Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
+    Alternately tools like Stowaway~\citep{Felt:2011kj} could be integrated to check apps.
+
+  \item[Digital Evidence]
+    In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
+    Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
+    For example that no audio can be recorded without the users consent~\citep{Seghir:2014uq}.
+    Using this tool someone generates a proof certificate that this property is met.
+    This could be the developer of the app, a third party certification service, or someone else;
+      when we check the proof we will find out if it is valid or not.
+    \begin{lstlisting}
+"alice" says Anyone can-say 0 
+  "recording-consent" shownIsMetBy(Evidence, App).
+
+"skb" says "recording-consent" 
+  shownIsMetBy("evicheck://evidence123", App).
+
+"alice" says "recording-consent" isMetBy(App)
+  if "recording-consent" shownIsMetBy(Evidence, App)
+  where
+    evicheckCheckEvidence(Evidence, App) = True. 
+    \end{lstlisting}
+    In the example here Alice allows anyone to present her with evidence for her policy.
+    A \ac{skb} offers an some evidence, and Alice accepts it as if it checks out.
+\end{description}
 
@@
 
+There has been a great amount of work on developing app analysis tools for Android.  
+Tools such as Stowaway~\citep{Felt:2011kj} detect overprivileged apps.
+TaintDroid~\citep{Enck:2010uw} and FlowDroid~\citep{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
+Other tools like QUIRE~\citep{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
+ScanDAL~\citep{Kim:2012vt} and SCanDroid~\citep{Fuchs:2009vi} help detect privacy leaks.
+Appscopy~\citep{Feng:kPGZr_ja} searches for specific kinds of malware.
+Tools like DroidRanger~\citep{Zhou:2012tb} scan app markets for malicious apps. 
+Many others exist checking and certifying other aspects of app behaviour.
+
+One tool, Kirin~\citep{Enck:2009ko}, also created a policy language and tool for enforcing app installation policies.
+Kirin's policies were concerned with preventing malware.
+Policy authors could specify combinations of permissions that should not appear together.
+For example an author might wish to stop malware sending premium rate text messages.
+To do this they restrict an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions (\autoref{fig:kirin}).
+Using this approach they found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
+\begin{figure}
+\begin{lstlisting}
+restrict permission [SEND_SMS] and permission [WRITE_SMS]
+\end{lstlisting}
+\begin{lstlisting}
+"user" says "no-write-send-sms" isMetBy(App)
+  where hasPermission(App, "SEND_SMS") = False.
+"user" says "no-write-send-sms" isMetBy(App)
+  where hasPermission(App, "WRITE_SMS") = False.
+\end{lstlisting}
+\caption{Kirin and AppPAL policies for stopping apps monetized by premium rate text messages.}
+\label{fig:kirin}
+\end{figure}
+
+Whilst this approach is great for malware, it is less suitable for finding \ac{pus}.
+The behaviours and permissions \ac{pus} apps display aren't necessarily malicious.
+One user may consider apps which need in-app-purchases to play malware, but another may enjoy them.
+AppPAL tries to stop these \ac{pus} apps.
+Because we can use external checking tools which go further than permissions checks, our policies can be richer.
+By allowing delegation relationships we can understand the provenance and trust relationships in these rules.
+
+Work by \citeauthor{Sadeh:2014vq} looked at modelling users' willingness to grant different kinds of permissions~\citep{Sadeh:2014vq}.
+They found four clusters of different users each with different kinds of policies:
+  conservatives who don't like granting many permissions,
+  advanced users who seem to have more complex policies,
+  unconcerned users who'll comfortably grant anything,
+  and fence-sitters (about half of all users) who seemed not to really care the majority of the time.
+This suggests that users are applying policies and do care how apps behave.
+Further more \citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
+This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
+Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
+
+We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
+Further more attempts to apply it to other areas had already succeeded~\citep{Aziz:2011vt,Humphrey:2007wc}.
+Other policy languages such as RT~\citep{Li:2003ua}, or Cassandra~\citep{Becker:2004il} might also have made a good starting point.
+However neither had the precise set of features we we're looking for (namely explicit speakers and external checking functions).
+Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
+DKAL also contains more modalities than \emph{says}.
+
 \section{Conclusions and further work \comment{$\frac{1}{2}$ page}}
+
+We have presented AppPAL: an authorization logic for describing app installation policies.
+The language is implemented in Java and runs on Android using a custom evaluation algorithm.
+This allows us to enforce app installation policies on Android devices.
+We've shown how the language can be used to describe an app installation policy;
+  and given brief descriptions of how other policies might be described.
+
+Further work is required to tightly integrate AppPAL into Android.
+The obvious place to put it would be as a \emph{required checker}: a program that checks all apps before installation.
+Google currently use this API to check for known malware and jailbreak apps.
+We would like AppPAL to sit there and check apps on the basis of policies.
+Unfortunately the API is protected and it would require the phone to be rooted to run there.
+
+Developing, and testing, policies for users is a key next step.
+Here we described a policy being specified by a company boss.
+For most end-users writing a policy in a formal language is too much work.
+Ad-blocking software works by users subscribing to filter policies written by experts\footnote{EasyList is a popular choice and the default in most ad-blocking software.
+They offer many different policies for specific use-cases however.
+\url{https://easylist.adblockplus.org/en/}}.
+We can imagine a similar scheme working well for app installation policies.
+Users subscribe to different policies by experts (e.g. no tracking apps, nothing with adult content, no spammy in-app-purchase apps).
+Optionally they can customize them further.
+
+We should also attempt to learn policies from existing users behavior.
+Given app usage data, from a project like Carat~\citep{Oliner:2013ht}, we could identify security conscious users.
+If we can infer these users policies we may be able to describe new policies that the less technical users may want.
+Given a set of apps one user has already installed, we could learn policies about what their personal installation policy is.
+This may help stores show users apps they're more likely to buy, and users apps that already behave as they want.
+
+AppPAL is a powerful language for describing app installation policies.
+It gives us a framework for describing and evaluating policies for Android apps.
+The work provides new ways for users to enforce their own rules about how apps should behave.
+Users policies can be enforced more reliably, and with less interaction;
+thus making apps more pleasant and safer for everyone.
 

### 2015-05-21 paper.tex
Updates

@@
   %   \item \comment{Sentence 1: State the problem}
-  When a user buys an app on Android they can only judge available apps based only on the information presented to them by the store.
+  Users must judge apps based only on the information presented to them by the store.
   %   \item \comment{Sentence 2: Why is it an interesting problem}
-  Employers must rely on employees using company devices properly to enforce security policies about apps at work.
+  Employers rely on employees using company devices appropriately to enforce security policies at work.
   Some users spend time to pick apps with care others don't.
-  This can cause frustration when they realise what data the apps have access to.
+  This may cause frustration when they realise what data the apps have access to.
   %   \item \comment{Sentence 3: What does my solution achieve}
@@
 %\comment{The problem is\ldots}
-How do you pick which apps to install on your phone?  
+Finding good apps can be tricky.
+Users need to work out which apps are well written.
+They also need to find the apps which suit how they want to use their phone.
+This can be difficult.
+It isn't obvious how apps use the data each has access to.
+
 App stores give some information about their apps; such as permissions and review scores.
 Android apps show a confusing list of permissions (\autoref{fig:facebook}).
-Whilst improving, users don't all understand how this relates to their device~\citep{Felt:2012hm,Thompson:2013eb}.
-Ultimately the decision which apps to use must be made by the user.
+Users don't all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
+Ultimately the decision which apps to use must be made by a human.
 
 Not all apps are suitable. 
-There is a large amount of \ac{pus} available for Android~\citep{Truong:2014bi,Svajcer:2013tp}. 
+A large amount of \ac{pus} is being propagated for Android devices~\citep{Truong:2014bi,Svajcer:2013tp}. 
 Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
-Their company may want to restrict which apps their employees can use.
+A employers may wish to restrict which apps their employees can use.
 The \ac{ciso} or IT department may set a policy to prevent information leaks.
-Some users worry apps will misuse their personal data.
-They avoid apps which can access their location, or contact data.
+Some users worry apps will misuse their personal data;
+  such a user avoids apps which can access their location, or address book.
 They apply their own personal security policy when downloading apps.
 
-Currently these policies are enforced by hand.
-This is error-prone and can be inconsistent.
-A better solution is to use a logic of authorization.
-The policy can be written in the logic and enforced automatically.
+These policies are enforced by hand, if at all.
+This is error-prone and mistakes can be made.
+We believe this can be bettered.
+An alternative would be to write the policy down and make the computer do it.
+To implement this we can use a logic of authorization.
+The policy may be written in the logic and enforced automatically.
 
@@
 The language is an instantiation of SecPAL~\citep{Becker:2006vh} tailored to making statements about apps.
-It allows us to reason about apps using statements from third parties.
-We can express trust relationships in these speakers and use constraints to perform additional checks at evaluation time.
-This allows us to enforce deeper and more complex policies than existing policy tools like Kirin~\citep{Enck:2009ko}.
+The language allows us to reason about apps using statements from third parties.
+The implementation allows us to enforce the policies on a device.
+We can express trust relationships amongst these parties; use constraints to perform additional checks.
+This lets us enforce deeper and more complex policies than existing tools such as Kirin~\citep{Enck:2009ko}.
 
-Specifically we have:
+Using AppPAL we can make statements about apps:
+\begin{lstlisting}
+"Alice" says "com.rovio.angrybirds" isInstallable.
+\end{lstlisting}
+We can delegate decisions to others:
+\begin{lstlisting}
+"Alice" says "Bob" can-say 0 App isCool.
+\end{lstlisting}.
+We can run analysis tools, and solve constraints:
+\begin{lstlisting}
+"Claire" says App isMalware
+  if App isAnApp
+  where
+    virusScanner(App) = True.
+\end{lstlisting}
+
+Specifically we:
 \begin{itemize}
@@
       and some idiomatic policies (\autoref{ssec:idioms}.)
+
+  \item When things go wrong we also need to attribute blame.
+    Authorization logic can help make trust relationships explicit.
+    In \autoref{ssec:language} we show a recent example of a broken trust relationship,
+    and how our authorization logic makes the problem explicit.
+
 \end{itemize}
 
-\section{The problem\comment{1 page}}
+\section{Enforcing a policy at work}
 \label{sec:problem}
 
-Suppose \emph{Alice} works for \emph{Bob}.
-Bob allows Alice to use her personal phone as a work phone, but he has some specific concerns.
+Suppose \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
+Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
 \begin{itemize}
@@
     The testing labs are at a secret location and it mustn't be leaked.
-  \item Apps should come from a reputable source, such as the Play store.
-  \item Bob uses an \ac{av} program by McAfee. 
+  \item Apps should come from a reputable source, such as the Google Play Store.
+  \item Emma uses an \ac{av} program by McAfee. 
     It should check all apps before they're installed.
@@
 
-To ensure this policy is fully met Alice could promise to follow it; but this is error-prone---what if she makes a mistake or misses something?
-Bob's policy could be implemented using existing tools.
-\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
-  could configure Alice's device to disallow apps from outside the Play Store.
-Various tools~\citep{Backes:2012vm,Jeon:2012ki,Hornyack:2011wq} can control the permissions or data an app can get.
+To ensure this policy is fully met Alice could promise to follow it.
+Perhaps she might even sign a document promising to do so.
+This is error-prone---what if she makes a mistake or misses something?
+Emma's policy could be implemented using existing tools.
+The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
+  could configure Alice's device to disallow apps from outside the Google Play Store.
+Various tools such as AppGuard~\citep{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\citep{Jeon:2012ki} or AppFence~\citep{Hornyack:2011wq} can control the permissions or data an app can get.
 These could be used to used to ensure no location data is ever obtained.
-Alternately other tools~\citep{Enck:2009ko,Fritz:2013vi,Gordon:2015et} could check that the locations are ever leaked to the web.
-Various anti-virus programs are available for Android---one could be installed on Alice's phone.
+Alternately other tools like Kirin~\citep{Enck:2009ko}, Flowdroid~\citep{Fritz:2013vi} or DroidSafe~\citep{Gordon:2015et} could check that the locations are ever leaked to the web.
+Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.
 
-Whilst we \emph{could} implement Bob's policy using existing tools, it isn't a great solution.
-It isn't flexible, Alice needs to alter and remove the software on her phone if Bob changes his policy or she changes jobs.
-It isn't clear what an app must do to be installed, or what checks must have been done if it already has been put on the phone.
-The link between Alice, the user; Bob, the policy setter; and the tools Bob trusts to implement his policy isn't immediately apparent.
+\begin{figure}
+\includegraphics{figures/overview.eps}
+\caption{Ecosystem of devices and stores with AppPAL.}
+\label{fig:ecosystem}
+\end{figure}
 
-This last point is particularly pertinent with respect to \textsc{CVE-2015-2077}.
-In 2015 Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
-Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
-Unfortunately they had configured it to ignore Superfish\footnote{The malware was an ad framework designed to show users products similar to those they viewed on the web.
-Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.}.
+Whilst we \emph{could} implement Emma's policy using existing tools, it isn't a great solution.
+It isn't flexible, Alice needs to alter and remove the software on her phone if Emma changes her policy or she changes jobs.
+It isn't clear what an app must do to be installed, or what checks have been done if it already has been put on the phone.
+The link between Alice, the user; Emma, the policy setter; and the tools Emma trusts to implement her policy isn't immediately apparent.
 
-From a user's perspective there was a delegation of trust to Lenovo to detect malware.
-Lenovo then delegated further to McAfee to supply the antivirus; but left an exception that they would allow Superfish.
-A user might wonder where the breach of trust occurred and who is to blame?  
-Is it the \ac{av} failing to spot the malware? 
-Has someone else configured something incorrectly?
-How should they fix the problem?
+Our research looks at the problem of picking software.
+Given there are some apps you want and others you don't:
+  how can you express your preferences in such a way that they can be enforced automatically?
+How can we translate policy documents from natural language into something machine checkable?
+Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?
 
-If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
-\begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
-"user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
-"lenovo" says "mcafee" can-say inf File isSafe.
-"lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
-\end{lstlisting}
-The fault lies with \autoref{lst:lenovo_unsafe}.  
-Lenovo has caused the problem (they \emph{said} it was safe).
-The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
-
-It is important to stress that we are not trying to solve the problem of detecting malware.
-There are many tools, and much research, that already does this.
-Our problem is different. 
-Given there are some apps you want and others you don't:
-  how can you express your preferences in such a way that they can be enforced automatically.
-Further more how can we show the trust relationships used to make these decisions clearly and precisely.
-
-\section{My idea \comment{2 pages}}
+\section{Expressing Policies in AppPAL}
 \label{sec:idea}
 
-Returning to Alice and Bob's dilemma in \autoref{sec:problem}, we propose a solution.
-Instead of using several different tools to enforce Bob's policy disjointedly, we use an authorization logic.
-In \autoref{lst:corporate} we give an AppPAL policy implementing Bob's app concerns on Alice's phone.
-A pictorial representation, is given in \autoref{fig:bobs_policy}; color is used to split statements by each speaker.
+Returning to Alice and Emma's dilemma in \autoref{sec:problem}, we propose a solution.
+Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
+In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
+A pictorial representation, is given in \autoref{fig:emmas_policy}; color is used to split statements by each speaker.
 
@@
   \caption{Pictorial representation of Alice's app install policy.}
-\label{fig:bobs_policy}
+\label{fig:emmas_policy}
 \end{figure}
@@
 
-\begin{lstlisting}[label={lst:corporate}, 
-                   caption={An AppPAL policy implementing Bob's security requirements},
-                   float,
-                   numbers=left,
+\begin{figure}
+\begin{lstlisting}[numbers=left,
                    escapeinside={@}{@}]
-"alice" says "bob" can-say inf App isInstallable. @\label{lst:corporate_1}@
-"bob" says App isInstallable @\label{lst:corporate_2}@
+"alice" says "emma" can-say inf App isInstallable. @\label{lst:corporate_1}@
+"emma" says App isInstallable @\label{lst:corporate_2}@
   if "no-tracking-policy" isMetBy(App),
@@
      "anti-virus-policy" isMetBy(App).
-"bob" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
-  if App isBuyable.
-"bob" says "google-play" can-say 0 App isBuyable. @\label{lst:corporate_4}@
-"bob" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
+"emma" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
+  if App isReputable.
+"emma" says "google-play" can-say 0 App isReputable. @\label{lst:corporate_4}@
+"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
   if App isAnApp
@@
     mcAfeeVirusCheck(App) = False.
-"bob" says "no-location-permissions" can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
-"bob" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
+"emma" says "no-location-permissions" can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
+"emma" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
   if App isAnApp
@@
 \end{lstlisting}
+\caption{AppPAL policy implementing Emma's security requirements}
+\label{lst:corporate}
+\end{figure}
 
-In \autoref{lst:corporate_1} Alice gives Bob the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
-She allows him to delegate the decision further if he chooses (\code{can-say inf}).
-Next in \autoref{lst:corporate_2} Bob specifies his concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
-If Bob can be convinced all these policies are met then he will say the \code{App isInstallable}.
-In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Bob specifies that an app meets the \code{reputable-source-policy} if the \code{App isBuyable};
+In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
+She allows her to delegate the decision further if she chooses (\code{can-say inf}).
+Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
+If Emma can be convinced all these policies are met then he will say the \code{App isInstallable}.
+In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-source-policy} if the \code{App isReputable};
   with \code{"google-play"} specified as the decider of what is buyable or not.
@@
 In other words Google is not allowed to specify Amazon as a supplier of apps as well.
-Google must say what is buyable directly for Bob to accept it.
-Bob specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
+Google must say what is buyable directly for Emma to accept it.
+Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
 Here we use a constraint.
@@
 Only if this returns \code{False} will the policy be met.
-To specify the \code{"no-tracking-policy"} Bob says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
-Bob specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.
+To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
+Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.
 
 Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
-To meet Bob's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isInstallable} predicate.
+To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isInstallable} predicate.
 Specifically it needs:
-\begin{description}
-  \item[\texttt{"bob" says "com.facebook.katana" isAnApp.}] 
+\begin{itemize}
+  \item\code{"emma" says "com.facebook.katana" isAnApp.} 
     A simple typing statement that can be generated for all apps as they are encountered.
-  \item[\texttt{"google-play" says "com.facebook.katana" isBuyable.}] 
-    Required to convince Bob the app came from a reputable source. 
+  \item\code{"google-play" says "com.facebook.katana" isReputable.} 
+    Required to convince Emma the app came from a reputable source. 
     It should be able to obtain this statement from the Play store as the app is available there.
-  \item[\texttt{"bob" says "anti-virus-policy" isMetBy("com.facebook.katana").}]
+  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
     She can obtain this by running the \ac{av} program on her app.
-  \item[\texttt{"bob" says "no-locations-permissions" isMetBy("com.facebook.katana").}]
-    Needed to show the App meets Bob's no-tracking-policy.
-    Bob will say this if after examining the app the location permissions are missing.
-\end{description}
+  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
+    Needed to show the App meets Emma's no-tracking-policy.
+    Emma will say this if after examining the app the location permissions are missing.
+\end{itemize}
 These last two statements require the checker to do some extra checks to satisfy the constraints.
@@
 In this scenario we have imagined Alice wanting to check the apps as she installs them.
-Alternatively we could imagine Bob wanting a personalised app store where all apps sold meet his policy.
+Alternatively we could imagine Emma wanting a personalised app store where all apps sold meet her policy.
 With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which meet the policy.
+This gives us a \emph{filtered store}.
+From an existing set of apps we produce a personalised store that meets a pre-defined policy.
 
-\section{The details \comment{5 pages}}
+\section{AppPAL}
 \label{sec:details}
@@
 \begin{description}
+  \item[\texttt{App isInstallable}]
+    Used to indicate that an App meets the install policy for the device.
+    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
+
   \item[\texttt{App isAnApp}, \texttt{Policy isAPolicy}\ldots] 
@@
 
-  \item[\texttt{App isInstallable}]
-    Used to indicate that an App meets the install policy for the device.
-    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
-
   \item[\texttt{Policy isMetBy(App)}]
     Policies let you split app behaviour into sub-policies.
-    For example in \autoref{sec:idea} we showed how Bob's installation policy could be written using three \texttt{isMetBy} statements.
+    For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
     Splitting policies allows greater control of how each one is checked.
@@
 
-We also add the following constraint functions which check simple properties of the app:
-In every case the check can be done by examining the app itself.
-\begin{description}
-  \item[\texttt{hasPermission(App, String)} : Bool]
-    Checks the permissions requested in the app's manifest to see if the string is included in them.
+Splitting the decision as to whether an app is installable into a series of policies that must be met gives us flexibility in how the decision is made.
+It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
+Some static analysis tools are not quick to run.
+Even taking minutes to run an expensive (in terms of battery power) analysis can be undesirable.
+If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
 
-  \item[\texttt{hasPackage(App, String)} : Bool]
-    Checks the app to see if a package called \emph{String} is included within it.
-    Examples might include aggressive advertisement frameworks a user wishes to ban.
+Consider the \emph{no-tracking-policy} from earlier.
+In Bob's policy we checked this by checking the permissions of the app.
+If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
+Some apps may want to access this data, but may not leak it.
+We could use a taint analysis tool to detect this (e.g.~Taintdroid~\citep{Fritz:2013vi}).
+Our policy now becomes:
 
-  \item[\texttt{hasNativeCode(App)} : Bool]
-    Checks to see whether the app contains native code.
-    Native (i.e. non-Dalvik) code is potentially more dangerous than the sandboxed Dalvik versions.
-    It is typically used for performance sensitive code (i.e. games).
+\begin{lstlisting}
+"bob" says "no-locations-permissions" can-act-as "no-tracking-policy".
+"bob" says "no locations-permissions" isMetBy(App)
+  if App isAnApp
+  where
+    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
+    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.
 
-  \item[\texttt{isSignedBy(App, Speaker)} : Bool]
-    Checks the \texttt{CERT.RSA} used to sign the App and compares it with the certificate used to identify a speaker.
+"bob" says "location-taint-analysis" can-act-as "no-tracking-policy".
+"bob" says "location-taint-analysis" isMetBy(App)
+  if App isAnApp
+  where
+    taintDroidCheckLeak(App, "Location", "Internet") = False.
+\end{lstlisting}
 
-  \item[\texttt{hasVersion(App)} : Int]
-    Gets the version number of an app.
+Sometimes we might want to use location data.
+For instance Bob might want to check that Alice is at her office.
+Bob might use a location tracking app to do this.
+Provided the app only talks to Bob, and it uses SSL correctly (something Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.
 
-  \item[\texttt{released(App)} : Date]
-    Checks when the app was last updated and released.
-\end{description}
+\begin{lstlisting}
+"bob" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
+"bob" says "relaxed-no-tracking-policy" isMetBy(App)
+  if App hasCategory("tracking")
+  where 
+    mallodroidSSLCheck(App) = False,
+    connections(App) = "[https://bob.com]".
+\end{lstlisting}
 
-We also add some constraints that do more complex checks:
-\begin{description}
-  \item[\texttt{googlePlayRating(App)} : Int]
-    Used to get the review score (out of 5) from the Google Play store.
-    This could be extended to other reviewers.
+This gives us four different ways of satisfying the \emph{no-tracking-policy}:
+  with permissions, 
+  with taint analysis,
+  with a relaxed version of the policy,
+  or by Bob directly saying the app meets it.
+When we come to check the policy if any of these ways give us a positive result we can can stop our search.
 
-  \item[tool\texttt{Check(App, Config)} : Bool]
-    Uses the \emph{tool} to run a check on an app.
-    This allows static analysis tools to run inferences on apps on demand.
-    Static analysis can be expensive for devices to perform.
-    An alternative would be to query a database of results~\citep{Noel:2014vt}.
+AppPAL also helps attribute blame when things go wrong.
+By modelling the trust relationships in systems we can work out precisely where mistakes were made.
+A recent example demonstrating this is \textsc{CVE-2015-2077}.
+Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
+Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
+Unfortunately they had configured it to ignore Superfish.
+The malware was an ad framework designed to show users products similar to those they viewed on the web.
+Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.
 
-  \item[tool\texttt{CheckEvidence(App, Config, Evidence)} : Bool]
-    Digital evidence and proof~carrying~code~\citep{Necula:1996tr,Stark:2009uc} allow you to check an analysis result without repeating it.
-    In a mobile phone this can be preferable to doing the inference as draining batteries to perform checks is undesirable.
-    Tools, such as \emph{evicheck}~\citep{Noel:2014vt}, can produce digital evidence for Android apps.
-    This constraint function allows us to incorporate these tools into AppPAL policies.
-\end{description}
+From a user's perspective there was a delegation of trust to Lenovo to detect malware.
+Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
+A user might wonder where the breach of trust occurred and who is to blame?  
+Is it the \ac{av} failing to spot the malware? 
+Has someone else configured something incorrectly?
+How should they fix the problem?
 
-subsection{Evaluation}
+If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
+\begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
+"user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
+"lenovo" says "mcafee" can-say inf File isSafe.
+"lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
+\end{lstlisting}
+The fault lies with \autoref{lst:lenovo_unsafe}.  
+Lenovo has caused the problem (they \emph{said} it was safe).
+The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
+
+%
+%We also add the following constraint functions which check simple properties of the app:
+%In every case the check can be done by examining the app itself.
+%\begin{description}
+%  \item[\texttt{hasPermission(App, String)} : Bool]
+%    Checks the permissions requested in the app's manifest to see if the string is included in them.
+%
+%  \item[\texttt{hasPackage(App, String)} : Bool]
+%    Checks the app to see if a package called \emph{String} is included within it.
+%    Examples might include aggressive advertisement frameworks a user wishes to ban.
+%
+%  \item[\texttt{hasNativeCode(App)} : Bool]
+%    Checks to see whether the app contains native code.
+%    Native (i.e. non-Dalvik) code is potentially more dangerous than the sandboxed Dalvik versions.
+%    It is typically used for performance sensitive code (i.e. games).
+%
+%  \item[\texttt{isSignedBy(App, Speaker)} : Bool]
+%    Checks the \texttt{CERT.RSA} used to sign the App and compares it with the certificate used to identify a speaker.
+%
+%  \item[\texttt{hasVersion(App)} : Int]
+%    Gets the version number of an app.
+%
+%  \item[\texttt{released(App)} : Date]
+%    Checks when the app was last updated and released.
+%\end{description}
+%
+%We also add some constraints that do more complex checks:
+%\begin{description}
+%  \item[\texttt{googlePlayRating(App)} : Int]
+%    Used to get the review score (out of 5) from the Google Play store.
+%    This could be extended to other reviewers.
+%
+%  \item[tool\texttt{Check(App, Config)} : Bool]
+%    Uses the \emph{tool} to run a check on an app.
+%    This allows static analysis tools to run inferences on apps on demand.
+%    Static analysis can be expensive for devices to perform.
+%    An alternative would be to query a database of results~\citep{Noel:2014vt}.
+%
+%  \item[tool\texttt{CheckEvidence(App, Config, Evidence)} : Bool]
+%    Digital evidence and proof~carrying~code~\citep{Necula:1996tr,Stark:2009uc} allow you to check an analysis result without repeating it.
+%    In a mobile phone this can be preferable to doing the inference as draining batteries to perform checks is undesirable.
+%    Tools, such as \emph{evicheck}~\citep{Noel:2014vt}, can produce digital evidence for Android apps.
+%    This constraint function allows us to incorporate these tools into AppPAL policies.
+%\end{description}
+
+\subsection{Evaluation}
+\label{ssec:eval}
 
@@
   we can fetch the assertion that the subject is an app based on the app in question.
-Similarly when we use a statement from \emph{Bob} that \emph{Google-Play can-say} whether an app is purchasable;
+Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is purchasable;
   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
@@
   \item[App white-listing by an IT department]
-    Inside Bob's company employees in the IT department may white-list certain apps as installable.
-    To do this Bob delegates to them to decide if something is installable.
+    Inside Emma's company employees in the IT department may white-list certain apps as installable.
+    To do this Emma delegates to them to decide if something is installable.
     \begin{lstlisting}
-"bob" says "it-department" can-say 0 App isInstallable.
+"emma" says "it-department" can-say 0 App isInstallable.
     \end{lstlisting}
@@
     \begin{lstlisting}
-"bob" says "charlie" can-act-as "it-department".
-"bob" says "diveena" can-act-as "it-department".
+"emma" says "charlie" can-act-as "it-department".
+"emma" says "diveena" can-act-as "it-department".
 
@@
 
+We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
+Further more attempts to apply it to other areas had already succeeded~\citep{Aziz:2011vt,Humphrey:2007wc}.
+Other policy languages such as RT~\citep{Li:2003ua}, or Cassandra~\citep{Becker:2004il} might also have made a good starting point.
+However neither had the precise set of features we we're looking for (namely explicit speakers and external checking functions).
+Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
+DKAL also contains more modalities than \emph{says}.
+
 There has been a great amount of work on developing app analysis tools for Android.  
@@
 
-We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
-Further more attempts to apply it to other areas had already succeeded~\citep{Aziz:2011vt,Humphrey:2007wc}.
-Other policy languages such as RT~\citep{Li:2003ua}, or Cassandra~\citep{Becker:2004il} might also have made a good starting point.
-However neither had the precise set of features we we're looking for (namely explicit speakers and external checking functions).
-Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
-\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
-DKAL also contains more modalities than \emph{says}.
-
 \section{Conclusions and further work \comment{$\frac{1}{2}$ page}}

### 2015-05-28 paper.tex
Updated formatting

@@
 % vim: set nowrap tw=0:
-\documentclass[a4paper]{scrartcl}
+\documentclass[twoside,letterpaper]{soups}
+
+\pdfpagewidth=8.5truein{}
+\pdfpageheight=11truein{}
+
+\usepackage{graphicx}
+\renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
+\renewcommand{\floatpagefraction}{0.99}
+\pagestyle{plain} % page numbers
+
 \usepackage{defaultstyle}
@@
 
-\title{AppPAL for Android\\
-Capturing and Checking App Installation Policies}
-\author{Joseph Hallett \and David Aspinall}
-\date{\today}
+\title{AppPAL for Android}
+\subtitle{Capturing and Checking App Installation Policies}
+\numberofauthors{2}
+\author{\alignauthor
+  Joseph Hallett \\ \email{J.Hallett@sms.ed.ac.uk}
+  \alignauthor
+  David Aspinall \\ \email{David.Aspinal@ed.ac.uk}}
 
 \begin{document}
-\maketitle
+
+\maketitle{}
 
@@
   Employers rely on employees using company devices appropriately to enforce security policies at work.
-  Some users spend time to pick apps with care others don't.
+  Some users spend time to pick apps with care others do not.
   This may cause frustration when they realise what data the apps have access to.
@@
 Android apps show a confusing list of permissions (\autoref{fig:facebook}).
-Users don't all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
+Users do not all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
 Ultimately the decision which apps to use must be made by a human.
@@
 Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
-A employers may wish to restrict which apps their employees can use.
+An employer may wish to restrict which apps their employees can use.
 The \ac{ciso} or IT department may set a policy to prevent information leaks.
@@
 The implementation allows us to enforce the policies on a device.
-We can express trust relationships amongst these parties; use constraints to perform additional checks.
+We can express trust relationships amongst these parties; use constraints to do additional checks.
 This lets us enforce deeper and more complex policies than existing tools such as Kirin~\citep{Enck:2009ko}.
@@
 Our research looks at the problem of picking software.
-Given there are some apps you want and others you don't:
+Given there are some apps you want and others you do not:
   how can you express your preferences in such a way that they can be enforced automatically?
@@
                    escapeinside={@}{@}]
-"alice" says "emma" can-say inf App isInstallable. @\label{lst:corporate_1}@
+"alice" says "emma" can-say inf
+  App isInstallable. @\label{lst:corporate_1}@
+
 "emma" says App isInstallable @\label{lst:corporate_2}@
@@
      "anti-virus-policy" isMetBy(App).
+
 "emma" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
   if App isReputable.
+
 "emma" says "google-play" can-say 0 App isReputable. @\label{lst:corporate_4}@
+
 "emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
@@
     mcAfeeVirusCheck(App) = False.
-"emma" says "no-location-permissions" can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
+
+"emma" says "no-location-permissions"
+  can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
+
 "emma" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
@@
   where
-    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
-    hasPermission(App, "ACCESS_COURSE_LOCATION") = False.
+    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
+    hasPermission(App, "ACCESS_COURSE_LOCATION") = false.
 \end{lstlisting}
@@
 \begin{lstlisting}
-"bob" says "no-locations-permissions" can-act-as "no-tracking-policy".
-"bob" says "no locations-permissions" isMetBy(App)
+"bob" says "no-locations-permissions"
+  can-act-as "no-tracking-policy".
+
+"bob" says "no-locations-permissions" isMetBy(App)
   if App isAnApp
@@
 
-"bob" says "location-taint-analysis" can-act-as "no-tracking-policy".
+"bob" says "location-taint-analysis"
+  can-act-as "no-tracking-policy".
+
 "bob" says "location-taint-analysis" isMetBy(App)
@@
   where
-    taintDroidCheckLeak(App, "Location", "Internet") = False.
+    taintDroidCheckLeak(App, "Location", "Internet")
+      = False.
 \end{lstlisting}
@@
 Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
-Similarly if a constant never appears as a subject of a statement, as most speakers don't,
+Similarly if a constant never appears as a subject of a statement, as most speakers do not,
   there is no need to check it as a possible subject for a \emph{can-act-as} statement.
@@
 They found four clusters of different users each with different kinds of policies:
-  conservatives who don't like granting many permissions,
+  conservatives who do not like granting many permissions,
   advanced users who seem to have more complex policies,

### 2015-06-02 paper.tex
Added to related work

@@
 % vim: set nowrap tw=0:
-\documentclass[twoside,letterpaper]{soups}
+\documentclass[]{scrartcl}
 
-\pdfpagewidth=8.5truein{}
-\pdfpageheight=11truein{}
+% \pdfpagewidth=8.5truein{}
+% \pdfpageheight=11truein{}
 
 \usepackage{graphicx}
-\renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
-\renewcommand{\floatpagefraction}{0.99}
-\pagestyle{plain} % page numbers
+% \renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
+% \renewcommand{\floatpagefraction}{0.99}
+% \pagestyle{plain} % page numbers
 
@@
 \subtitle{Capturing and Checking App Installation Policies}
-\numberofauthors{2}
-\author{\alignauthor
-  Joseph Hallett \\ \email{J.Hallett@sms.ed.ac.uk}
-  \alignauthor
-  David Aspinall \\ \email{David.Aspinal@ed.ac.uk}}
+% \numberofauthors{2}
+\author{Joseph Hallett \and David Aspinall }
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.3\linewidth]{figures/facebook.png}
-  \caption{Security information presented when a user downloads the Facebook app.}
-\label{fig:facebook}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.3\linewidth]{figures/facebook.png}
+%   \caption{Security information presented when a user downloads the Facebook app.}
+% \label{fig:facebook}
+% \end{figure}
 
@@
 \begin{lstlisting}
-"Alice" says "com.rovio.angrybirds" isInstallable.
+"Alice" says "com.rovio.angrybirds" isRunnable.
 \end{lstlisting}
@@
 "Alice" says "Bob" can-say 0 App isCool.
-\end{lstlisting}.
+\end{lstlisting}
 We can run analysis tools, and solve constraints:
@@
   where
-    virusScanner(App) = True.
+    virusScanner(App) = true.
 \end{lstlisting}
@@
 \begin{itemize}
-  \item Alice shouldn't have any apps that can track her movements.
+  \item Alice shouldn't run any apps that can track her movements.
     The testing labs are at a secret location and it mustn't be leaked.
@@
 Perhaps she might even sign a document promising to do so.
-This is error-prone---what if she makes a mistake or misses something?
+This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
 Emma's policy could be implemented using existing tools.
@@
 It isn't flexible, Alice needs to alter and remove the software on her phone if Emma changes her policy or she changes jobs.
-It isn't clear what an app must do to be installed, or what checks have been done if it already has been put on the phone.
+It isn't clear what an app must do to be run, or what checks have been done if it already has been put on the phone.
 The link between Alice, the user; Emma, the policy setter; and the tools Emma trusts to implement her policy isn't immediately apparent.
+
+What happens when Alice goes home?
+Emma shouldn't be able to overly control what Alice does in her private life.
+Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
+The some mobile OSs allow app permissions to be enabled and disable at run time.
+Can we enforce different policies at different times?
 
@@
   how can you express your preferences in such a way that they can be enforced automatically?
-How can we translate policy documents from natural language into something machine checkable?
+How can we translate policy documents from natural language into a machine checkable form?
 Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?
 
-\section{Expressing Policies in AppPAL}
+\section{Expressing policies in AppPAL}
 \label{sec:idea}
@@
 In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
-A pictorial representation, is given in \autoref{fig:emmas_policy}; color is used to split statements by each speaker.
+% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.
 
-\begin{figure}
-  \includegraphics{figures/bob-policy.eps}
-  \caption{Pictorial representation of Alice's app install policy.}
-\label{fig:emmas_policy}
-\end{figure}
+% \begin{figure}
+%   \includegraphics{figures/bob-policy.eps}
+%   \caption{Pictorial representation of Alice's app install policy.}
+% \label{fig:emmas_policy}
+% \end{figure}
 
-AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app installation policies.
+AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app running policies.
 SecPAL is a logic of authorization for access control decisions in distributed systems.
@@
 "alice" says "emma" can-say inf
-  App isInstallable. @\label{lst:corporate_1}@
+  App isRunnable. @\label{lst:corporate_1}@
 
-"emma" says App isInstallable @\label{lst:corporate_2}@
+"emma" says App isRunnable @\label{lst:corporate_2}@
   if "no-tracking-policy" isMetBy(App),
@@
   where
-    mcAfeeVirusCheck(App) = False.
+    mcAfeeVirusCheck(App) = false.
 
@@
 
-In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
+In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
 She allows her to delegate the decision further if she chooses (\code{can-say inf}).
 Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
-If Emma can be convinced all these policies are met then he will say the \code{App isInstallable}.
+If Emma can be convinced all these policies are met then he will say the \code{App isRunnable}.
 In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-source-policy} if the \code{App isReputable};
@@
 When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
-Only if this returns \code{False} will the policy be met.
+Only if this returns \code{false} will the policy be met.
 To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
@@
 Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
-To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isInstallable} predicate.
+To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
 Specifically it needs:
@@
 
-AppPAL is implemented as a library for Android.
-A standalone Java version also exists for testing.
+AppPAL is implemented as a library for Android and Java.
 The parser is implemented using ANTLR4.
-Code, tests, and prebuilt APKs are available from Github\footnote{\url{https://github.com/bogwonch/AppPAL}}.
+Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.
 
@@
 
-The structure of an AppPAL assertion is shown in \autoref{fig:assertion}.
+The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.
 
@@
       \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} &
-      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isInstallable}}^{\text{predicate}}}^{\text{fact}} \\
+      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isRunnable}}^{\text{predicate}}}^{\text{fact}} \\
       & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
-      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = True}}^{\text{constraint}}.
+      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = true}}^{\text{constraint}}.
     \end{array}
@@
 \begin{description}
-  \item[\texttt{App isInstallable}]
+  \item[\texttt{App isRunnable}]
     Used to indicate that an App meets the install policy for the device.
@@
 
-Splitting the decision as to whether an app is installable into a series of policies that must be met gives us flexibility in how the decision is made.
+Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
 It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
 Some static analysis tools are not quick to run.
-Even taking minutes to run an expensive (in terms of battery power) analysis can be undesirable.
+Even taking minutes to run a battery draining analysis can be undesirable.
 If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
@@
   where
-    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
-    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.
+    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
+    hasPermission(App, "ACCESS_COARSE_LOCATION") = false.
 
@@
     taintDroidCheckLeak(App, "Location", "Internet")
-      = False.
+      = false.
 \end{lstlisting}
@@
 For instance Bob might want to check that Alice is at her office.
-Bob might use a location tracking app to do this.
-Provided the app only talks to Bob, and it uses SSL correctly (something Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.
+Bob might track Alice using a location tracking app.
+Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.
 
@@
   where
-    mallodroidSSLCheck(App) = False,
+    mallodroidSSLCheck(App) = false,
     connections(App) = "[https://bob.com]".
@@
 Is it the \ac{av} failing to spot the malware?
-Has someone else configured something incorrectly?
+Has someone else configured the software incorrectly?
 How should they fix the problem?
@@
 
-%
-%We also add the following constraint functions which check simple properties of the app:
-%In every case the check can be done by examining the app itself.
-%\begin{description}
-%  \item[\texttt{hasPermission(App, String)} : Bool]
-%    Checks the permissions requested in the app's manifest to see if the string is included in them.
-%
-%  \item[\texttt{hasPackage(App, String)} : Bool]
-%    Checks the app to see if a package called \emph{String} is included within it.
-%    Examples might include aggressive advertisement frameworks a user wishes to ban.
-%
-%  \item[\texttt{hasNativeCode(App)} : Bool]
-%    Checks to see whether the app contains native code.
-%    Native (i.e. non-Dalvik) code is potentially more dangerous than the sandboxed Dalvik versions.
-%    It is typically used for performance sensitive code (i.e. games).
-%
-%  \item[\texttt{isSignedBy(App, Speaker)} : Bool]
-%    Checks the \texttt{CERT.RSA} used to sign the App and compares it with the certificate used to identify a speaker.
-%
-%  \item[\texttt{hasVersion(App)} : Int]
-%    Gets the version number of an app.
-%
-%  \item[\texttt{released(App)} : Date]
-%    Checks when the app was last updated and released.
-%\end{description}
-%
-%We also add some constraints that do more complex checks:
-%\begin{description}
-%  \item[\texttt{googlePlayRating(App)} : Int]
-%    Used to get the review score (out of 5) from the Google Play store.
-%    This could be extended to other reviewers.
-%
-%  \item[tool\texttt{Check(App, Config)} : Bool]
-%    Uses the \emph{tool} to run a check on an app.
-%    This allows static analysis tools to run inferences on apps on demand.
-%    Static analysis can be expensive for devices to perform.
-%    An alternative would be to query a database of results~\citep{Noel:2014vt}.
-%
-%  \item[tool\texttt{CheckEvidence(App, Config, Evidence)} : Bool]
-%    Digital evidence and proof~carrying~code~\citep{Necula:1996tr,Stark:2009uc} allow you to check an analysis result without repeating it.
-%    In a mobile phone this can be preferable to doing the inference as draining batteries to perform checks is undesirable.
-%    Tools, such as \emph{evicheck}~\citep{Noel:2014vt}, can produce digital evidence for Android apps.
-%    This constraint function allows us to incorporate these tools into AppPAL policies.
-%\end{description}
-
 \subsection{Evaluation}
@@
 We would like to keep the context as small as possible.
-To do this we allow some assertions to be imported at evaluation-time.
+For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
 
@@
 
-In SecPAL there is no notion of type.
+In SecPAL there is no type.
 All entities are either constants (effectively strings) or variables.
-There is no differentiation made between a \emph{voiced} constant (one who speaks),
-and a \emph{subject} (one that is the subject of a verb-phrase, or predicate argument).
+There is no differentiation made between a \emph{voiced} constant (who says assertions),
+and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
 In AppPAL the subjects are usually apps or policies.
@@
 
-In this section we give examples of policies and queries and show how they can be implemented in AppPAL.
+We give examples of policies and queries and show how they can be implemented in AppPAL.
 
 \begin{description}
+  \item[Policies for home and work]
+    In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
+    Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
+    To implement this we add rules that use constraints to decide whether an app should be run.
+    Alice already has the Emma's work policy as the default.
+    Now she just needs to add an exception to the rule.
+    \begin{lstlisting}
+"alice" says App isRunnable
+  if "home-policy" isMetBy(App)
+  where TimeOfDay() > 17.00.
+
+"alice" says App isRunnable
+  if "home-policy" isMetBy(App)
+  where At("work") = false.
+    \end{lstlisting}
+
   \item[App white-listing by an IT department]
-    Inside Emma's company employees in the IT department may white-list certain apps as installable.
-    To do this Emma delegates to them to decide if something is installable.
+    Inside Emma's company employees in the IT department may white-list certain apps as runnable.
+    Emma delegates to them to decide if an app is runnable.
     \begin{lstlisting}
-"emma" says "it-department" can-say 0 App isInstallable.
+"emma" says "it-department" can-say 0 App isRunnable.
     \end{lstlisting}
     Charlie and Diveena work in the IT department.
-    When they have checked an app they say it is installable.
+    When they have checked an app they say it is runnable.
     \begin{lstlisting}
@@
 
-"charlie" says "com.facebook.katana" isInstallable.
-"diveena" says "org.thoughtcrime.securesms" isInstallable.
+"charlie" says "com.facebook.katana" isRunnable.
+"diveena" says "org.thoughtcrime.securesms" isRunnable.
     \end{lstlisting}
@@
     \begin{lstlisting}
-"alice" says App isInstallable
+"alice" says App isRunnable
    if "permissionsPolicy" isMetBy(App).
@@
       category(App, "Photography"),
-      hasPermission(App, "ACCESS_COARSE_LOCATION") = False,
-      hasPermission(App, "CAMERA") = True.
+      hasPermission(App, "ACCESS_COARSE_LOCATION") = false,
+      hasPermission(App, "CAMERA") = true.
     \end{lstlisting}
@@
     Using this tool someone generates a proof certificate that this property is met.
-    This could be the developer of the app, a third party certification service, or someone else;
-      when we check the proof we will find out if it is valid or not.
+    This might be the developer of the app, a third party certification service, or someone else;
+      when we check the proof we will discover if it is valid or not.
     \begin{lstlisting}
@@
   where
-    evicheckCheckEvidence(Evidence, App) = True.
+    evicheckCheckEvidence(Evidence, App) = true.
     \end{lstlisting}
     In the example here Alice allows anyone to present her with evidence for her policy.
-    A \ac{skb} offers an some evidence, and Alice accepts it as if it checks out.
+    A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
 \end{description}
 
-\section{Related work \comment{1--2 pages}}
+\section{Related work}
 
+Authorization logics have been successfuly used to enforce policies in several other domains.
+The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general, if undecidable.
+Logics that followed like KeyNote~\citep{Blaze:1999fa} and SPKI/SDSI~\citep{Ellison:1999ui} looked at public key infrastructure.
+The RT-languages~\citep{Li:2002if,Li:2003ix,Li:2003to} were designed for credential management.
+Cassandra~\citep{Becker:2004fi} was used to model trust relationships in the british national health service.
+
+SELinux is used to describe policies for Linux processes, and for access control (on top of the Linux discretionary controls).
+It was ported to Android~\citep{Smalley:2013vl} and is used in the implementation of the permissions system.
+Google also offer an app for configuring a \emph{device policy}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}.
+This lets businesses to configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
+It cannot be used to describe policies about apps, or describe trust relationships, however.
+
+The SecPAL language was initially used for access control in distributed systems.
 We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
-Further more attempts to apply it to other areas had already succeeded~\citep{Aziz:2011vt,Humphrey:2007wc}.
-Other policy languages such as RT~\citep{Li:2003ua}, or Cassandra~\citep{Becker:2004il} might also have made a good starting point.
-However neither had the precise set of features we we're looking for (namely explicit speakers and external checking functions).
+It has already been used to describe data usage policies~\citep{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
 Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
@@
 DKAL also contains more modalities than \emph{says}.
-
-There has been a great amount of work on developing app analysis tools for Android.
-Tools such as Stowaway~\citep{Felt:2011kj} detect overprivileged apps.
-TaintDroid~\citep{Enck:2010uw} and FlowDroid~\citep{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
-Other tools like QUIRE~\citep{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
-ScanDAL~\citep{Kim:2012vt} and SCanDroid~\citep{Fuchs:2009vi} help detect privacy leaks.
-Appscopy~\citep{Feng:kPGZr_ja} searches for specific kinds of malware.
-Tools like DroidRanger~\citep{Zhou:2012tb} scan app markets for malicious apps.
-Many others exist checking and certifying other aspects of app behaviour.
 
@@
 For example an author might wish to stop malware sending premium rate text messages.
-To do this they restrict an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions (\autoref{fig:kirin}).
+To might implement this by restricting an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions (\autoref{fig:kirin}).
 Using this approach they found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
@@
 "user" says "no-write-send-sms" isMetBy(App)
-  where hasPermission(App, "SEND_SMS") = False.
+  where hasPermission(App, "SEND_SMS") = false.
 "user" says "no-write-send-sms" isMetBy(App)
-  where hasPermission(App, "WRITE_SMS") = False.
+  where hasPermission(App, "WRITE_SMS") = false.
 \end{lstlisting}
@@
   unconcerned users who'll comfortably grant anything,
-  and fence-sitters (about half of all users) who seemed not to really care the majority of the time.
+  and fence-sitters (about half of all users) who seemed not to have varying opinions.
 This suggests that users are applying policies and do care how apps behave.
-Further more \citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
+\citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
 This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
 Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
+
+There has been a great amount of work on developing app analysis tools for Android.
+Tools such as Stowaway~\citep{Felt:2011kj} detect overprivileged apps.
+TaintDroid~\citep{Enck:2010uw} and FlowDroid~\citep{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
+Other tools like QUIRE~\citep{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
+ScanDAL~\citep{Kim:2012vt} and SCanDroid~\citep{Fuchs:2009vi} help detect privacy leaks.
+Appscopy~\citep{Feng:kPGZr_ja} searches for specific kinds of malware.
+Tools like DroidRanger~\citep{Zhou:2012tb} scan app markets for malicious apps.
+Many others exist checking and certifying other aspects of app behaviour.
 
@@
 The language is implemented in Java and runs on Android using a custom evaluation algorithm.
-This allows us to enforce app installation policies on Android devices.
-We've shown how the language can be used to describe an app installation policy;
+This lets us to enforce app installation policies on Android devices.
+We have shown how the language can be used to describe an app installation policy;
   and given brief descriptions of how other policies might be described.
@@
 Further work is required to tightly integrate AppPAL into Android.
-The obvious place to put it would be as a \emph{required checker}: a program that checks all apps before installation.
-Google currently use this API to check for known malware and jailbreak apps.
-We would like AppPAL to sit there and check apps on the basis of policies.
+One way to integrate AppPAL on Android would be as a \emph{required checker}: a program that checks all apps before installation.
+Google uses this API to check for known malware and jailbreak apps.
+We would use AppPAL to check apps meet policies before installation.
 Unfortunately the API is protected and it would require the phone to be rooted to run there.
+Alternatively it could be integrated as a service to reconfigure app permissions.
+The latest version of Android\footnote{Called \emph{Android M}.} is moving to a more iOS like permissions model where permissions can be granted and revoked at any time.
 
@@
 We can imagine a similar scheme working well for app installation policies.
-Users subscribe to different policies by experts (e.g. no tracking apps, nothing with adult content, no spammy in-app-purchase apps).
+Users subscribe to different policies by experts (examples could include no tracking apps, nothing with adult content, no spammy in-app-purchase apps).
 Optionally they can customize them further.

### 2015-06-10 paper.tex
Updates and more editing

@@
 \title{AppPAL for Android}
-\subtitle{Capturing and Checking App Installation Policies}
+\subtitle{Capturing and Checking Mobile App Policies}
 % \numberofauthors{2}
@@
   %   \item \comment{Sentence 1: State the problem}
-  Users must judge apps based only on the information presented to them by the store.
+  Users must judge apps by the information shown to them by the store.
   %   \item \comment{Sentence 2: Why is it an interesting problem}
-  Employers rely on employees using company devices appropriately to enforce security policies at work.
-  Some users spend time to pick apps with care others do not.
-  This may cause frustration when they realise what data the apps have access to.
+  Employers rely on employees enforcing company policies correctly on devices at their workplace.
+  Some users take time to pick apps with care whilst others do not.
+  Users feel frustrated when they realise what data the apps have access to.
+  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
   %   \item \comment{Sentence 3: What does my solution achieve}
-  We present AppPAL: a policy language that can enforce app installation policies.
-  Policies use statements from third parties and delegation relationships.
+  We present AppPAL: a policy language to enforce policies about apps.
+  AppPAL Policies use statements from third parties and delegation relationships to
   %   \item \comment{Sentence 4: What follows from my solution}
-  This gives us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
+  give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
   % \end{itemize}
@@
 %\comment{The problem is\ldots}
-Finding good apps can be tricky.
-Users need to work out which apps are well written.
-They also need to find the apps which suit how they want to use their phone.
-This can be difficult.
-It isn't obvious how apps use the data each has access to.
+Finding the right apps can be tricky:
+  users need to work out which apps are well written, which are not going to abuse their data, which ones will work in the way they want
+  and to find the apps which suit how they want to use their device.
+This can be difficult as it isn't obvious how apps use the data each has access to.
 
-App stores give some information about their apps; such as permissions and review scores.
-Android apps show a confusing list of permissions (\autoref{fig:facebook}).
+App stores give some information about their apps; such as permissions, descriptions of the app and screenshots as well as review scores.
+Android apps currently show a confusing list of permissions when they're first installed.
+Soon new Android apps will display permissions requests when the app first tries to access sensitive data such as contacts or location information.
 Users do not all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
-Ultimately the decision which apps to use must be made by a human.
+Ultimately the decision which apps to use and which permissions to grant must be made by the user.
 
@@
 A large amount of \ac{pus} is being propagated for Android devices~\citep{Truong:2014bi,Svajcer:2013tp}.
-Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
+Employees are increasingly using their own phones for work (bring your own device or BYOD).
 An employer may wish to restrict which apps their employees can use.
-The \ac{ciso} or IT department may set a policy to prevent information leaks.
+The IT department may set a policy to prevent information leaks.
 Some users worry apps will misuse their personal data;
   such a user avoids apps which can access their location, or address book.
-They apply their own personal security policy when downloading apps.
+They may apply their own personal security policy when downloading and running apps.
 
-These policies are enforced by hand, if at all.
-This is error-prone and mistakes can be made.
-We believe this can be bettered.
-An alternative would be to write the policy down and make the computer do it.
-To implement this we can use a logic of authorization.
-The policy may be written in the logic and enforced automatically.
+These policies can only be enforced by the users continuously making decisions guided by these policies when prompted about apps.
+This is error-prone.
+Mistakes can be made.
+We believe this can be improved.
+An alternative would be to write the policy down and make the computer enforce it.
+To implement this we use a logic of authorization.
+The policy is written in the logic and enforced by checking the policy is satisfied.
 
@@
 %\comment{My contribution is\ldots}
-We present AppPAL, an authorization logic for reasoning about Android apps.
-The language is an instantiation of SecPAL~\citep{Becker:2006vh} tailored to making statements about apps.
+We present AppPAL, an authorization logic for reasoning about apps.
+The language is an instantiation of SecPAL~\citep{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
@@
 
-Using AppPAL we can make statements about apps:
+Using AppPAL we can say an app is installable:
 \begin{lstlisting}
-"Alice" says "com.rovio.angrybirds" isRunnable.
+"Alice" says "com.rovio.angrybirds" isInstallable.
 \end{lstlisting}
@@
 \end{lstlisting}
+Constraints allow up to express facts that are true at some times but not others:
+\begin{lstlisting}
+"Emma" says "com.facebook.katana" isRunnable
+  where timeOfDay() > 1700.
+\end{lstlisting}
 
-Specifically we:
+Specifically we have:
 \begin{itemize}
   \item
-    Described a scenario where a user has a policy they want to enforce for their employees~(\autoref{sec:problem}.)
+    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem}.)
     We show how the trust relationships in a system can be difficult to see;
@@
   \item
-    Shown how the user's policy could be implemented using AppPAL~(\autoref{sec:idea});
-      and describe what AppPAL would check for an app to meet this policy.
+    Shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
+    We have described what AppPAL would check for an app to meet this policy and how these statements might be collected.
 
-  \item Implemented the AppPAL language.
-    Our implementation runs on Android (\autoref{sec:details}.)
-    We show how we have tailored the language to Android (\autoref{ssec:language}),
+  \item Implemented the AppPAL language on Android and the JVM.
+    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}),
       how we evaluate it (\autoref{ssec:eval}),
-      and some idiomatic policies (\autoref{ssec:idioms}.)
+      and some idiomatic policies (\autoref{ssec:idioms}).
 
@@
     Authorization logic can help make trust relationships explicit.
-    In \autoref{ssec:language} we show a recent example of a broken trust relationship,
-    and how our authorization logic makes the problem explicit.
+    We show a recent example of a broken trust relationship (\autoref{ssec:language})
+    and how our authorization logic makes explicit where the breach of trust is.
 
@@
 
-Suppose \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
+An employee \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
 Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
@@
 
-To ensure this policy is fully met Alice could promise to follow it.
-Perhaps she might even sign a document promising to do so.
-This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
+To ensure this policy is met Alice promises to follow it.
+She might even sign a document promising never to break the rules within the policy.
+This is error-prone though---what if she makes a mistake or misses an app that breaks her policy?
 Emma's policy could be implemented using existing tools.
 The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
-  could configure Alice's device to disallow apps from outside the Google Play Store.
+could configure Alice's device to disallow apps from outside the Google Play Store (and soon in the newest version of Android let Emma set the permissions of each app on an app by app basis).
 Various tools such as AppGuard~\citep{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\citep{Jeon:2012ki} or AppFence~\citep{Hornyack:2011wq} can control the permissions or data an app can get.
@@
 Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.
+
+Whilst we could implement Emma's policy using existing tools, it is a clumsy solution.
+They are not flexible: If Emma changes her policy or Alice changes jobs she needs to recheck and then to alter and remove the software on her phone accordingly.
+It isn't clear what an app must do to be run, or what checks have been done if it already running on the phone.
+The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.
+
+What happens when Alice goes home?
+Emma shouldn't be able to overly control what Alice does in her private life.
+Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
+Some mobile OSs allow app permissions to be enabled and disable at run time.
+Can we enforce different policies at different times or locations?
+
+Our research looks at the problem of picking software.
+Given there are some apps you want to install and run and others you do not want to, at least some of the time:
+  how can you express your preferences in such a way that they can be enforced automatically?
+How can we translate policy documents from natural language into a machine checkable form?
+Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?
 
@@
 
-Whilst we \emph{could} implement Emma's policy using existing tools, it isn't a great solution.
-It isn't flexible, Alice needs to alter and remove the software on her phone if Emma changes her policy or she changes jobs.
-It isn't clear what an app must do to be run, or what checks have been done if it already has been put on the phone.
-The link between Alice, the user; Emma, the policy setter; and the tools Emma trusts to implement her policy isn't immediately apparent.
-
-What happens when Alice goes home?
-Emma shouldn't be able to overly control what Alice does in her private life.
-Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
-The some mobile OSs allow app permissions to be enabled and disable at run time.
-Can we enforce different policies at different times?
-
-Our research looks at the problem of picking software.
-Given there are some apps you want and others you do not:
-  how can you express your preferences in such a way that they can be enforced automatically?
-How can we translate policy documents from natural language into a machine checkable form?
-Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?
+We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
+People have policies which are enforced by AppPAL on their devices.
+The device can make use of vetting services which run tools to infer complex properties about apps.
+Users can buy from stores which ensure the only apps they see are the apps which meet their policies.
 
@@
 
-Returning to Alice and Emma's dilemma in \autoref{sec:problem}, we propose a solution.
+In \autoref{sec:problem} Alice and Emma's had policies they wanted to enforce but no means to do so.
 Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
@@
 
-AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app running policies.
+AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for describing app policies.
 SecPAL is a logic of authorization for access control decisions in distributed systems.
@@
 \begin{figure}
+  \begin{minipage}[t]{0.5\textwidth}
     \begin{lstlisting}[numbers=left,
@@
   if "no-tracking-policy" isMetBy(App),
-     "reputable-source-policy" isMetBy(App),
+     "reputable-policy" isMetBy(App),
      "anti-virus-policy" isMetBy(App).
 
-"emma" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
+"emma" says
+  "reputable-policy" isMetBy(App) @\label{lst:corporate_3}@
       if App isReputable.
 
-"emma" says "google-play" can-say 0 App isReputable. @\label{lst:corporate_4}@
-
+"emma" says "google-play" can-say 0
+  App isReputable. @\label{lst:corporate_4}@
+    \end{lstlisting}
+  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
+    \begin{lstlisting}[numbers=left,
+                       escapeinside={@}{@},
+                       firstnumber=14]
 "emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
@@
 
-"emma" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
+"emma" says
+  "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
     if App isAnApp
     where
-    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
-    hasPermission(App, "ACCESS_COURSE_LOCATION") = false.
+      hasPermission(App, "LOCATION")=false.
 \end{lstlisting}
+\end{minipage}
 \caption{AppPAL policy implementing Emma's security requirements}
@@
 If Emma can be convinced all these policies are met then he will say the \code{App isRunnable}.
-In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-source-policy} if the \code{App isReputable};
+In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
   with \code{"google-play"} specified as the decider of what is buyable or not.
@@
     A simple typing statement that can be generated for all apps as they are encountered.
+    This helps keep the number of  assertions in the policy low aiding readability.
   \item\code{"google-play" says "com.facebook.katana" isReputable.}
@@
 To get the third statement it must run the \ac{av} program on her app and check the result.
-The results from the \ac{av} program may change with time as it's signatures are updated.
-Consequently the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
+The results from the \ac{av} program may change with time as it's signatures are updated;
+  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
 For the forth statement the checker needs to check the permissions of the app.
-It does this by looking in the \texttt{MANIFEST.xml} inside the app itself.
+It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself.
 
@@
 Alternatively we could imagine Emma wanting a personalised app store where all apps sold meet her policy.
-With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which meet the policy.
+With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which will meet the user's policy.
 This gives us a \emph{filtered store}.
@@
   \item[\texttt{App isAnApp}, \texttt{Policy isAPolicy}\ldots]
-    Simple typing relations.
-    The app \text{App} is an App, and so on.
+    Expresses a simple typing relations.
+    AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
+    These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
+    This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
 
@@
   where
-    taintDroidCheckLeak(App, "Location", "Internet")
-      = false.
+    taintDroidCheckLeak(App, "Location", "Internet") = false.
 \end{lstlisting}
@@
 Rather we implement the rules directly in Java.
-Brief pseudo-code is given in \autoref{fig:pseudocode}
+Pseudo-code is given in \autoref{fig:pseudocode}.
 Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
@@
   \models \mathtt{c}\theta &
-  \text{\sl vars}(\text{\tt fact$\theta$})}
+  \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
 \]
@@
   we can fetch the assertion that the subject is an app based on the app in question.
-Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is purchasable;
+Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
@@
 In SecPAL there is no type.
-All entities are either constants (effectively strings) or variables.
+All entities are either constants (implemented as strings) or variables.
 There is no differentiation made between a \emph{voiced} constant (who says assertions),
@@
 In AppPAL the subjects are usually apps or policies.
-Apps do not utter facts.
-Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
+Apps do not utter facts;
+  so there is no need to check them as potential speakers of \emph{can-say} statements.
 Similarly if a constant never appears as a subject of a statement, as most speakers do not,
@@
 We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
+This means we need to search less constants when attempting to replace a variable, speeding evaluation.
 
@@
 def checkConditions(ac, rt, a, d)
-  getVarSubstitutions(a, ac.constants).each do |s|
+  getVarSubstitutions(a,ac.constants).each do |s|
     sa = s.substitute a
@@
 
+When writing policies some schemes often come up.
 We give examples of policies and queries and show how they can be implemented in AppPAL.
@@
 "alice" says App isRunnable
-   if "permissionsPolicy" isMetBy(App).
-"alice" says "permissionsPolicy" isMetBy(App)
+   if "permissions-policy" isMetBy(App).
+"alice" says "permissions-policy" isMetBy(App)
    if App isAnApp
@@
       category(App, "Photography"),
-      hasPermission(App, "ACCESS_COARSE_LOCATION") = false,
+      hasPermission(App, "LOCATION") = false,
       hasPermission(App, "CAMERA") = true.
@@
 It was ported to Android~\citep{Smalley:2013vl} and is used in the implementation of the permissions system.
-Google also offer an app for configuring a \emph{device policy}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}.
+Google also offer the \emph{Device Policy for Android} app.
 This lets businesses to configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
@@
 \citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
-DKAL also contains more modalities than \emph{says}.
+DKAL also contains more modalities than \emph{says}, which lets policies describe actions principals carry out rather than just their oppinions.
+For example in AppPAL a user might \emph{say} an app is installable if they would install it (\code{"user" says App isInstallable})
+In DKAL they can describe the conditions that would force them to install it (\code{"user" installs App}).
+The distinction is that in AppPAL whilst the user thinks the app could be installed we do not know for sure whether the user has installed it.
+With the DKAL we can guarantee that the action was completed.
 
@@
 Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
+We have created an Android app that can check approximations of \citeauthor{Sadeh:2014vq}'s policies against apps on a users device.
+This gives us a demonstration of the power of AppPAL using known policies\footnote{\url{https://github.com/bogwonch/apppal-checker}}.
 
@@
 Unfortunately the API is protected and it would require the phone to be rooted to run there.
-Alternatively it could be integrated as a service to reconfigure app permissions.
+Alternatively AppPAL could be integrated as a service to reconfigure app permissions.
 The latest version of Android\footnote{Called \emph{Android M}.} is moving to a more iOS like permissions model where permissions can be granted and revoked at any time.
+These will be manually configurable by the user through the settings app.
+We can imagine AppPAL working to reconfigure these settings (and set their initial grant or deny states) based on a user's policy, as well as the time of day or the user's location.
+A policy could deny pop-up notices while a user is driving for example.
 
@@
 Users policies can be enforced more reliably, and with less interaction;
-thus making apps more pleasant and safer for everyone.
+making apps more pleasant for everyone and helping to reduce user fatigue.
 

### 2015-07-31 paper.tex
Updates and switched to LNCS style for ESSOS

@@
 % vim: set nowrap tw=0:
-\documentclass[]{scrartcl}
+\documentclass[]{llncs}
 
@@
 \author{Joseph Hallett \and David Aspinall }
+\institute{University of Edinburgh}
 
@@
 Soon new Android apps will display permissions requests when the app first tries to access sensitive data such as contacts or location information.
-Users do not all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
+Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
 Ultimately the decision which apps to use and which permissions to grant must be made by the user.
@@
 Not all apps are suitable.
-A large amount of \ac{pus} is being propagated for Android devices~\citep{Truong:2014bi,Svajcer:2013tp}.
+A large amount of \ac{pus} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
 Employees are increasingly using their own phones for work (bring your own device or BYOD).
@@
 We present AppPAL, an authorization logic for reasoning about apps.
-The language is an instantiation of SecPAL~\citep{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
+The language is an instantiation of SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
@@
 We can express trust relationships amongst these parties; use constraints to do additional checks.
-This lets us enforce deeper and more complex policies than existing tools such as Kirin~\citep{Enck:2009ko}.
+This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.
 
@@
 could configure Alice's device to disallow apps from outside the Google Play Store (and soon in the newest version of Android let Emma set the permissions of each app on an app by app basis).
-Various tools such as AppGuard~\citep{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\citep{Jeon:2012ki} or AppFence~\citep{Hornyack:2011wq} can control the permissions or data an app can get.
+Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
 These could be used to used to ensure no location data is ever obtained.
-Alternately other tools like Kirin~\citep{Enck:2009ko}, Flowdroid~\citep{Fritz:2013vi} or DroidSafe~\citep{Gordon:2015et} could check that the locations are ever leaked to the web.
+Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are ever leaked to the web.
 Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.
@@
 
-AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for describing app policies.
+AppPAL is an instantiation of SecPAL~\cite{Becker:2006vh} for describing app policies.
 SecPAL is a logic of authorization for access control decisions in distributed systems.
 It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
-SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\citep{Becker:2009ula} and data-sharing~\citep{Aziz:2011vt}.
+SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
 We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.
@@
 
-In the \citeauthor{Becker:2006vh}'s paper~\citep{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
+In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
 With AppPAL we make explicit our predicates and how they relate to Android.
@@
 Some apps may want to access this data, but may not leak it.
-We could use a taint analysis tool to detect this (e.g.~Taintdroid~\citep{Fritz:2013vi}).
+We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
 Our policy now becomes:
@@
 Bob might track Alice using a location tracking app.
-Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.
+Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) he is happy to relax the policy.
 
@@
 To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
-We do not use the DatalogC~\citep{Li:2003ix}~based translation and evaluation algorithm suggested by \citeauthor{Becker:2006vh}.
+We do not use the DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm suggested by Becker~\etal.
 Rather we implement the rules directly in Java.
 Pseudo-code is given in \autoref{fig:pseudocode}.
-Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
+Like Becker~\etal we make use of an assertion context to store known statements.
 We also store intermediate results to avoid re-computation.
@@
     Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
-    Alternately tools like Stowaway~\citep{Felt:2011kj} could be integrated to check apps.
+    Alternately tools like Stowaway~\cite{Felt:2011kj} could be integrated to check apps.
 
@@
     Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
-    For example that no audio can be recorded without the users consent~\citep{Seghir:2014uq}.
+    For example that no audio can be recorded without the users consent~\cite{Seghir:2014uq}.
     Using this tool someone generates a proof certificate that this property is met.
@@
 The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general, if undecidable.
-Logics that followed like KeyNote~\citep{Blaze:1999fa} and SPKI/SDSI~\citep{Ellison:1999ui} looked at public key infrastructure.
-The RT-languages~\citep{Li:2002if,Li:2003ix,Li:2003to} were designed for credential management.
-Cassandra~\citep{Becker:2004fi} was used to model trust relationships in the british national health service.
+Logics that followed like KeyNote~\cite{Blaze:1999fa} and SPKI/SDSI~\cite{Ellison:1999ui} looked at public key infrastructure.
+The RT-languages~\cite{Li:2002if,Li:2003ix,Li:2003to} were designed for credential management.
+Cassandra~\cite{Becker:2004fi} was used to model trust relationships in the british national health service.
 
 SELinux is used to describe policies for Linux processes, and for access control (on top of the Linux discretionary controls).
-It was ported to Android~\citep{Smalley:2013vl} and is used in the implementation of the permissions system.
+It was ported to Android~\cite{Smalley:2013vl} and is used in the implementation of the permissions system.
 Google also offer the \emph{Device Policy for Android} app.
@@
 The SecPAL language was initially used for access control in distributed systems.
-We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
-It has already been used to describe data usage policies~\citep{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
-Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
-\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
+We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\cite{Hallett:2014un}.
+It has already been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
+Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+Gruevich and Neeman showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
 DKAL also contains more modalities than \emph{says}, which lets policies describe actions principals carry out rather than just their oppinions.
@@
 
-One tool, Kirin~\citep{Enck:2009ko}, also created a policy language and tool for enforcing app installation policies.
+One tool, Kirin~\cite{Enck:2009ko}, also created a policy language and tool for enforcing app installation policies.
 Kirin's policies were concerned with preventing malware.
@@
 
-Work by \citeauthor{Sadeh:2014vq} looked at modelling users' willingness to grant different kinds of permissions~\citep{Sadeh:2014vq}.
+Work by Lin~\etal looked at modelling users' willingness to grant different kinds of permissions~\cite{Sadeh:2014vq}.
 They found four clusters of different users each with different kinds of policies:
@@
 This suggests that users are applying policies and do care how apps behave.
-\citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
+Lin~\etal speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
 This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
 Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
-We have created an Android app that can check approximations of \citeauthor{Sadeh:2014vq}'s policies against apps on a users device.
+We have created an Android app that can check approximations of Lin~\etal's policies against apps on a users device.
 This gives us a demonstration of the power of AppPAL using known policies\footnote{\url{https://github.com/bogwonch/apppal-checker}}.
@@
 There has been a great amount of work on developing app analysis tools for Android.
-Tools such as Stowaway~\citep{Felt:2011kj} detect overprivileged apps.
-TaintDroid~\citep{Enck:2010uw} and FlowDroid~\citep{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
-Other tools like QUIRE~\citep{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
-ScanDAL~\citep{Kim:2012vt} and SCanDroid~\citep{Fuchs:2009vi} help detect privacy leaks.
-Appscopy~\citep{Feng:kPGZr_ja} searches for specific kinds of malware.
-Tools like DroidRanger~\citep{Zhou:2012tb} scan app markets for malicious apps.
+Tools such as Stowaway~\cite{Felt:2011kj} detect overprivileged apps.
+TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
+Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
+ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
+Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
+Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
 Many others exist checking and certifying other aspects of app behaviour.
@@
 We should also attempt to learn policies from existing users behavior.
-Given app usage data, from a project like Carat~\citep{Oliner:2013ht}, we could identify security conscious users.
+Given app usage data, from a project like Carat~\cite{Oliner:2013ht}, we could identify security conscious users.
 If we can infer these users policies we may be able to describe new policies that the less technical users may want.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{splncs03}
 \bibliography{paper}

### 2015-08-05 paper.tex
updates and editing

@@
 
-% \pdfpagewidth=8.5truein{}
-% \pdfpageheight=11truein{}
-
 \usepackage{graphicx}
-% \renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
-% \renewcommand{\floatpagefraction}{0.99}
-% \pagestyle{plain} % page numbers
-
 \usepackage{defaultstyle}
@@
 \begin{abstract}
-  % \begin{itemize}
-  %   \item \comment{Sentence 1: State the problem}
   Users must judge apps by the information shown to them by the store.
-  %   \item \comment{Sentence 2: Why is it an interesting problem}
   Employers rely on employees enforcing company policies correctly on devices at their workplace.
@@
   They want greater control over what data they give away but they do not want to spend time reading permissions lists.
-  %   \item \comment{Sentence 3: What does my solution achieve}
   We present AppPAL: a policy language to enforce policies about apps.
   AppPAL Policies use statements from third parties and delegation relationships to
-  %   \item \comment{Sentence 4: What follows from my solution}
   give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
@@
 
-%\comment{The problem is\ldots}
-Finding the right apps can be tricky:
-  users need to work out which apps are well written, which are not going to abuse their data, which ones will work in the way they want
+Finding the right apps can be tricky.
+Users need to work out which apps are well written, which are not going to abuse their data, which will work in the way they want
   and to find the apps which suit how they want to use their device.
@@
 
-App stores give some information about their apps; such as permissions, descriptions of the app and screenshots as well as review scores.
-Android apps currently show a confusing list of permissions when they're first installed.
-Soon new Android apps will display permissions requests when the app first tries to access sensitive data such as contacts or location information.
+App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
+Android apps show a list of permissions when they're first installed.
+Soon new Android apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
 Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
@@
 Not all apps are suitable.
-A large amount of \ac{pus} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
-Employees are increasingly using their own phones for work (bring your own device or BYOD).
+A large amount of \ac{pup} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
+Employees are increasingly using their own phones for work (\ac{byod}).
 An employer may wish to restrict which apps their employees can use.
@@
 
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=0.3\linewidth]{figures/facebook.png}
-%   \caption{Security information presented when a user downloads the Facebook app.}
-% \label{fig:facebook}
-% \end{figure}
-
-%\comment{My contribution is\ldots}
 We present AppPAL, an authorization logic for reasoning about apps.
@@
 
-Using AppPAL we can say an app is installable:
+Using AppPAL we write policies for work and home, and decide which policy to use using a user's location, or the time of day:
 \begin{lstlisting}
-"Alice" says "com.rovio.angrybirds" isInstallable.
+"alice" says App isRunnable
+  if "home-policy" isMetBy(App)
+  where At("work") = false.
+
+"alice" says App isRunnable
+  if "work-policy" isMetBy(App)
+  where BeforeHourOfDay("17") = true.
 \end{lstlisting}
-We can delegate decisions to others:
+We can delegate policy specification to third parties or roles, and assign principals to those rolls:
 \begin{lstlisting}
-"Alice" says "Bob" can-say 0 App isCool.
+"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
+"alice" says "bob" can-act-as "it-department".
 \end{lstlisting}
-We can run analysis tools, and solve constraints:
+We can write policies specifying which permissions an app must or must not have by its app store categorization:
     \begin{lstlisting}
-"Claire" says App isMalware
+"alice" says App isRunnable
+   if "permissions-policy" isMetBy(App).
+"alice" says "permissions-policy" isMetBy(App)
    if App isAnApp
    where
-    virusScanner(App) = true.
-\end{lstlisting}
-Constraints allow up to express facts that are true at some times but not others:
-\begin{lstlisting}
-"Emma" says "com.facebook.katana" isRunnable
-  where timeOfDay() > 1700.
+      category(App, "Photography"),
+      hasPermission(App, "LOCATION") = false,
+      hasPermission(App, "CAMERA") = true.
     \end{lstlisting}
 
-Specifically we have:
+% Using AppPAL we can say an app is installable:
+% \begin{lstlisting}
+% "Alice" says "com.rovio.angrybirds" isInstallable.
+% \end{lstlisting}
+% We can delegate decisions to others:
+% \begin{lstlisting}
+% "Alice" says "ITDepartment" can-say 0 App isInstallable.
+% \end{lstlisting}
+% We can run analysis tools, and solve constraints:
+% \begin{lstlisting}
+% "Claire" says App isMalware
+%   if App isAnApp
+%   where
+%     virusScanner(App) = true.
+% \end{lstlisting}
+% Constraints allow up to express facts that are true at some times but not others:
+% \begin{lstlisting}
+% "Emma" says "com.facebook.katana" isInstallable
+%   where timeOfDay() > 1700.
+% \end{lstlisting}
+
+\noindent
+Specifically our contribution is:
 \begin{itemize}
   \item
-    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem}.)
-    We show how the trust relationships in a system can be difficult to see;
-      and how AppPAL makes the trust between users explicit.
-  \item
-    Shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
-    We have described what AppPAL would check for an app to meet this policy and how these statements might be collected.
+    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
+    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
 
   \item Implemented the AppPAL language on Android and the JVM.
-    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}),
-      how we evaluate it (\autoref{ssec:eval}),
-      and some idiomatic policies (\autoref{ssec:idioms}).
+    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).
 
-  \item When things go wrong we also need to attribute blame.
-    Authorization logic can help make trust relationships explicit.
-    We show a recent example of a broken trust relationship (\autoref{ssec:language})
-    and how our authorization logic makes explicit where the breach of trust is.
-
+  \item Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
 \end{itemize}
@@
 Emma's policy could be implemented using existing tools.
-The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
-could configure Alice's device to disallow apps from outside the Google Play Store (and soon in the newest version of Android let Emma set the permissions of each app on an app by app basis).
+The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store (and in the newest version of Android let Emma set the permissions of each app on an app by app basis~\cite{AndroidMPermission:2015uq}).
+AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and formalize the decision making process.
 Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
@@
 Users can buy from stores which ensure the only apps they see are the apps which meet their policies.
+Policies can be combined
 
@@
 % A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.
-
-% \begin{figure}
-%   \includegraphics{figures/bob-policy.eps}
-%   \caption{Pictorial representation of Alice's app install policy.}
-% \label{fig:emmas_policy}
-% \end{figure}
 
@@
                        escapeinside={@}{@},
-                       firstnumber=14]
+                       firstnumber=15]
 "emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
@@
 
-Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
+Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
 To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
@@
 For the forth statement the checker needs to check the permissions of the app.
-It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself.
+It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if running on device.
 
@@
 \label{sec:details}
+\label{ssec:language}
 
-AppPAL is implemented as a library for Android and Java.
+AppPAL is an instantiation of SecPAL for Android app installation policies.
+It is implemented as a library for Android and Java.
 The parser is implemented using ANTLR4.
 Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.
-
-\subsection{The Language}
-\label{ssec:language}
 
@@
 
-In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
-With AppPAL we make explicit our predicates and how they relate to Android.
-Some of these predicates require arguments.
-We add a light typing system to help ensure these predicates have the correct arguments.
+In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of
+predicates, and constraints for their SecPAL open.  With AppPAL we make
+explicit our predicates and how they relate to Android.  Some of these
+predicates require arguments.  AppPAL policies typically make use of the
+predicates and constraints in \autoref{tab:predicates}.  Additional predicates
+can be created in the policy files, but adding or modifying constraints
+requires a code change\footnote{But a small one.  The Android version of the
+\texttt{hasPermission} constraint, for example, uses the Android package
+manager to determine what permissions an app requests, but the Java version
+uses the Android platform tools.  This required a code change to one function.}
 
-Specifically we define the following predicates:
-\begin{description}
-  \item[\texttt{App isRunnable}]
-    Used to indicate that an App meets the install policy for the device.
-    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
+\begin{table}
+\begin{tabulary}{\linewidth}{l L}
+  \toprule
+  Name & Description \\
+  \midrule
+  App \texttt{isRunnable} & Says an app can be run. \\
+  App \texttt{isInstallable} & Says an app can be installed. \\
+  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
+  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
+  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
+  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
+  \bottomrule
+\end{tabulary}
+\caption{Typical AppPAL predicates and constraints}
+\label{tab:predicates}
+\end{table}
 
-  \item[\texttt{App isAnApp}, \texttt{Policy isAPolicy}\ldots]
-    Expresses a simple typing relations.
-    AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
-    These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
-    This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
-
-  \item[\texttt{Policy isMetBy(App)}]
-    Policies let you split app behaviour into sub-policies.
-    For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
-    Splitting policies allows greater control of how each one is checked.
-    We can delegate checking a policy to an expert using the \code{can-say} statement.
-    We can specify more detailed checks using the \code{can-act-as} statements.
-
-  \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
-    A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
-\end{description}
+% \begin{description}
+%   \item[\texttt{App isRunnable}]
+%     Used to indicate that an App meets the install policy for the device.
+%     Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
+%
+%   \item[\texttt{App isAnApp}]
+%     Expresses a simple typing relations.
+%     AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
+%     These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
+%     This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
+%
+%   \item[\texttt{Policy isMetBy(App)}]
+%     Policies let you split app behaviour into sub-policies.
+%     For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
+%     Splitting policies allows greater control of how each one is checked.
+%     We can delegate checking a policy to an expert using the \code{can-say} statement.
+%     We can specify more detailed checks using the \code{can-act-as} statements.
+%
+%   \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
+%     A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
+% \end{description}
 
@@
 
-Consider the \emph{no-tracking-policy} from earlier.
+Earlier we described a \emph{no-tracking-policy} to prevent a user's location being tracked.
 In Bob's policy we checked this by checking the permissions of the app.
@@
 
-AppPAL also helps attribute blame when things go wrong.
-By modelling the trust relationships in systems we can work out precisely where mistakes were made.
-A recent example demonstrating this is \textsc{CVE-2015-2077}.
-Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
-Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
-Unfortunately they had configured it to ignore Superfish.
-The malware was an ad framework designed to show users products similar to those they viewed on the web.
-Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.
+% AppPAL also helps attribute blame when things go wrong.
+% By modelling the trust relationships in systems we can work out precisely where mistakes were made.
+% A recent example demonstrating this is \textsc{CVE-2015-2077}.
+% Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
+% Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
+% Unfortunately they had configured it to ignore Superfish.
+% The malware was an ad framework designed to show users products similar to those they viewed on the web.
+% Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.
+%
+% From a user's perspective there was a delegation of trust to Lenovo to detect malware.
+% Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
+% A user might wonder where the breach of trust occurred and who is to blame?
+% Is it the \ac{av} failing to spot the malware?
+% Has someone else configured the software incorrectly?
+% How should they fix the problem?
+%
+% If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
+% \begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
+% "user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
+% "lenovo" says "mcafee" can-say inf File isSafe.
+% "lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
+% \end{lstlisting}
+% The fault lies with \autoref{lst:lenovo_unsafe}.
+% Lenovo has caused the problem (they \emph{said} it was safe).
+% The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
 
-From a user's perspective there was a delegation of trust to Lenovo to detect malware.
-Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
-A user might wonder where the breach of trust occurred and who is to blame?
-Is it the \ac{av} failing to spot the malware?
-Has someone else configured the software incorrectly?
-How should they fix the problem?
-
-If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
-\begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
-"user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
-"lenovo" says "mcafee" can-say inf File isSafe.
-"lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
-\end{lstlisting}
-The fault lies with \autoref{lst:lenovo_unsafe}.
-Lenovo has caused the problem (they \emph{said} it was safe).
-The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
-
-\subsection{Evaluation}
+\subsection{Policy checking}
 \label{ssec:eval}
 
-To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
-We do not use the DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm suggested by Becker~\etal.
-Rather we implement the rules directly in Java.
+Since AppPAL is an instantiation of SecPAL the policy checking rules are the same.
+We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm, as no DatalogC library exists for Android.
+We have implemented the rules directly in Java.
 Pseudo-code is given in \autoref{fig:pseudocode}.
@@
 
-\begin{figure}
-\[\footnotesize
-% AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
-% ----------------------------------------------------------- (can-act-as)
-%                 AC, D ⊧ A says B verbphrase
-\infer[\text{\tiny can-act-as}]{
-  AC, D \models \text{\tt A says B verbphrase}}{%
-  AC, D \models \text{\tt A says B can-act-as C} & %
-  AC, D \models \text{\tt A says C verbphrase}}
-\]
-\[\footnotesize
-% AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
-% ----------------------------------------------------- (can-say)
-%                   AC, ∞ ⊧ A says fact
-\infer[\text{\tiny can-say}]{
-  AC, \infty \models \text{\tt A says fact}}{%
-  AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
-  AC, D \models \text{\tt B says fact}}
-\]
-\[\footnotesize
-% (A says fact if fact₁, …, factₖ, c) ∈ AC
-% AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
-% ----------------------------------------------------------- (cond)
-%                    AC, D ⊧ A says factθ
-\infer[\text{\tiny cond}]{
-  AC, D \models \text{\tt A says fact$\theta$}}{
-  \begin{array}{c}
-    (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
-    \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
-  \end{array} &
-  \models \mathtt{c}\theta &
-  \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
-\]
-\caption{AppPAL's evaluation rules.}
-\label{fig:rules}
-\end{figure}
+% \begin{figure}
+% \[\footnotesize
+% % AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
+% % ----------------------------------------------------------- (can-act-as)
+% %                 AC, D ⊧ A says B verbphrase
+% \infer[\text{\tiny can-act-as}]{
+%   AC, D \models \text{\tt A says B verbphrase}}{%
+%   AC, D \models \text{\tt A says B can-act-as C} & %
+%   AC, D \models \text{\tt A says C verbphrase}}
+% \]
+% \[\footnotesize
+% % AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
+% % ----------------------------------------------------- (can-say)
+% %                   AC, ∞ ⊧ A says fact
+% \infer[\text{\tiny can-say}]{
+%   AC, \infty \models \text{\tt A says fact}}{%
+%   AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
+%   AC, D \models \text{\tt B says fact}}
+% \]
+% \[\footnotesize
+% % (A says fact if fact₁, …, factₖ, c) ∈ AC
+% % AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
+% % ----------------------------------------------------------- (cond)
+% %                    AC, D ⊧ A says factθ
+% \infer[\text{\tiny cond}]{
+%   AC, D \models \text{\tt A says fact$\theta$}}{
+%   \begin{array}{c}
+%     (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
+%     \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
+%   \end{array} &
+%   \models \mathtt{c}\theta &
+%   \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
+% \]
+% \caption{AppPAL's evaluation rules.}
+% \label{fig:rules}
+% \end{figure}
 
@@
 
-In SecPAL there is no type.
-All entities are either constants (implemented as strings) or variables.
-There is no differentiation made between a \emph{voiced} constant (who says assertions),
-and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
-In AppPAL the subjects are usually apps or policies.
-Apps do not utter facts;
-  so there is no need to check them as potential speakers of \emph{can-say} statements.
-Similarly if a constant never appears as a subject of a statement, as most speakers do not,
-  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
-We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
-This means we need to search less constants when attempting to replace a variable, speeding evaluation.
+% In SecPAL there is no type.
+% All entities are either constants (implemented as strings) or variables.
+% There is no differentiation made between a \emph{voiced} constant (who says assertions),
+% and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
+% In AppPAL the subjects are usually apps or policies.
+% Apps do not utter facts;
+%   so there is no need to check them as potential speakers of \emph{can-say} statements.
+% Similarly if a constant never appears as a subject of a statement, as most speakers do not,
+%   there is no need to check it as a possible subject for a \emph{can-act-as} statement.
+% We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
+% This means we need to search less constants when attempting to replace a variable, speeding evaluation.
 
@@
 \begin{minipage}[b]{0.5\linewidth}
-\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
-# ac, d |-? q (using previous-results table rt)
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
 def evaluate(ac, rt, q, d)
@@
 
-# Try can-say and can-act-as
 def canSay_CanActAs(ac, rt, q, d)
@@
 \begin{minipage}[b]{0.5\linewidth}
-\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
-# Try cond rule
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
 def cond(ac, rt, q, d)
@@
 
-# Evaluate the antecedents
 def checkConditions(ac, rt, a, d)
@@
 
-\subsection{Policy Examples}
-\label{ssec:idioms}
+% \subsection{Policy Examples}
+% \label{ssec:idioms}
+%
+% When writing policies some schemes often come up.
+% We give examples of policies and queries and show how they can be implemented in AppPAL.
+%
+% \begin{description}
+%   \item[Policies for home and work]
+%     In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
+%     Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
+%     To implement this we add rules that use constraints to decide whether an app should be run.
+%     Alice already has the Emma's work policy as the default.
+%     Now she just needs to add an exception to the rule.
+%     \begin{lstlisting}
+% "alice" says App isRunnable
+%   if "home-policy" isMetBy(App)
+%   where TimeOfDay() > 17.00.
+%
+% "alice" says App isRunnable
+%   if "home-policy" isMetBy(App)
+%   where At("work") = false.
+%     \end{lstlisting}
+%
+%   \item[App white-listing by an IT department]
+%     Inside Emma's company employees in the IT department may white-list certain apps as runnable.
+%     Emma delegates to them to decide if an app is runnable.
+%     \begin{lstlisting}
+% "emma" says "it-department" can-say 0 App isRunnable.
+%     \end{lstlisting}
+%     Charlie and Diveena work in the IT department.
+%     When they have checked an app they say it is runnable.
+%     \begin{lstlisting}
+% "emma" says "charlie" can-act-as "it-department".
+% "emma" says "diveena" can-act-as "it-department".
+%
+% "charlie" says "com.facebook.katana" isRunnable.
+% "diveena" says "org.thoughtcrime.securesms" isRunnable.
+%     \end{lstlisting}
+%
+%   \item[Overpriviliged applications]
+%     Alice is particularly worried about apps stealing her data.
+%     She knows that certain apps need certain permissions.
+%     For example a photography app needs access to the camera.
+%     She checks each app's permissions carefully for things that seem unusual.
+%
+%     Alice's policy can be implemented by checking the permissions of the app, using constraints.
+%     \begin{lstlisting}
+% "alice" says App isRunnable
+%    if "permissions-policy" isMetBy(App).
+% "alice" says "permissions-policy" isMetBy(App)
+%    if App isAnApp
+%    where
+%       category(App, "Photography"),
+%       hasPermission(App, "LOCATION") = false,
+%       hasPermission(App, "CAMERA") = true.
+%     \end{lstlisting}
+%
+%     Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
+%     Alternately tools like Stowaway~\cite{Felt:2011kj} could be integrated to check apps.
 
-When writing policies some schemes often come up.
-We give examples of policies and queries and show how they can be implemented in AppPAL.
+%   \item[Digital Evidence]
+%     In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
+%     Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
+%     For example that no audio can be recorded without the users consent~\cite{Seghir:2014uq}.
+%     Using this tool someone generates a proof certificate that this property is met.
+%     This might be the developer of the app, a third party certification service, or someone else;
+%       when we check the proof we will discover if it is valid or not.
+%     \begin{lstlisting}
+% "alice" says Anyone can-say 0
+%   "recording-consent" shownIsMetBy(Evidence, App).
+%
+% "skb" says "recording-consent"
+%   shownIsMetBy("evicheck://evidence123", App).
+%
+% "alice" says "recording-consent" isMetBy(App)
+%   if "recording-consent" shownIsMetBy(Evidence, App)
+%   where
+%     evicheckCheckEvidence(Evidence, App) = true.
+%     \end{lstlisting}
+%     In the example here Alice allows anyone to present her with evidence for her policy.
+%     A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
+% \end{description}
 
-\begin{description}
-  \item[Policies for home and work]
-    In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
-    Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
-    To implement this we add rules that use constraints to decide whether an app should be run.
-    Alice already has the Emma's work policy as the default.
-    Now she just needs to add an exception to the rule.
+\subsection{Benchmarks}
+\label{ssec:benchmarks}
+
+Using the AppPAL library we wrote a benchmarking app that used policy that combined permissions checks and delegation relationships.
+The app measured the time it took to decide whether XX installed apps met the policy.
+We ran the app on X devices, and...
+
+\section{Demonstration}
+\label{sec:demonstation}
+
+AppPAL can be used as a tool for exploring policy compliance in app installation data sets.
+The \emph{privacy paradox} is that whilst people often have opinions about use of their personal data, they do not always follow through with their actions.
+Users seem to have opinions about apps~\cite{Sadeh:2014vq} but are they picking apps which follow their policies?
+If the privacy paradox occurs here too then it suggests the need for tools to at least make users aware they're breaking their own policies.
+
+In a user study of 725 Android users, Lin~\etal found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
+The \emph{Conservative} (C) users were uncomfortable allowing an app access to any personal data for any reason.
+The \emph{Unconcerned} (U) users felt okay allowing access to most data for almost any reason.
+The \emph{Advanced} (A) users were comfortable allowing apps access to location data but not if it was for advertising reasons.
+Opinions in the largest cluster, \emph{Fencesitters} (F), varied but were broadly against collection of personal data for advertising purposes.
+We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
+These simplify privacy policies identified as by Lin~\etal as we do not take into account the reason each app might have been collecting each permission.
+AppPAL can describe more complex policies however.
+
+\newcommand{\tabtitle}[1]{\textbf{\footnotesize #1}}
+\begin{center}
+  \begin{tabular}{ r l l l l }
+%                                       & \rothead{Conservative} & \rothead{Advanced} & \rothead{Fencesitter} & \rothead{Unconcerned} \\
+    \toprule
+    \tabtitle{Policy}                  & \tabtitle{C}           & \tabtitle{A}       & \tabtitle{F}          & \tabtitle{U}          \\
+    \midrule
+    \lstinline{GET_ACCOUNTS}           & \xmark                 & \xmark             & \xmark                & \xmark                \\
+    \lstinline{ACCESS_FINE_LOCATION}   & \xmark                 & \xmark             & \xmark                &                       \\
+    \lstinline{READ_CONTACT}           & \xmark                 & \xmark             & \xmark                &                       \\
+    \lstinline{READ_PHONE_STATE}       & \xmark                 & \xmark             &                       &                       \\
+    \lstinline{SEND_SMS}               & \xmark                 & \xmark             &                       &                       \\
+    \lstinline{ACCESS_COARSE_LOCATION} & \xmark                 &                    &                       &                       \\
+    \bottomrule
+  \end{tabular}
+\end{center}
+
+Like other vendors, McAfee classify malware into several categories.
+The \emph{malicious} and \emph{trojan} categories describe traditional malware.
+Other categories classify \ac{pup} such as aggressive adware.
+Using AppPAL we can write policies to differentiate between different kinds of malware, characterising users who allow dangerous apps or those who install merely ``unsavoury'' apps.
 \begin{lstlisting}
-"alice" says App isRunnable
-  if "home-policy" isMetBy(App)
-  where TimeOfDay() > 17.00.
-
-"alice" says App isRunnable
-  if "home-policy" isMetBy(App)
-  where At("work") = false.
+"user" says "mcafee" can-say
+  "malware" isKindOf(App).
+"mcafee" says "trojan" can-act-as "malware".
+"mcafee" says "pup" can-act-as "malware".
 \end{lstlisting}
 
-  \item[App white-listing by an IT department]
-    Inside Emma's company employees in the IT department may white-list certain apps as runnable.
-    Emma delegates to them to decide if an app is runnable.
-    \begin{lstlisting}
-"emma" says "it-department" can-say 0 App isRunnable.
-    \end{lstlisting}
-    Charlie and Diveena work in the IT department.
-    When they have checked an app they say it is runnable.
-    \begin{lstlisting}
-"emma" says "charlie" can-act-as "it-department".
-"emma" says "diveena" can-act-as "it-department".
+We want to test how well policies capture user behaviour.
+Installation data was taken from a partially anonymized\footnote{Users are replaced with incrementing numbers, app names are replaced with hashes to protect sensitive names.} database of installed apps captured by Carat~\cite{Oliner:2013ht}.
+By calculating the hash of known package names we see who installed what.
+The initial database has over 90,000 apps and 55,000 users.
+On average each user installed around 90 apps each; 4,300 apps have known names.
+Disregarding system apps (such as \texttt{com.android.vending}) and very common apps (Facebook, Dropbox, Whatsapp, and Twitter) we reduced the set to an average of 20 known apps per user.
+To see some variations in app type, we considered only the 44,000 users who had more than 20 known apps.
+Using this data, and the apps themselves taken from the Google Play Store and Android Observatory~\cite{Barrera:2012iba}, we checked which apps satisfied which policies.
 
-"charlie" says "com.facebook.katana" isRunnable.
-"diveena" says "org.thoughtcrime.securesms" isRunnable.
-    \end{lstlisting}
+\begin{figure}\centering
+  \subfigure[Use of policies modelling user behaviour.]{%
+    \includegraphics[width=0.48\linewidth]{./figures/lin.eps}
+    \label{fig:lin}}
+  \subfigure[Percentage of malware in installed apps for users installing some malicious apps.]{%
+    \includegraphics[width=0.48\linewidth]{./figures/malware.eps}
+    \label{fig:malware}}
+  \caption{Policy compliance graphs.}
+\end{figure}
 
-  \item[Overpriviliged applications]
-    Alice is particularly worried about apps stealing her data.
-    She knows that certain apps need certain permissions.
-    For example a photography app needs access to the camera.
-    She checks each app's permissions carefully for things that seem unusual.
+Figure~\ref{fig:lin} shows that the very few users follow these policies, but a few users who do seem to be installing apps meeting these policies most of the time.
+For the unconcerned policy (the most permissive) only 1,606 users (4\%) had 50\% compliance;
+  and only 120 users (0.3\%) had 80\% compliance.
+For the stricter conservative policy only 60 users were complying half the time, and just 7 more than 80\% of the time.
+This suggests that while users may have privacy preferences the majority are not attempting to enforce them.
 
-    Alice's policy can be implemented by checking the permissions of the app, using constraints.
-    \begin{lstlisting}
-"alice" says App isRunnable
-   if "permissions-policy" isMetBy(App).
-"alice" says "permissions-policy" isMetBy(App)
-   if App isAnApp
-   where
-      category(App, "Photography"),
-      hasPermission(App, "LOCATION") = false,
-      hasPermission(App, "CAMERA") = true.
-    \end{lstlisting}
+\begin{figure}\centering
+  \includegraphics[width=\linewidth]{./figures/c_v_pup.eps}
+  \caption{Compliance with the advanced policy and non-\ac{pup} policy.}
+  \label{fig:versus}
+\end{figure}
 
-    Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
-    Alternately tools like Stowaway~\cite{Felt:2011kj} could be integrated to check apps.
+We found 1\% of the users had a \ac{pup} or malicious app installed.
+Figure~\ref{fig:malware} shows that infection rates for \ac{pup}s and malware is small.
+A user is 3 times more likely to have a \ac{pup} installed than malware.
+Only 9 users had both a \ac{pup} and malware installed.
+Users who were complying more than half the time with the conservative or advanced policies complied with the malware or \ac{pup} policies fully (\autoref{fig:versus}).
+This is significant (P-value $< 0.05$) and suggests that users who pick their apps carefully are less likely to experience malware.
 
-  \item[Digital Evidence]
-    In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
-    Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
-    For example that no audio can be recorded without the users consent~\cite{Seghir:2014uq}.
-    Using this tool someone generates a proof certificate that this property is met.
-    This might be the developer of the app, a third party certification service, or someone else;
-      when we check the proof we will discover if it is valid or not.
-    \begin{lstlisting}
-"alice" says Anyone can-say 0
-  "recording-consent" shownIsMetBy(Evidence, App).
+Most users seem to use apps irrespective of how uncomfortable they are with the permissions they request.
+A small set of users do seem to enforce these policies at least some of the time however.
+Exploring where this disconnect comes from is an avenue for future research.
+Do users not understand the relationship between apps and permissions~\cite{Felt:2012hma}?
+Is enforcing them informally too difficult?
 
-"skb" says "recording-consent"
-  shownIsMetBy("evicheck://evidence123", App).
-
-"alice" says "recording-consent" isMetBy(App)
-  if "recording-consent" shownIsMetBy(Evidence, App)
-  where
-    evicheckCheckEvidence(Evidence, App) = true.
-    \end{lstlisting}
-    In the example here Alice allows anyone to present her with evidence for her policy.
-    A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
-\end{description}
+We claim that AppPAL can capture the differences in informal user policies.
+Using AppPAL we have written short policies describing user behaviour and used these to identify the users following them to varying degrees.
+Some limitations of our results include:
+\begin{itemize}
+\item We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
+  So a user may have apps installed that break the policy without us knowing.
+\item Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
+  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may actually be more conservative than our analysis suggests.
+\end{itemize}
 
@@
 
-Whilst this approach is great for malware, it is less suitable for finding \ac{pus}.
-The behaviours and permissions \ac{pus} apps display aren't necessarily malicious.
+This approach could help identify malware, but it is less suitable for detecting \ac{pup}.
+The behaviours and permissions \ac{pup} apps display aren't necessarily malicious.
 One user may consider apps which need in-app-purchases to play malware, but another may enjoy them.
-AppPAL tries to stop these \ac{pus} apps.
+AppPAL tries to stop these \ac{pup} apps.
 Because we can use external checking tools which go further than permissions checks, our policies can be richer.
 By allowing delegation relationships we can understand the provenance and trust relationships in these rules.
-
-Work by Lin~\etal looked at modelling users' willingness to grant different kinds of permissions~\cite{Sadeh:2014vq}.
-They found four clusters of different users each with different kinds of policies:
-  conservatives who do not like granting many permissions,
-  advanced users who seem to have more complex policies,
-  unconcerned users who'll comfortably grant anything,
-  and fence-sitters (about half of all users) who seemed not to have varying opinions.
-This suggests that users are applying policies and do care how apps behave.
-Lin~\etal speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
-This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
-Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
-We have created an Android app that can check approximations of Lin~\etal's policies against apps on a users device.
-This gives us a demonstration of the power of AppPAL using known policies\footnote{\url{https://github.com/bogwonch/apppal-checker}}.
 

### 2015-09-04 paper.tex
edited

@@
   We present AppPAL: a policy language to enforce policies about apps.
-  AppPAL Policies use statements from third parties and delegation relationships to
-  give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
-  % \end{itemize}
+  AppPAL policies use statements from third parties and delegation relationships to give us a rigorous and flexible framework for enforcing and comparing app policies; and the trust relationships surrounding them.
 \end{abstract}
@@
 Finding the right apps can be tricky.
-Users need to work out which apps are well written, which are not going to abuse their data, which will work in the way they want
+Users need to discover which apps are well written, which are not going to abuse their data
   and to find the apps which suit how they want to use their device.
@@
 Android apps show a list of permissions when they're first installed.
-Soon new Android apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
-Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
+Soon new apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
+Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
 Ultimately the decision which apps to use and which permissions to grant must be made by the user.
@@
 Not all apps are suitable.
-A large amount of \ac{pup} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
-Employees are increasingly using their own phones for work (\ac{byod}).
+Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
+Employees are increasingly using their own phones for work (a system known as \ac{byod}).
 An employer may wish to restrict which apps their employees can use.
@@
 
-These policies can only be enforced by the users continuously making decisions guided by these policies when prompted about apps.
+These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
 This is error-prone.
-Mistakes can be made.
 We believe this can be improved.
@@
 We present AppPAL, an authorization logic for reasoning about apps.
-The language is an instantiation of SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
+The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
@@
 "alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
-"alice" says "bob" can-act-as "it-department".
+"alice" says "alice" can-act-as "it-department".
 \end{lstlisting}
@@
 \noindent
-Specifically our contribution is:
+In this paper we have:
 \begin{itemize}
@@
     and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
-
-  \item Implemented the AppPAL language on Android and the JVM.
-    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).
-
-  \item Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
+  \item
+    Implemented the AppPAL language as a library on Android and the JVM.
+    We show how the language can describe properties of Android apps and Android security policies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).
+  \item
+    Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
 \end{itemize}
@@
 She might even sign a document promising never to break the rules within the policy.
-This is error-prone though---what if she makes a mistake or misses an app that breaks her policy?
+This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
 Emma's policy could be implemented using existing tools.
-The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store (and in the newest version of Android let Emma set the permissions of each app on an app by app basis~\cite{AndroidMPermission:2015uq}).
-AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and formalize the decision making process.
+\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}).
+AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and provide a rigorous means for deciding whether an app meets a policy.
 Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
 These could be used to used to ensure no location data is ever obtained.
-Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are ever leaked to the web.
+Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are never leaked to the web.
 Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.
 
-Whilst we could implement Emma's policy using existing tools, it is a clumsy solution.
-They are not flexible: If Emma changes her policy or Alice changes jobs she needs to recheck and then to alter and remove the software on her phone accordingly.
-It isn't clear what an app must do to be run, or what checks have been done if it already running on the phone.
+Whilst we could implement Emma's policy using existing tools it is a clumsy solution---they are not flexible.
+If Emma changes her policy or Alice changes jobs she needs to recheck her apps and then alter or remove the software on her phone to ensure compliance.
+It isn't clear what an app must do to be run, or what checks have been done if it is already running on the phone.
 The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.
@@
 Emma shouldn't be able to overly control what Alice does in her private life.
-Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
-Some mobile OSs allow app permissions to be enabled and disable at run time.
+Alice might not be allowed to use location tracking apps at work but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
+Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disable at run time.
 Can we enforce different policies at different times or locations?
@@
 Our research looks at the problem of picking software.
-Given there are some apps you want to install and run and others you do not want to, at least some of the time:
+Given there are some apps you want to install and run and others you do not:
   how can you express your preferences in such a way that they can be enforced automatically?
 How can we translate policy documents from natural language into a machine checkable form?
-Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?
+How can we show the trust relationships used to make these decisions clearly and precisely?
 
@@
 People have policies which are enforced by AppPAL on their devices.
+They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
 The device can make use of vetting services which run tools to infer complex properties about apps.
-Users can buy from stores which ensure the only apps they see are the apps which meet their policies.
-Policies can be combined
+Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
+Developers could decide which stores to sell their apps in on the basis of policies about stores.
 
@@
 
-In \autoref{sec:problem} Alice and Emma's had policies they wanted to enforce but no means to do so.
-Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
+In \autoref{sec:problem} Alice and Emma had policies they wanted to enforce but no means to do so.
+Instead of using several different tools to enforce Emma's policy disjointedly, we could use an authorization logic.
 In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
@@
 
-AppPAL is an instantiation of SecPAL~\cite{Becker:2006vh} for describing app policies.
 SecPAL is a logic of authorization for access control decisions in distributed systems.
@@
 SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
-We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.
+We present AppPAL as a modified form of SecPAL, targeting apps on mobile devices.
 
@@
     \begin{lstlisting}[numbers=left,
-                       escapeinside={@}{@}]
+                       escapeinside={@}{@},
+                       basicstyle=\scriptsize\ttfamily,
+                       stringstyle=\scriptsize\sffamily,
+                       keywordstyle=\scriptsize\slshape]
 "alice" says "emma" can-say inf
@@
                        escapeinside={@}{@},
-                       firstnumber=15]
+                       firstnumber=15,
+                       basicstyle=\scriptsize\ttfamily,
+                       stringstyle=\scriptsize\sffamily,
+                       keywordstyle=\scriptsize\slshape]
 "emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
@@
 Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
-If Emma can be convinced all these policies are met then he will say the \code{App isRunnable}.
+If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
 In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
@@
 Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
-To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
+To meet Emma's policy the AppPAL checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
 Specifically it needs:
@@
 For the forth statement the checker needs to check the permissions of the app.
-It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if running on device.
+It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if it is running on a device.
 
 In this scenario we have imagined Alice wanting to check the apps as she installs them.
-Alternatively we could imagine Emma wanting a personalised app store where all apps sold meet her policy.
+We could also imagine Emma wanting a personalised app store where all apps sold meet her policy.
 With AppPAL this can be implemented by taking an existing store and selectively offering only the apps which will meet the user's policy.
-This gives us a \emph{filtered store}.
-From an existing set of apps we produce a personalised store that meets a pre-defined policy.
+This gives us a \emph{filtered store} which, from an existing set of apps, we get a personalised store that only sells apps that meet a policy.
 
@@
 
-AppPAL is an instantiation of SecPAL for Android app installation policies.
-It is implemented as a library for Android and Java.
+AppPAL is implemented as a library for Android and Java.
 The parser is implemented using ANTLR4.
-Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.
-
 The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.
@@
 \begin{figure}
+  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
+  \newcommand{\smalltext}[1]{\text{\ttfamily\footnotesize #1}}
   \centering
   \begin{equation*}\small
-    \begin{array}{r l}
-      \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} &
-      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isRunnable}}^{\text{predicate}}}^{\text{fact}} \\
-      & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
-      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = true}}^{\text{constraint}}.
+    \begin{array}{r l}\footnotesize
+      \overbrace{\smalltext{"user"}}^{\bracetext{speaker}} &
+      \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
+      & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
+      & \overbrace{\smalltext{ where hasPermission(App, "INTERNET") = true}}^{\bracetext{constraint}}.
     \end{array}
@@
 
-In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of
-predicates, and constraints for their SecPAL open.  With AppPAL we make
-explicit our predicates and how they relate to Android.  Some of these
-predicates require arguments.  AppPAL policies typically make use of the
-predicates and constraints in \autoref{tab:predicates}.  Additional predicates
-can be created in the policy files, but adding or modifying constraints
-requires a code change\footnote{But a small one.  The Android version of the
-\texttt{hasPermission} constraint, for example, uses the Android package
-manager to determine what permissions an app requests, but the Java version
-uses the Android platform tools.  This required a code change to one function.}
+In Becker~\etal's work~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
+With AppPAL we make explicit our predicates and how they relate to Android.
+AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
+Additional predicates can be created in the policy files, but adding or modifying constraints required a code change\footnote{The Android version of the \texttt{hasPermission} constraint, for example, uses the Android package manager to determine what permissions an app requests, but the Java version uses the Android platform tools.
+This requires a small code change.}
 
@@
   \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
+  Tool\texttt{Check(}App, Property) &  General form of a constraint used to run a static analysis tool over an app. \\
   \bottomrule
@@
 Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
-It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
+It allows us to describe multiple means of making the same decision, and provide backup routes when one fails.
 Some static analysis tools are not quick to run.
@@
 
-Earlier we described a \emph{no-tracking-policy} to prevent a user's location being tracked.
-In Bob's policy we checked this by checking the permissions of the app.
+In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
+In Emma's policy we checked this using the app's permissions.
 If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
@@
 \begin{lstlisting}
-"bob" says "no-locations-permissions"
+"emma" says "no-locations-permissions"
   can-act-as "no-tracking-policy".
 
-"bob" says "no-locations-permissions" isMetBy(App)
+"emma" says "no-locations-permissions" isMetBy(App)
   if App isAnApp
@@
 
-"bob" says "location-taint-analysis"
+"emma" says "location-taint-analysis"
   can-act-as "no-tracking-policy".
 
-"bob" says "location-taint-analysis" isMetBy(App)
+"emma" says "location-taint-analysis" isMetBy(App)
   if App isAnApp
   where
-    taintDroidCheckLeak(App, "Location", "Internet") = false.
+    taintDroidCheck(App, "Location", "Internet") = false.
 \end{lstlisting}
@@
 Sometimes we might want to use location data.
-For instance Bob might want to check that Alice is at her office.
-Bob might track Alice using a location tracking app.
-Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) he is happy to relax the policy.
+For instance Emma might want to check that Alice is at her office.
+Emma might track Alice using a location tracking app.
+Provided the app only talks to Emma, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) she is happy to relax the policy.
 
 \begin{lstlisting}
-"bob" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
-"bob" says "relaxed-no-tracking-policy" isMetBy(App)
+"emma" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
+"emma" says "relaxed-no-tracking-policy" isMetBy(App)
   if App hasCategory("tracking")
@@
     mallodroidSSLCheck(App) = false,
-    connections(App) = "[https://bob.com]".
+    connectionsCheck(App, "[https://emma.com]") = true.
 \end{lstlisting}
@@
   with a relaxed version of the policy,
-  or by Bob directly saying the app meets it.
-When we come to check the policy if any of these ways give us a positive result we can can stop our search.
+  or by Emma directly saying the app meets it.
+When we come to check the policy if any of these ways give us a positive result we can stop our search.
 
@@
 
-Since AppPAL is an instantiation of SecPAL the policy checking rules are the same.
-We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm, as no DatalogC library exists for Android.
+AppPAL has the same policy inference rules as SecPAL.
+We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
 We have implemented the rules directly in Java.
 Pseudo-code is given in \autoref{fig:pseudocode}.
-Like Becker~\etal we make use of an assertion context to store known statements.
-We also store intermediate results to avoid re-computation.
+
+Like Becker~\etal~we make use of an assertion context to store known statements.
+We also cache intermediate results, and ground facts to avoid re-computation.
 On a mobile device memory is at a premium.
@@
 For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
-
-% \begin{figure}
-% \[\footnotesize
-% % AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
-% % ----------------------------------------------------------- (can-act-as)
-% %                 AC, D ⊧ A says B verbphrase
-% \infer[\text{\tiny can-act-as}]{
-%   AC, D \models \text{\tt A says B verbphrase}}{%
-%   AC, D \models \text{\tt A says B can-act-as C} & %
-%   AC, D \models \text{\tt A says C verbphrase}}
-% \]
-% \[\footnotesize
-% % AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
-% % ----------------------------------------------------- (can-say)
-% %                   AC, ∞ ⊧ A says fact
-% \infer[\text{\tiny can-say}]{
-%   AC, \infty \models \text{\tt A says fact}}{%
-%   AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
-%   AC, D \models \text{\tt B says fact}}
-% \]
-% \[\footnotesize
-% % (A says fact if fact₁, …, factₖ, c) ∈ AC
-% % AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
-% % ----------------------------------------------------------- (cond)
-% %                    AC, D ⊧ A says factθ
-% \infer[\text{\tiny cond}]{
-%   AC, D \models \text{\tt A says fact$\theta$}}{
-%   \begin{array}{c}
-%     (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
-%     \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
-%   \end{array} &
-%   \models \mathtt{c}\theta &
-%   \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
-% \]
-% \caption{AppPAL's evaluation rules.}
-% \label{fig:rules}
-% \end{figure}
-
 This gives us greater control of the evaluation and how the assertion context is created.
@@
 
-\begin{figure}
-\begin{minipage}[b]{0.5\linewidth}
-\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
+\begin{figure}\centering
+\begin{minipage}[b]{0.49\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
 def evaluate(ac, rt, q, d)
@@
   p = cond(ac, rt, q, d)
-  return (Proven, rt.update q, d, p) if p.isValid
+  if p.isValid then
+    return (Proven, rt.update q, d, p)
   p = canSay_CanActAs(ac, rt, q, d)
-  return (Proven, rt.update q, d, p) if p.isValid
+  if p.isValid then
+    return (Proven, rt.update q, d, p)
+  else
     return (Failure, rt.update q, d, Failure)
-
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[b]{0.49\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
 def canSay_CanActAs(ac, rt, q, d)
@@
 \end{minipage}
-\begin{minipage}[b]{0.5\linewidth}
-\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
+
+\begin{minipage}[b]{0.49\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
 def cond(ac, rt, q, d)
@@
   return Failure
-
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}[b]{0.49\linewidth}
+\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
 def checkConditions(ac, rt, a, d)
-  getVarSubstitutions(a,ac.constants).each do |s|
-    sa = s.substitute a
+  getVarSubs(a,ac.constants).each do |s|
+    sa = s.sub a
     if sa.antecedents.all
@@
 
-Using the AppPAL library we wrote a benchmarking app that used policy that combined permissions checks and delegation relationships.
-The app measured the time it took to decide whether XX installed apps met the policy.
-We ran the app on X devices, and...
+We envisage AppPAL running on a mobile phone checking apps.
+For it to be a viable solution the policy checking performance must be reasonable: capable of checking apps against a policy as quickly as possible.
+The policy checking search procedure is at its slowest when performing large repeated delegations.
+Two synthetic benchmarks were created to check that the search procedure performed acceptably.
+Each benchmark consisted of a repeated chain of delegations.
+The \emph{straight} benchmark consists of a single long chain of delegation.
+The \emph{forking} benchmark consists of a binary tree of principals delegating to each other.
+These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by forking even more.
+A snippet of each policy is shown in \autoref{fig:policy-snippet}.
+
+On a \emph{Nexus 4} checking times are measured in seconds when there are hundreds of delegations, and in minutes when there are thousands (\autoref{fig:benchmarks}).
+We have only used a few delegations per decision when describing hypothetical user policies.
+Since the benchmarks show that long chains of delegation can be used, we believe the performance is acceptable.
+
+\begin{figure}\centering
+  \begin{minipage}{0.49\linewidth}
+    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
+    '1' says '2' can-say App isInstallable.
+    '2' says '3' can-say App isInstallable.
+    '3' says '4' can-say App isInstallable.
+    '4' says '5' can-say App isInstallable.
+    '5' says '6' can-say App isInstallable.
+    '6' says '7' can-say App isInstallable.
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}
+    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
+    '1' says '2' can-say App isInstallable.
+    '1' says '3' can-say App isInstallable.
+    '2' says '4' can-say App isInstallable.
+    '2' says '5' can-say App isInstallable.
+    '3' says '6' can-say App isInstallable.
+    '3' says '7' can-say App isInstallable.
+    \end{lstlisting}
+  \end{minipage}
+  \caption{Straight and forking policies used for benchmarking.}
+  \label{fig:policy-snippet}
+\end{figure}
+
+\begin{figure}
+  \begin{minipage}{0.49\linewidth}
+    \begin{tabular}{c c c c}
+       \toprule
+       Policy   & Nodes & Parse (s) & Check (s) \\
+       \midrule
+       Straight & 10    & 0.01      & 0.04      \\
+       Straight & 100   & 0.05      & 1.00      \\
+       Straight & 500   & 0.32      & 21.07     \\
+       Straight & 1000  & 0.43      & 87.75     \\
+       \midrule
+       Forking  & 256   & 0.10      & 1.93      \\
+       Forking  & 512   & 0.23      & 7.47      \\
+       Forking  & 1024  & 0.45      & 28.29     \\
+       Forking  & 2048  & 0.85      & 117.68    \\
+       \bottomrule
+    \end{tabular}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}
+    \includegraphics[width=\linewidth]{figures/benchmarks.eps}
+  \end{minipage}
+  \caption{Benchmarking results on a Nexus 4 Android phone.}
+  \label{fig:benchmarks}
+\end{figure}
 
@@
 
-AppPAL can be used as a tool for exploring policy compliance in app installation data sets.
-The \emph{privacy paradox} is that whilst people often have opinions about use of their personal data, they do not always follow through with their actions.
-Users seem to have opinions about apps~\cite{Sadeh:2014vq} but are they picking apps which follow their policies?
-If the privacy paradox occurs here too then it suggests the need for tools to at least make users aware they're breaking their own policies.
+Throughout we have asserted that user's have policies and that there is a need for policy enforcement tools.
+Corporate mobile security \ac{BYOD policies} have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
+In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
+Using app installation data from Carat~\cite{Oliner:2013ht} we used AppPAL to find the apps satisfying each policy Lin~\etal~identified and measured the extent each user was following them.
 
-In a user study of 725 Android users, Lin~\etal found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
+Lin~\etal~identified four types of user.
 The \emph{Conservative} (C) users were uncomfortable allowing an app access to any personal data for any reason.
@@
 We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
-These simplify privacy policies identified as by Lin~\etal as we do not take into account the reason each app might have been collecting each permission.
-AppPAL can describe more complex policies however.
+These simplify the privacy policies identified as by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
 
@@
 
+Malware is available for Android.
 Like other vendors, McAfee classify malware into several categories.
@@
 Other categories classify \ac{pup} such as aggressive adware.
-Using AppPAL we can write policies to differentiate between different kinds of malware, characterising users who allow dangerous apps or those who install merely ``unsavoury'' apps.
+Using AppPAL we can write policies to differentiate between different kinds of malware, characterising users who allow dangerous apps and those who install poor quality ones.
 \begin{lstlisting}
@@
 \end{lstlisting}
+If a user is enforcing a privacy policy we might also expect them to install less malware.
+We can check this by using AppPAL policies to find the malware and again measure the number of malwares each user had installed.
 
@@
 Installation data was taken from a partially anonymized\footnote{Users are replaced with incrementing numbers, app names are replaced with hashes to protect sensitive names.} database of installed apps captured by Carat~\cite{Oliner:2013ht}.
-By calculating the hash of known package names we see who installed what.
+By calculating the hashes of known package names we see who installed what.
 The initial database has over 90,000 apps and 55,000 users.
@@
 
-Figure~\ref{fig:lin} shows that the very few users follow these policies, but a few users who do seem to be installing apps meeting these policies most of the time.
-For the unconcerned policy (the most permissive) only 1,606 users (4\%) had 50\% compliance;
-  and only 120 users (0.3\%) had 80\% compliance.
-For the stricter conservative policy only 60 users were complying half the time, and just 7 more than 80\% of the time.
+Figure~\ref{fig:lin} shows that very few users follow Lin~\etal's policies most of the time.
+Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (often referred to as the \emph{privacy paradox}).
+A few users, however, did seem to be installing apps meeting these policies most of the time.
 This suggests that while users may have privacy preferences the majority are not attempting to enforce them.
+Policy enforcement tools, like AppPAL, can help users enforce their own policies which they cannot do easily using the current means available to them.
 
 \begin{figure}\centering
-  \includegraphics[width=\linewidth]{./figures/c_v_pup.eps}
+  \includegraphics[width=0.5\linewidth]{./figures/c_v_pup.eps}
   \caption{Compliance with the advanced policy and non-\ac{pup} policy.}
@@
 We found 1\% of the users had a \ac{pup} or malicious app installed.
-Figure~\ref{fig:malware} shows that infection rates for \ac{pup}s and malware is small.
-A user is 3 times more likely to have a \ac{pup} installed than malware.
-Only 9 users had both a \ac{pup} and malware installed.
+Figure~\ref{fig:malware} shows that infection rates for \ac{pup}s and malware is low;
+though a user is 3 times more likely to have a \ac{pup} installed than malware.
 Users who were complying more than half the time with the conservative or advanced policies complied with the malware or \ac{pup} policies fully (\autoref{fig:versus}).
-This is significant (P-value $< 0.05$) and suggests that users who pick their apps carefully are less likely to experience malware.
+This suggests that policy enforcement is worthwhile: users who can enforce policies about their apps experience less malware.
+%This is significant (P-value $< 0.05$) and suggests that users who pick their apps carefully are less likely to experience malware.
 
-Most users seem to use apps irrespective of how uncomfortable they are with the permissions they request.
-A small set of users do seem to enforce these policies at least some of the time however.
-Exploring where this disconnect comes from is an avenue for future research.
-Do users not understand the relationship between apps and permissions~\cite{Felt:2012hma}?
-Is enforcing them informally too difficult?
-
-We claim that AppPAL can capture the differences in informal user policies.
 Using AppPAL we have written short policies describing user behaviour and used these to identify the users following them to varying degrees.
@@
 \begin{itemize}
-\item We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
+\item
+  We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
   So a user may have apps installed that break the policy without us knowing.
-\item Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
-  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may actually be more conservative than our analysis suggests.
+\item
+  Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
+  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
+\item
+  The AppPAL policies we used are a simplification of the identified privacy preferences.
+  User's could be following policies more nuanced than our implementation.
+  We believe our approximation is valid, however, as users will not know the precise reason an app uses a permission, and may decline to install it anyway.
+  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND_SMS} permission to run).
 \end{itemize}
@@
 Google also offer the \emph{Device Policy for Android} app.
-This lets businesses to configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
+This lets businesses configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
 It cannot be used to describe policies about apps, or describe trust relationships, however.
 
-The SecPAL language was initially used for access control in distributed systems.
+The SecPAL language is designed for access control in distributed systems.
 We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\cite{Hallett:2014un}.
 It has already been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
-Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and ultimately spawned the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
 Gruevich and Neeman showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
@@
 The distinction is that in AppPAL whilst the user thinks the app could be installed we do not know for sure whether the user has installed it.
-With the DKAL we can guarantee that the action was completed.
+With DKAL we can guarantee that the action was completed.
 
-One tool, Kirin~\cite{Enck:2009ko}, also created a policy language and tool for enforcing app installation policies.
-Kirin's policies were concerned with preventing malware.
+Kirin~\cite{Enck:2009ko} is a policy language and tool for enforcing app installation policies preventing malware.
 Policy authors could specify combinations of permissions that should not appear together.
@@
 \end{lstlisting}
-\begin{lstlisting}
-"user" says "no-write-send-sms" isMetBy(App)
-  where hasPermission(App, "SEND_SMS") = false.
-"user" says "no-write-send-sms" isMetBy(App)
-  where hasPermission(App, "WRITE_SMS") = false.
-\end{lstlisting}
-\caption{Kirin and AppPAL policies for stopping apps monetized by premium rate text messages.}
+% \begin{lstlisting}
+% "user" says "no-write-send-sms" isMetBy(App)
+%   where hasPermission(App, "SEND_SMS") = false.
+% "user" says "no-write-send-sms" isMetBy(App)
+%   where hasPermission(App, "WRITE_SMS") = false.
+% \end{lstlisting}
+\caption{Kirin policy stopping apps monetized by premium rate text messages.}
 \label{fig:kirin}
@@
 
-This approach could help identify malware, but it is less suitable for detecting \ac{pup}.
-The behaviours and permissions \ac{pup} apps display aren't necessarily malicious.
+This approach could help identify malware, but it is less suitable for detecting \acp{pup}.
+The behaviours and permissions \ac{pup} displays aren't necessarily malicious.
 One user may consider apps which need in-app-purchases to play malware, but another may enjoy them.
-AppPAL tries to stop these \ac{pup} apps.
-Because we can use external checking tools which go further than permissions checks, our policies can be richer.
+With Kirin we are restricted to permitting or allowing apps.
+AppPAL is more general: we can describe more scenarios than just permit or allow, and use more app information than just permissions.
 By allowing delegation relationships we can understand the provenance and trust relationships in these rules.
+With AppPAL we can incorporate static analysis results and take into account the user's location and other constraints.
 
@@
 
-\section{Conclusions and further work \comment{$\frac{1}{2}$ page}}
+\section{Conclusions and further work}
 
 We have presented AppPAL: an authorization logic for describing app installation policies.
-The language is implemented in Java and runs on Android using a custom evaluation algorithm.
-This lets us to enforce app installation policies on Android devices.
+AppPAL lets us enforce app installation policies on Android devices.
 We have shown how the language can be used to describe an app installation policy;
@@
 We would use AppPAL to check apps meet policies before installation.
-Unfortunately the API is protected and it would require the phone to be rooted to run there.
+Unfortunately the API is protected and it would require the phone to be rooted.
 Alternatively AppPAL could be integrated as a service to reconfigure app permissions.
-The latest version of Android\footnote{Called \emph{Android M}.} is moving to a more iOS like permissions model where permissions can be granted and revoked at any time.
+The latest version of Android\footnote{Called \emph{Android M}.} is moving to an iOS like permissions model where permissions can be granted and revoked at any time.
 These will be manually configurable by the user through the settings app.
 We can imagine AppPAL working to reconfigure these settings (and set their initial grant or deny states) based on a user's policy, as well as the time of day or the user's location.
-A policy could deny pop-up notices while a user is driving for example.
+A policy could deny notifications while a user is driving for example.
 
 Developing, and testing, policies for users is a key next step.
-Here we described a policy being specified by a company boss.
+Here we described a policy being specified by a user's employer.
 For most end-users writing a policy in a formal language is too much work.
@@
 Users policies can be enforced more reliably, and with less interaction;
-making apps more pleasant for everyone and helping to reduce user fatigue.
+  making apps more pleasant for everyone.
 

### 2015-09-20 paper.tex
updates

@@
   \newcommand{\bracetext}[1]{\text{\sffamily #1}}
-  \newcommand{\smalltext}[1]{\text{\ttfamily\footnotesize #1}}
+  \newcommand{\smalltext}[1]{\text{\ttfamily\scriptsize #1}}
   \centering
+  \begin{minipage}{0.49\linewidth}
     \begin{equation*}\small
       \begin{array}{r l}\footnotesize
-      \overbrace{\smalltext{"user"}}^{\bracetext{speaker}} &
+        \overbrace{\smalltext{`user'}}^{\bracetext{speaker}} &
         \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
         & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
-      & \overbrace{\smalltext{ where hasPermission(App, "INTERNET") = true}}^{\bracetext{constraint}}.
+        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
       \end{array}
     \end{equation*}
-  \caption{Structure of an AppPAL assertion.}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}
+    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize,keywordstyle=\ttfamily\scriptsize]
+      Assertion := <E> says <Fact>
+                     (if (<Fact>,)+)?
+                     (where <Constraint>)?.
+      Fact := <E> (isRunable | ...)
+            | <E> can-say (0 | inf) <Fact>
+            | <E> can-act-as <E>
+      E := Variable | `constant'
+    \end{lstlisting}
+  \end{minipage}
+  \caption{Structure and simplified grammar of an AppPAL assertion.}
   \label{fig:assertion}
@@
   we can fetch the assertion that the subject is an app based on the app in question.
-Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
-  it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
+% Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
+%   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
 
@@
     if (u = q.unify a.consequent) &&
-       (a = u.substitution a).variables == none
+       (a = u.sub a).variables == none
       return checkConditions ac, rt, a, d
@@
 
-We envisage AppPAL running on a mobile phone checking apps.
-For it to be a viable solution the policy checking performance must be reasonable: capable of checking apps against a policy as quickly as possible.
-The policy checking search procedure is at its slowest when performing large repeated delegations.
-Two synthetic benchmarks were created to check that the search procedure performed acceptably.
-Each benchmark consisted of a repeated chain of delegations.
-The \emph{straight} benchmark consists of a single long chain of delegation.
-The \emph{forking} benchmark consists of a binary tree of principals delegating to each other.
-These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by forking even more.
-A snippet of each policy is shown in \autoref{fig:policy-snippet}.
+We envisage AppPAL running on a mobile phone checking apps when they are installed.
+Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
+Downloading and installing an app takes roughly 30s on an Android phone.
+If checking a policy delays this even further user's may become annoyed and disable AppPAL.
+
+The policy checking search procedure is at its slowest when having to delegate repeatedly;
+  with the depth of the delegation tree being the biggest factor for slowing the search.
+Synthetic benchmarks were created to check that the search procedure performed acceptably.
+Each benchmark consisted of a chain of delegations.
+The \emph{1 to 1} benchmark consists of a single long chain of delegation.
+The \emph{1 to 2} benchmark consists of a binary tree of principals delegating to each other; where each principal delegated to 2 others.
+Similarly the \emph{1 to 3} benchmark consists of a tree where each principal delegated to 3 others.
+These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by delegating even more.
+% A snippet of the \emph{1 to 1} and \emph{1 to 3} policies is shown in \autoref{fig:policy-snippet}.
+For each benchmark we controlled the number of principals in the policy file:
+  as the number of principals increased so did the size of the policy.
 
@@
 
-\begin{figure}\centering
-  \begin{minipage}{0.49\linewidth}
-    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
-    '1' says '2' can-say App isInstallable.
-    '2' says '3' can-say App isInstallable.
-    '3' says '4' can-say App isInstallable.
-    '4' says '5' can-say App isInstallable.
-    '5' says '6' can-say App isInstallable.
-    '6' says '7' can-say App isInstallable.
-    \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}{0.49\linewidth}
-    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
-    '1' says '2' can-say App isInstallable.
-    '1' says '3' can-say App isInstallable.
-    '2' says '4' can-say App isInstallable.
-    '2' says '5' can-say App isInstallable.
-    '3' says '6' can-say App isInstallable.
-    '3' says '7' can-say App isInstallable.
-    \end{lstlisting}
-  \end{minipage}
-  \caption{Straight and forking policies used for benchmarking.}
-  \label{fig:policy-snippet}
-\end{figure}
+% \begin{figure}\centering
+%   \begin{minipage}{0.49\linewidth}
+%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
+%     '1' says '2' can-say App isInstallable.
+%     '2' says '3' can-say App isInstallable.
+%     '3' says '4' can-say App isInstallable.
+%     '4' says '5' can-say App isInstallable.
+%     '5' says '6' can-say App isInstallable.
+%     '6' says '7' can-say App isInstallable.
+%     \end{lstlisting}
+%   \end{minipage}
+%   \begin{minipage}{0.49\linewidth}
+%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
+%     '1' says '2' can-say App isInstallable.
+%     '1' says '3' can-say App isInstallable.
+%     '2' says '4' can-say App isInstallable.
+%     '2' says '5' can-say App isInstallable.
+%     '3' says '6' can-say App isInstallable.
+%     '3' says '7' can-say App isInstallable.
+%     \end{lstlisting}
+%   \end{minipage}
+%   \caption{Snippets of the \emph{1 to 1} and \emph{1 to 3} policies used for benchmarking.}
+%   \label{fig:policy-snippet}
+% \end{figure}
 
 \begin{figure}
-  \begin{minipage}{0.49\linewidth}
-    \begin{tabular}{c c c c}
+  \begin{minipage}{0.40\linewidth}
+    \scriptsize
+    \begin{tabular*}{0.80\linewidth}{c c r@{.}l}
       \toprule
-       Policy   & Nodes & Parse (s) & Check (s) \\
+      Delegations & Principals & \multicolumn{2}{c}{Time (s)} \\
       \midrule
-       Straight & 10    & 0.01      & 0.04      \\
-       Straight & 100   & 0.05      & 1.00      \\
-       Straight & 500   & 0.32      & 21.07     \\
-       Straight & 1000  & 0.43      & 87.75     \\
+      1 to 1 & 10   &  0&01 \\
+      1 to 1 & 100  &  1&00 \\
+      1 to 1 & 500  & 20&90 \\
+      1 to 1 & 1000 & 88&73 \\
       \midrule
-       Forking  & 256   & 0.10      & 1.93      \\
-       Forking  & 512   & 0.23      & 7.47      \\
-       Forking  & 1024  & 0.45      & 28.29     \\
-       Forking  & 2048  & 0.85      & 117.68    \\
+      1 to 2 & 10   &  0&01 \\
+      1 to 2 & 100  &  0&43 \\
+      1 to 2 & 500  &  7&36 \\
+      1 to 2 & 1000 & 27&47 \\
+      \midrule
+      1 to 3 & 10   &  0&01 \\
+      1 to 3 & 100  &  0&24 \\
+      1 to 3 & 500  &  3&99 \\
+      1 to 3 & 1000 & 15&28 \\
       \bottomrule
-    \end{tabular}
+    \end{tabular*}
   \end{minipage}
-  \begin{minipage}{0.49\linewidth}
+  \begin{minipage}{0.60\linewidth}
     \includegraphics[width=\linewidth]{figures/benchmarks.eps}
@@
 Throughout we have asserted that user's have policies and that there is a need for policy enforcement tools.
-Corporate mobile security \ac{BYOD policies} have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
+Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
 In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
@@
   We believe our approximation is valid, however, as users will not know the precise reason an app uses a permission, and may decline to install it anyway.
-  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND_SMS} permission to run).
+  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND\_SMS} permission to run).
 \end{itemize}

### 2015-09-20 figures/benchmarks.R

@@ -1,0 +1,17 @@
+library('ggplot2')
+library('ggthemes')
+
+setEPS()
+postscript("benchmarks.eps", width=6, height=4)
+
+data <- read.csv('benchmarks.csv', header=TRUE)
+qplot(Principals, Check, data=data, group=Delegations, color=Delegations, geom="line") +
+  scale_x_log10(breaks=c(10,100,500,1000)) +
+  scale_y_log10(breaks=c(1,10,100)) +
+  ylab("Check time (s)") +
+  geom_point(size=1.1) +
+  geom_rug() +
+  theme_tufte(ticks=F, base_family="Helvetica")
+
+dev.off()
+

### 2015-09-21 paper.tex
updates and grammar

@@
 \begin{abstract}
+  It can be difficult to find appropriate mobile apps.
+  At work, employers rely on employees enforcing company policies about apps at the workplace correctly.
   Users must judge apps by the information shown to them by the store.
-  Employers rely on employees enforcing company policies correctly on devices at their workplace.
-  Some users take time to pick apps with care whilst others do not.
+  Whilst some users take time to pick with care others do not.
   Users feel frustrated when they realise what data the apps have access to.
@@
   AppPAL policies use statements from third parties and delegation relationships to give us a rigorous and flexible framework for enforcing and comparing app policies; and the trust relationships surrounding them.
+  Finally we show that AppPAL can be used to model user's privacy policies but that most users do not seem to enforce them when picking apps for their phone: a demonstration of the privacy paradox.
 \end{abstract}
@@
 Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
-Employees are increasingly using their own phones for work (a system known as \ac{byod}).
+Employees are increasingly using their own phones for work (a scheme called \ac{byod}).
 An employer may wish to restrict which apps their employees can use.
-The IT department may set a policy to prevent information leaks.
+The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
 Some users worry apps will misuse their personal data;
   such a user avoids apps which can access their location, or address book.
-They may apply their own personal security policy when downloading and running apps.
+They may apply their own personal security policies when downloading and running apps.
 
@@
 We believe this can be improved.
-An alternative would be to write the policy down and make the computer enforce it.
-To implement this we use a logic of authorization.
+An alternative is be to write the policy down and make the computer enforce it.
+To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
 The policy is written in the logic and enforced by checking the policy is satisfied.
@@
 We present AppPAL, an authorization logic for reasoning about apps.
-The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
+The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool}.
+and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
@@
 \end{lstlisting}
-We can delegate policy specification to third parties or roles, and assign principals to those rolls:
+We can delegate policy specification to third parties or roles, and assign principals (entities which speak) to roles:
 \begin{lstlisting}
@@
     Implemented the AppPAL language as a library on Android and the JVM.
-    We show how the language can describe properties of Android apps and Android security policies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).
+    We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
   \item
@@
   \begin{minipage}{0.49\linewidth}
-    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize,keywordstyle=\ttfamily\scriptsize]
-      Assertion := <E> says <Fact>
-                     (if (<Fact>,)+)?
-                     (where <Constraint>)?.
-      Fact := <E> (isRunable | ...)
-            | <E> can-say (0 | inf) <Fact>
-            | <E> can-act-as <E>
-      E := Variable | `constant'
-    \end{lstlisting}
+  \newcommand{\nonterminal}[1]{$\langle$#1$\rangle$}
+  \newcommand{\terminal}[1]{\textbf{#1}}
+  \begin{tabular}{r c l}
+    \footnotesize
+    \nonterminal{Assertion} & $\coloneqq$ & \nonterminal{E} \terminal{says} \nonterminal{Fact} \\
+                            &             & \hspace{1em}(\terminal{if} (\nonterminal{Fact}\terminal{,})+)?\terminal{.} \\
+                            &             & \hspace{1em}(\terminal{where} \nonterminal{Constraint})? \\
+    \nonterminal{Fact}      & $\coloneqq$ & \nonterminal{E} (\terminal{isRunable} $\vert$ $\ldots$) \\
+                            & $\vert$     & \nonterminal{E} \terminal{can-say} (\terminal{0} $\vert$ \terminal{inf}) \nonterminal{Fact} \\
+                            & $\vert$     & \nonterminal{E} \terminal{can-act-as} \nonterminal{E} \\
+    \nonterminal{E}         & $\coloneqq$ & \terminal{Variable} $\vert$ \terminal{`constant'}
+  \end{tabular}
   \end{minipage}
@@
 Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
-Downloading and installing an app takes roughly 30s on an Android phone.
-If checking a policy delays this even further user's may become annoyed and disable AppPAL.
+Downloading and installing an app takes about 30 seconds on a typical Android phone over wifi.
+If checking a policy delays this even further a user may become annoyed and disable AppPAL.
 
-The policy checking search procedure is at its slowest when having to delegate repeatedly;
+The policy checking procedure is at its slowest when having to delegate repeatedly;
   with the depth of the delegation tree being the biggest factor for slowing the search.
-Synthetic benchmarks were created to check that the search procedure performed acceptably.
+Synthetic benchmarks were created to check that the checking procedure performed acceptably.
 Each benchmark consisted of a chain of delegations.
-The \emph{1 to 1} benchmark consists of a single long chain of delegation.
-The \emph{1 to 2} benchmark consists of a binary tree of principals delegating to each other; where each principal delegated to 2 others.
-Similarly the \emph{1 to 3} benchmark consists of a tree where each principal delegated to 3 others.
-These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by delegating even more.
-% A snippet of the \emph{1 to 1} and \emph{1 to 3} policies is shown in \autoref{fig:policy-snippet}.
+The \emph{1 to 1} benchmark consists of a repeated delegation between all of the principals.
+In the \emph{1 to 2} benchmark each principal delegated to 2 others and in the \emph{1 to 3} benchmark each principal delegated to 3 others.
+These benchmarks are reasonable as they model the slowest kinds of policies to evaluate---though worse ones could be designed by delegating even more.
+
 For each benchmark we controlled the number of principals in the policy file:
   as the number of principals increased so did the size of the policy.
-
-On a \emph{Nexus 4} checking times are measured in seconds when there are hundreds of delegations, and in minutes when there are thousands (\autoref{fig:benchmarks}).
+The results are shown in \autoref{fig:benchmarks}.
 We have only used a few delegations per decision when describing hypothetical user policies.
-Since the benchmarks show that long chains of delegation can be used, we believe the performance is acceptable.
+We believe the policy checking performance of AppPAL is acceptable as unless a policy consists of hundreds of delegating principals the overhead of checking an AppPAL policy is negligable.
 

### 2015-09-23 paper.tex
updates

@@
 \documentclass[]{llncs}
+
+% Not using NatBib but tired of fixing issues with citations being inserted in that format
+\newcommand{\citep}[1]{\cite{#1}}
 
@@
 \begin{abstract}
-  It can be difficult to find appropriate mobile apps.
+  It can be difficult to find mobile apps that respect your security and privacy.
   At work, employers rely on employees enforcing company policies about apps at the workplace correctly.
   Users must judge apps by the information shown to them by the store.
-  Whilst some users take time to pick with care others do not.
-  Users feel frustrated when they realise what data the apps have access to.
-  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
-  We present AppPAL: a policy language to enforce policies about apps.
-  AppPAL policies use statements from third parties and delegation relationships to give us a rigorous and flexible framework for enforcing and comparing app policies; and the trust relationships surrounding them.
-  Finally we show that AppPAL can be used to model user's privacy policies but that most users do not seem to enforce them when picking apps for their phone: a demonstration of the privacy paradox.
+  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and
+  most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
+  We present AppPAL: a policy language for Android to describe app installation  policies.
+  AppPAL goes beyond existing policy enforcement tools, such as Kirin~\cite{Enck:2008vn}, with delegation relationships and the use of static analysis tools.
+  We use AppPAL to explore what kinds of policies user's use when installing apps, finding a novel example of the privacy paradox when comparing users privacy views to their actual behaviour.
 \end{abstract}
@@
 Finding the right apps can be tricky.
-Users need to discover which apps are well written, which are not going to abuse their data
-  and to find the apps which suit how they want to use their device.
+Users need to discover which are not going to abuse their data.
 This can be difficult as it isn't obvious how apps use the data each has access to.
+Consider a user attempting to buy a flashlight app.
+By searching the Play store the user is presented with a long list of apps.
+Clicking through each one they can find the permissions each requests but not the reasons why each was needed.
+They can see review scores from users but not experts who can run tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
+If they want to use the app at work will it break their employers rules for mobile usage?
 
@@
 Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
-Ultimately the decision which apps to use and which permissions to grant must be made by the user.
+Ultimately the decision which apps to use and which permissions to grant must be made by the device user.
 
-Not all apps are suitable.
+Some apps are highly undesirable.
 Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
-Employees are increasingly using their own phones for work (a scheme called \ac{byod}).
-An employer may wish to restrict which apps their employees can use.
+Employees are increasingly using their own phones for work.
+An employer may restrict which apps their employees can use.
 The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
-Some users worry apps will misuse their personal data;
-  such a user avoids apps which can access their location, or address book.
+Some users worry apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
+Such a user avoids apps which can access their location, or address book.
 They may apply their own personal security policies when downloading and running apps.
@@
 This is error-prone.
-We believe this can be improved.
-An alternative is be to write the policy down and make the computer enforce it.
+An alternative is to write the policy down and make the computer enforce it.
 To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
@@
 
-We present AppPAL, an authorization logic for reasoning about apps.
-The language is an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool}.
-and predicates that allow us to decide which apps to run or install.
+We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool} and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
 The implementation allows us to enforce the policies on a device.
-We can express trust relationships amongst these parties; use constraints to do additional checks.
-This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.
+We can express trust relationships amongst these parties; use constraints to do additional checks, such as using static analysis results.
+This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.
 
-Using AppPAL we write policies for work and home, and decide which policy to use using a user's location, or the time of day:
+A user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
+Using AppPAL we can write for work and home, and decide which policy to enforce using a user's location, or the time of day:
 \begin{lstlisting}
@@
   if "home-policy" isMetBy(App)
-  where At("work") = false.
+  where at("work") = false.
 
@@
   if "work-policy" isMetBy(App)
-  where BeforeHourOfDay("17") = true.
+  where beforeHourOfDay("17") = true.
 \end{lstlisting}
-We can delegate policy specification to third parties or roles, and assign principals (entities which speak) to roles:
+We can delegate policy specification to third parties or roles, and assign principals to roles:
 \begin{lstlisting}
-"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
+"alice" says "it-department" can-say "work-policy" isMetBy(App).
 "alice" says "alice" can-act-as "it-department".
@@
 
-% Using AppPAL we can say an app is installable:
-% \begin{lstlisting}
-% "Alice" says "com.rovio.angrybirds" isInstallable.
-% \end{lstlisting}
-% We can delegate decisions to others:
-% \begin{lstlisting}
-% "Alice" says "ITDepartment" can-say 0 App isInstallable.
-% \end{lstlisting}
-% We can run analysis tools, and solve constraints:
-% \begin{lstlisting}
-% "Claire" says App isMalware
-%   if App isAnApp
-%   where
-%     virusScanner(App) = true.
-% \end{lstlisting}
-% Constraints allow up to express facts that are true at some times but not others:
-% \begin{lstlisting}
-% "Emma" says "com.facebook.katana" isInstallable
-%   where timeOfDay() > 1700.
-% \end{lstlisting}
+There has been a great amount of work on developing app analysis tools for Android.
+Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
+TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
+Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
+ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
+Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
+Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
+Many others exist checking and certifying other aspects of app behaviour.
+
+AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owner's are trying to enforce.
+For example a user might not want to install apps with SSL errors (which Mallodroid can check for~\citep{Fahl:2012dj}) or malicious apps as determined by their anti-virus.
+Using AppPAL we can combine tools for checking apps to implement the user's policies.
+\begin{lstlisting}
+"alice" says App isInstallable
+  if App isAnApp
+  where mallodroidCheck(App) = true,
+        avCheck(App) = true.
+\end{lstlisting}
 
@@
   \item
-    Implemented the AppPAL language as a library on Android and the JVM.
+    Described our implementation of the AppPAL language as a library for Android and the JVM.
     We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
@@
 
-An employee \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
-Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
+An employee \emph{Alice} works for her employer \emph{Emma}.
+Emma allows Alice to use her personal phone as a work phone but has some specific concerns.
 \begin{itemize}
@@
 This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
-Emma's policy could be implemented using existing tools.
-\emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}).
-AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and provide a rigorous means for deciding whether an app meets a policy.
+Alternately Emma's policy could be partially enforced using existing tools.
+\emph{Google's Device Policy for Android}~\cite{GoogleAppsDeviceP:tz} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}.
+
+AppPAL builds on these tools, but make the trust and delegation relationships explicit and provide a rigorous and extensible means for deciding whether an app meets a policy.
 Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
@@
 
-Our research looks at the problem of picking software.
-Given there are some apps you want to install and run and others you do not:
-  how can you express your preferences in such a way that they can be enforced automatically?
-How can we translate policy documents from natural language into a machine checkable form?
-How can we show the trust relationships used to make these decisions clearly and precisely?
+% Our research looks at the problem of picking software.
+% Given there are some apps you want to install and run and others you do not:
+%   how can you express your preferences in such a way that they can be enforced automatically?
+% How can we translate policy documents from natural language into a machine checkable form?
+% How can we show the trust relationships used to make these decisions clearly and precisely?
 
@@
 
-We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
+We propose a change to the mobile ecosystem, shown in \autoref{fig:ecosystem}.
 People have policies which are enforced by AppPAL on their devices.
@@
 
-"emma" says "google-play" can-say 0
+"emma" says "google-play" can-say
   App isReputable. @\label{lst:corporate_4}@
@@
   with \code{"google-play"} specified as the decider of what is buyable or not.
-This time Google is not allowed to delegate the decision further (\code{can-say 0}).
+This time Google is not allowed to delegate the decision further (\code{can-say}).
 In other words Google is not allowed to specify Amazon as a supplier of apps as well.
@@
 The parser is implemented using ANTLR4.
-The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.
+The structure of an AppPAL is inherited from SecPAL~\cite{Becker:2006vh} can be seen in \autoref{fig:assertion}.
 
@@
     \nonterminal{Fact}      & $\coloneqq$ & \nonterminal{E} (\terminal{isRunable} $\vert$ $\ldots$) \\
-                            & $\vert$     & \nonterminal{E} \terminal{can-say} (\terminal{0} $\vert$ \terminal{inf}) \nonterminal{Fact} \\
+                            & $\vert$     & \nonterminal{E} \terminal{can-say} \terminal{inf}? \nonterminal{Fact} \\
                             & $\vert$     & \nonterminal{E} \terminal{can-act-as} \nonterminal{E} \\
@@
 
-In Becker~\etal's work~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
-With AppPAL we make explicit our predicates and how they relate to Android.
+In SecPAL the precise nature of predicates and constraints is left open.
+In instantiating SecPAL, AppPAL makes the predicates and constraints explicit.
 AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
-Additional predicates can be created in the policy files, but adding or modifying constraints required a code change\footnote{The Android version of the \texttt{hasPermission} constraint, for example, uses the Android package manager to determine what permissions an app requests, but the Java version uses the Android platform tools.
-This requires a small code change.}
+Additional predicates can be created in the policy files, however constraints are implemented individually.
+For example on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.
 
 \begin{table}
+  \footnotesize
+  \newcommand{\style}[1]{{\footnotesize #1}}
   \begin{tabulary}{\linewidth}{l L}
   \toprule
-  Name & Description \\
+  \style{Name}                                             & \style{Description}                                                                                                                        \\
   \midrule
-  App \texttt{isRunnable} & Says an app can be run. \\
-  App \texttt{isInstallable} & Says an app can be installed. \\
-  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
-  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
-  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
-  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
-  Tool\texttt{Check(}App, Property) &  General form of a constraint used to run a static analysis tool over an app. \\
+  \style{App \texttt{isRunnable}}                          & \style{Says an app can be run.}                                                                                                            \\
+  \style{App \texttt{isInstallable}                      } & \style{Says an app can be installed.
+} \\
+  \style{App \texttt{isAnApp}                            } & \style{Tells AppPAL that an app exists.
+} \\
+  \style{Policy \texttt{isMetBy(}App\texttt{)}           } & \style{Used to split policies into smaller components.}                                                                                    \\
+  \style{\texttt{hasPermission(}App, Permission\texttt{)}} & \style{Constraint to check if an app has a permission.
+} \\
+  \style{\texttt{beforeHourOfDay(}time\texttt{)}         } & \style{Constraint used to check the time.
+} \\
+  \style{Tool\texttt{Check(}App, Property)               } & \style{Constraint to run an analysis tool on an app.
+}              \\
   \bottomrule
 \end{tabulary}
-\caption{Typical AppPAL predicates and constraints}
+\caption{AppPAL predicates and constraints.}
 \label{tab:predicates}
 \end{table}
-
-% \begin{description}
-%   \item[\texttt{App isRunnable}]
-%     Used to indicate that an App meets the install policy for the device.
-%     Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
-%
-%   \item[\texttt{App isAnApp}]
-%     Expresses a simple typing relations.
-%     AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
-%     These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
-%     This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
-%
-%   \item[\texttt{Policy isMetBy(App)}]
-%     Policies let you split app behaviour into sub-policies.
-%     For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
-%     Splitting policies allows greater control of how each one is checked.
-%     We can delegate checking a policy to an expert using the \code{can-say} statement.
-%     We can specify more detailed checks using the \code{can-act-as} statements.
-%
-%   \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
-%     A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
-% \end{description}
 
@@
 Some apps may want to access this data, but may not leak it.
-We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
+We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Fritz:2013vi}).
 Our policy now becomes:
@@
   where
-    taintDroidCheck(App, "Location", "Internet") = false.
+    flowDroidCheck(App, "Location", "Internet") = false.
 \end{lstlisting}
@@
 
-AppPAL has the same policy inference rules as SecPAL.
+AppPAL has the same policy checking rules as SecPAL~\cite{Becker:2006vh}.
+AppPAL uses an assertion context of known facts and rules
 We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
@@
 
-Like Becker~\etal~we make use of an assertion context to store known statements.
-We also cache intermediate results, and ground facts to avoid re-computation.
+Like Becker~\etal~we make use of an assertion context to store known statements and rules, as well as facts that are deduced during checking.
 On a mobile device memory is at a premium.
-We would like to keep the context as small as possible.
+We would like to keep the assertion context as small as possible.
 For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
@@
   we can fetch the assertion that the subject is an app based on the app in question.
-% Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
-%   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.
-
-% In SecPAL there is no type.
-% All entities are either constants (implemented as strings) or variables.
-% There is no differentiation made between a \emph{voiced} constant (who says assertions),
-% and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
-% In AppPAL the subjects are usually apps or policies.
-% Apps do not utter facts;
-%   so there is no need to check them as potential speakers of \emph{can-say} statements.
-% Similarly if a constant never appears as a subject of a statement, as most speakers do not,
-%   there is no need to check it as a possible subject for a \emph{can-act-as} statement.
-% We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
-% This means we need to search less constants when attempting to replace a variable, speeding evaluation.
 
@@
 
-% \subsection{Policy Examples}
-% \label{ssec:idioms}
-%
-% When writing policies some schemes often come up.
-% We give examples of policies and queries and show how they can be implemented in AppPAL.
-%
-% \begin{description}
-%   \item[Policies for home and work]
-%     In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
-%     Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
-%     To implement this we add rules that use constraints to decide whether an app should be run.
-%     Alice already has the Emma's work policy as the default.
-%     Now she just needs to add an exception to the rule.
-%     \begin{lstlisting}
-% "alice" says App isRunnable
-%   if "home-policy" isMetBy(App)
-%   where TimeOfDay() > 17.00.
-%
-% "alice" says App isRunnable
-%   if "home-policy" isMetBy(App)
-%   where At("work") = false.
-%     \end{lstlisting}
-%
-%   \item[App white-listing by an IT department]
-%     Inside Emma's company employees in the IT department may white-list certain apps as runnable.
-%     Emma delegates to them to decide if an app is runnable.
-%     \begin{lstlisting}
-% "emma" says "it-department" can-say 0 App isRunnable.
-%     \end{lstlisting}
-%     Charlie and Diveena work in the IT department.
-%     When they have checked an app they say it is runnable.
-%     \begin{lstlisting}
-% "emma" says "charlie" can-act-as "it-department".
-% "emma" says "diveena" can-act-as "it-department".
-%
-% "charlie" says "com.facebook.katana" isRunnable.
-% "diveena" says "org.thoughtcrime.securesms" isRunnable.
-%     \end{lstlisting}
-%
-%   \item[Overpriviliged applications]
-%     Alice is particularly worried about apps stealing her data.
-%     She knows that certain apps need certain permissions.
-%     For example a photography app needs access to the camera.
-%     She checks each app's permissions carefully for things that seem unusual.
-%
-%     Alice's policy can be implemented by checking the permissions of the app, using constraints.
-%     \begin{lstlisting}
-% "alice" says App isRunnable
-%    if "permissions-policy" isMetBy(App).
-% "alice" says "permissions-policy" isMetBy(App)
-%    if App isAnApp
-%    where
-%       category(App, "Photography"),
-%       hasPermission(App, "LOCATION") = false,
-%       hasPermission(App, "CAMERA") = true.
-%     \end{lstlisting}
-%
-%     Discovering the \emph{normal} permissions for an app can be done by mining large stores of apps, or apps a user has previously installed.
-%     Alternately tools like Stowaway~\cite{Felt:2011kj} could be integrated to check apps.
-
-%   \item[Digital Evidence]
-%     In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
-%     Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
-%     For example that no audio can be recorded without the users consent~\cite{Seghir:2014uq}.
-%     Using this tool someone generates a proof certificate that this property is met.
-%     This might be the developer of the app, a third party certification service, or someone else;
-%       when we check the proof we will discover if it is valid or not.
-%     \begin{lstlisting}
-% "alice" says Anyone can-say 0
-%   "recording-consent" shownIsMetBy(Evidence, App).
-%
-% "skb" says "recording-consent"
-%   shownIsMetBy("evicheck://evidence123", App).
-%
-% "alice" says "recording-consent" isMetBy(App)
-%   if "recording-consent" shownIsMetBy(Evidence, App)
-%   where
-%     evicheckCheckEvidence(Evidence, App) = true.
-%     \end{lstlisting}
-%     In the example here Alice allows anyone to present her with evidence for her policy.
-%     A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
-% \end{description}
-
 \subsection{Benchmarks}
@@
 
-We envisage AppPAL running on a mobile phone checking apps when they are installed.
+We envisage AppPAL running on a mobile phone, checking apps when they are installed.
 Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
@@
 The policy checking procedure is at its slowest when having to delegate repeatedly;
-  with the depth of the delegation tree being the biggest factor for slowing the search.
+   the depth of the delegation tree is the biggest factor for slowing the search.
 Synthetic benchmarks were created to check that the checking procedure performed acceptably.
@@
 
-% \begin{figure}\centering
-%   \begin{minipage}{0.49\linewidth}
-%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
-%     '1' says '2' can-say App isInstallable.
-%     '2' says '3' can-say App isInstallable.
-%     '3' says '4' can-say App isInstallable.
-%     '4' says '5' can-say App isInstallable.
-%     '5' says '6' can-say App isInstallable.
-%     '6' says '7' can-say App isInstallable.
-%     \end{lstlisting}
-%   \end{minipage}
-%   \begin{minipage}{0.49\linewidth}
-%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
-%     '1' says '2' can-say App isInstallable.
-%     '1' says '3' can-say App isInstallable.
-%     '2' says '4' can-say App isInstallable.
-%     '2' says '5' can-say App isInstallable.
-%     '3' says '6' can-say App isInstallable.
-%     '3' says '7' can-say App isInstallable.
-%     \end{lstlisting}
-%   \end{minipage}
-%   \caption{Snippets of the \emph{1 to 1} and \emph{1 to 3} policies used for benchmarking.}
-%   \label{fig:policy-snippet}
-% \end{figure}
-
 \begin{figure}
+  \centering
   \begin{minipage}{0.40\linewidth}
@@
   \end{minipage}
-  \begin{minipage}{0.60\linewidth}
+  \begin{minipage}{0.59\linewidth}
     \includegraphics[width=\linewidth]{figures/benchmarks.eps}
@@
 
-\section{Demonstration}
+\section{Measuring Policy Compliance}
 \label{sec:demonstation}
@@
 
-Malware is available for Android.
-Like other vendors, McAfee classify malware into several categories.
+McAfee classify malware into several categories.
 The \emph{malicious} and \emph{trojan} categories describe traditional malware.
@@
 If a user is enforcing a privacy policy we might also expect them to install less malware.
-We can check this by using AppPAL policies to find the malware and again measure the number of malwares each user had installed.
+We can check this by using AppPAL policies to measure the number of malwares each user had installed.
 
@@
 
-Using AppPAL we have written short policies describing user behaviour and used these to identify the users following them to varying degrees.
-Some limitations of our results include:
+Some limitations of this investigation include:
 \begin{itemize}
@@
 
-Authorization logics have been successfuly used to enforce policies in several other domains.
+Authorization logics have been successfully used to enforce policies in several other domains.
 The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general, if undecidable.
 Logics that followed like KeyNote~\cite{Blaze:1999fa} and SPKI/SDSI~\cite{Ellison:1999ui} looked at public key infrastructure.
-The RT-languages~\cite{Li:2002if,Li:2003ix,Li:2003to} were designed for credential management.
-Cassandra~\cite{Becker:2004fi} was used to model trust relationships in the british national health service.
+The RT-languages~\cite{Li:2002if} were designed for credential management.
+Cassandra~\cite{Becker:2004fi} was used to model trust relationships in the British national health service.
 
@@
 It was ported to Android~\cite{Smalley:2013vl} and is used in the implementation of the permissions system.
+SELinux describes the capabilities (in terms of system calls and file access) of processes, it cannot describe app installation policies or delegation relationships.
 Google also offer the \emph{Device Policy for Android} app.
@@
 The SecPAL language is designed for access control in distributed systems.
-We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\cite{Hallett:2014un}.
-It has already been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
-Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and ultimately spawned the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
-Gruevich and Neeman showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
-DKAL also contains more modalities than \emph{says}, which lets policies describe actions principals carry out rather than just their oppinions.
-For example in AppPAL a user might \emph{say} an app is installable if they would install it (\code{"user" says App isInstallable})
+We picked SecPAL as the basis for AppPAL because it is readable, extensible, and is a good for for the mobile ecosystem setting~\citep{Hallett:2014un}.
+It has also been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
+Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+DKAL contains more modalities than \emph{says}, which lets policies describe actions principals carry out rather than just their opinions.
+For example in AppPAL a user might \emph{say} an app is installable if they would install it (\code{"user" says App isInstallable}).
 In DKAL they can describe the conditions that would force them to install it (\code{"user" installs App}).
-The distinction is that in AppPAL whilst the user thinks the app could be installed we do not know for sure whether the user has installed it.
-With DKAL we can guarantee that the action was completed.
+With DKAL we can guarantee that the action was completed, whereas in AppPAL we do not know if the user actually installed a particular app.
+We chose to use SecPAL as the basis for AppPAL as we did not need the extra features DKAL added to express app installation policies.
+If, in future work, we need additional modalities AppPAL could be extended to support additional DKAL features as SecPAL has been shown to be a subset of the DKAL language.
 
-Kirin~\cite{Enck:2009ko} is a policy language and tool for enforcing app installation policies preventing malware.
-Policy authors could specify combinations of permissions that should not appear together.
-For example an author might wish to stop malware sending premium rate text messages.
-To might implement this by restricting an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions (\autoref{fig:kirin}).
-Using this approach they found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
-\begin{figure}
+Kirin~\cite{Enck:2009ko} is a policy language and tool for enforcing app installation policies to prevent malware.
+Policy authors can specify combinations of permissions and broadcast events that should not appear together.
+For example to stop malware sending premium rate text messages, we prevent
+ an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions.
 \begin{lstlisting}
@@
 \end{lstlisting}
-% \begin{lstlisting}
-% "user" says "no-write-send-sms" isMetBy(App)
-%   where hasPermission(App, "SEND_SMS") = false.
-% "user" says "no-write-send-sms" isMetBy(App)
-%   where hasPermission(App, "WRITE_SMS") = false.
-% \end{lstlisting}
-\caption{Kirin policy stopping apps monetized by premium rate text messages.}
-\label{fig:kirin}
-\end{figure}
+By analyzing apps which broke their policies Enck~\etal~found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
 
@@
 The behaviours and permissions \ac{pup} displays aren't necessarily malicious.
-One user may consider apps which need in-app-purchases to play malware, but another may enjoy them.
+One user may not want apps which need in-app-purchases to play, but another may enjoy them.
 With Kirin we are restricted to permitting or allowing apps.
-AppPAL is more general: we can describe more scenarios than just permit or allow, and use more app information than just permissions.
+AppPAL can describe more scenarios than just permit or allow, and use more app information than just permissions, such as constraints and static analysis results.
 By allowing delegation relationships we can understand the provenance and trust relationships in these rules.
-With AppPAL we can incorporate static analysis results and take into account the user's location and other constraints.
-
-There has been a great amount of work on developing app analysis tools for Android.
-Tools such as Stowaway~\cite{Felt:2011kj} detect overprivileged apps.
-TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
-Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
-ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
-Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
-Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
-Many others exist checking and certifying other aspects of app behaviour.
 
@@
 
-We have presented AppPAL: an authorization logic for describing app installation policies.
-AppPAL lets us enforce app installation policies on Android devices.
+We have presented AppPAL: an authorization logic for describing app installation policies primarily designed to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways.
 We have shown how the language can be used to describe an app installation policy;
   and given brief descriptions of how other policies might be described.
+Using app installation data and user privacy policies expressed using AppPAL we showed how the privacy paradox can be seen in the Android ecosystem.
+This motivates the need for tools
 
@@
 One way to integrate AppPAL on Android would be as a \emph{required checker}: a program that checks all apps before installation.
-Google uses this API to check for known malware and jailbreak apps.
+Google uses the required checker API to check for known malware and jailbreak apps.
 We would use AppPAL to check apps meet policies before installation.
 Unfortunately the API is protected and it would require the phone to be rooted.
-Alternatively AppPAL could be integrated as a service to reconfigure app permissions.
-The latest version of Android\footnote{Called \emph{Android M}.} is moving to an iOS like permissions model where permissions can be granted and revoked at any time.
+Alternatively, AppPAL could be integrated as a service to reconfigure app permissions.
+Android Marshmallow has an iOS like permissions model where permissions can be granted and revoked at any time.
 These will be manually configurable by the user through the settings app.
 We can imagine AppPAL working to reconfigure these settings (and set their initial grant or deny states) based on a user's policy, as well as the time of day or the user's location.
-A policy could deny notifications while a user is driving for example.
+A policy could deny notifications while a user is driving, for example, by checking if they are using Android Auto~\cite{AndroidAuto:uw} (an app to interact with a car's center console) or moving along a road at high speed.
 
@@
 For most end-users writing a policy in a formal language is too much work.
-Ad-blocking software works by users subscribing to filter policies written by experts\footnote{EasyList is a popular choice and the default in most ad-blocking software.
-They offer many different policies for specific use-cases however.
-\url{https://easylist.adblockplus.org/en/}}.
+Ad-blocking software works by users subscribing to filter policies written by experts.
+EasyList is a popular choice and they offer many different policies for specific use-cases\footnote{\url{https://easylist.adblockplus.org/en/}}.
 We can imagine a similar scheme working well for app installation policies.
@@
 
-We should also attempt to learn policies from existing users behavior.
+We might attempt to learn policies from existing users behavior.
 Given app usage data, from a project like Carat~\cite{Oliner:2013ht}, we could identify security conscious users.
@@
 The work provides new ways for users to enforce their own rules about how apps should behave.
-Users policies can be enforced more reliably, and with less interaction;
-  making apps more pleasant for everyone.
+Users policies can be enforced more reliably, and with less interaction.
 

### 2015-09-29 paper.tex
updates

@@
   We present AppPAL: a policy language for Android to describe app installation  policies.
-  AppPAL goes beyond existing policy enforcement tools, such as Kirin~\cite{Enck:2008vn}, with delegation relationships and the use of static analysis tools.
-  We use AppPAL to explore what kinds of policies user's use when installing apps, finding a novel example of the privacy paradox when comparing users privacy views to their actual behaviour.
+  AppPAL goes beyond existing policy enforcement tools, such as Kirin~\cite{Enck:2008vn}, adding delegation relationships and the connecting to static analysis tools.
+  We use AppPAL to explore what kinds of policies users use when installing apps, demonstrating the privacy paradox when comparing user's privacy views to their actual behaviour.
 \end{abstract}
@@
 Using AppPAL we can write for work and home, and decide which policy to enforce using a user's location, or the time of day:
+
+\noindent
+\begin{minipage}{\linewidth}
+  \begin{minipage}{0.5\linewidth}
     \begin{lstlisting}
@@
   where at("work") = false.
-
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.5\linewidth}
+    \begin{lstlisting}
 "alice" says App isRunnable
@@
     \end{lstlisting}
+  \end{minipage}
+\end{minipage}
+
 We can delegate policy specification to third parties or roles, and assign principals to roles:
@@
 
-\noindent
-In this paper we have:
-\begin{itemize}
-  \item
-    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
-    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
-  \item
-    Described our implementation of the AppPAL language as a library for Android and the JVM.
-    We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
-  \item
-    Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
-\end{itemize}
+% \noindent
+% In this paper we have:
+% \begin{itemize}
+%   \item
+%     Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
+%     and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
+%   \item
+%     Described our implementation of the AppPAL language as a library for Android and the JVM.
+%     We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
+%   \item
+%     Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
+% \end{itemize}
 
@@
 
-% AppPAL also helps attribute blame when things go wrong.
-% By modelling the trust relationships in systems we can work out precisely where mistakes were made.
-% A recent example demonstrating this is \textsc{CVE-2015-2077}.
-% Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
-% Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
-% Unfortunately they had configured it to ignore Superfish.
-% The malware was an ad framework designed to show users products similar to those they viewed on the web.
-% Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.
-%
-% From a user's perspective there was a delegation of trust to Lenovo to detect malware.
-% Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
-% A user might wonder where the breach of trust occurred and who is to blame?
-% Is it the \ac{av} failing to spot the malware?
-% Has someone else configured the software incorrectly?
-% How should they fix the problem?
-%
-% If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
-% \begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
-% "user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
-% "lenovo" says "mcafee" can-say inf File isSafe.
-% "lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
-% \end{lstlisting}
-% The fault lies with \autoref{lst:lenovo_unsafe}.
-% Lenovo has caused the problem (they \emph{said} it was safe).
-% The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.
-
 \subsection{Policy checking}
@@
 \begin{figure}\centering
-  \includegraphics[width=0.5\linewidth]{./figures/c_v_pup.eps}
-  \caption{Compliance with the advanced policy and non-\ac{pup} policy.}
+  \subfigure[Advanced and non-\ac{pup} policies.]{%
+    \includegraphics[width=0.48\linewidth]{./figures/advanced-v-pup.eps}
+    \label{fig:versus_pup}}
+    \subfigure[Advanced and Mallodroid policies.]{%
+    \includegraphics[width=0.48\linewidth]{./figures/advanced-v-mallodroid_m.eps}
+    \label{fig:versus_mallodroid}}
+  \caption{Compliance with the advanced policy and the non-\ac{pup} and SSL policies.}
   \label{fig:versus}
@@
 though a user is 3 times more likely to have a \ac{pup} installed than malware.
-Users who were complying more than half the time with the conservative or advanced policies complied with the malware or \ac{pup} policies fully (\autoref{fig:versus}).
+Users who were complying more than half the time with the conservative or advanced policies complied with the malware or \ac{pup} policies fully (Figure~\autoref{fig:versus_pup}).
 This suggests that policy enforcement is worthwhile: users who can enforce policies about their apps experience less malware.
+
+The \emph{MalloDroid} tool~\citep{Fahl:2012dj} can scan apps for SSL misconfigurations.
+SSL misconfigurations are dangerous as they can undermine any privacy guarantees that SSL/TLS usage gives.
+We set up AppPAL to use Mallodroid results as a constraint and measured the percentage of apps each user had installed that did not have issues or suspected issues when scanned with Mallodroid.
+Users who were complying with the advanced policy were no better at avoiding apps with SSL errors than any other users (Figure~\autoref{fig:versus_mallodroid}).
+
 %This is significant (P-value $< 0.05$) and suggests that users who pick their apps carefully are less likely to experience malware.

### 2015-10-01 paper.tex
updates

@@
   It can be difficult to find mobile apps that respect your security and privacy.
-  At work, employers rely on employees enforcing company policies about apps at the workplace correctly.
+  Employers rely on employees enforcing company policies about apps at the workplace correctly.
   Users must judge apps by the information shown to them by the store.
@@
   We present AppPAL: a policy language for Android to describe app installation  policies.
-  AppPAL goes beyond existing policy enforcement tools, such as Kirin~\cite{Enck:2008vn}, adding delegation relationships and the connecting to static analysis tools.
-  We use AppPAL to explore what kinds of policies users use when installing apps, demonstrating the privacy paradox when comparing user's privacy views to their actual behaviour.
+  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2008vn}, adding delegation relationships and the connecting to static analysis tools.
+  We use AppPAL to explore what kinds of policies are used when installing apps, demonstrating the privacy paradox when comparing user's privacy views to their actual behaviour.
 \end{abstract}
@@
 
-We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool} and predicates that allow us to decide which apps to run or install.
+We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool.} and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
@@
 \end{lstlisting}
-We can write policies specifying which permissions an app must or must not have by its app store categorization:
+We can write policies specifying which permissions an app must or must not have by its app store categorization.
+For example it would be okay allowing a photography app access to the camera, but not to the location data if the user doesn't want their photos geotagged.
 \begin{lstlisting}
@@
 Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
+Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
+MalloDroid~\citep{Fahl:2012dj} looks for apps configured to use SSL incorrectly (for instance by not verifying hostnames or certificates).
 Many others exist checking and certifying other aspects of app behaviour.
@@
 AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owner's are trying to enforce.
-For example a user might not want to install apps with SSL errors (which Mallodroid can check for~\citep{Fahl:2012dj}) or malicious apps as determined by their anti-virus.
-Using AppPAL we can combine tools for checking apps to implement the user's policies.
+This avoids creating tools with hard-coded fixed policies
+For example a store might not want to sell apps with SSL errors or malicious apps as determined by their anti-virus.
+Using AppPAL we can combine tools for checking apps to implement the store's policies.
 \begin{lstlisting}
-"alice" says App isInstallable
+"play-store" says App isInstallable
   if App isAnApp
   where mallodroidCheck(App) = true,
-        avCheck(App) = true.
+         mcafeeAVCheck(App) = true.
 \end{lstlisting}
@@
 
-An employee \emph{Alice} works for her employer \emph{Emma}.
+An employee \emph{Alice} works for \emph{Emma}.
 Emma allows Alice to use her personal phone as a work phone but has some specific concerns.
@@
   \item Alice shouldn't run any apps that can track her movements.
-    The testing labs are at a secret location and it mustn't be leaked.
+    Alice's workplace is at a secret location and it mustn't be leaked.
   \item Apps should come from a reputable source, such as the Google Play Store.
@@
 She might even sign a document promising never to break the rules within the policy.
-This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
+This is error-prone---what if she makes a mistake or misses an app that breaks her policy
 Alternately Emma's policy could be partially enforced using existing tools.
@@
 
-AppPAL builds on these tools, but make the trust and delegation relationships explicit and provide a rigorous and extensible means for deciding whether an app meets a policy.
-Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
-These could be used to used to ensure no location data is ever obtained.
-Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are never leaked to the web.
-Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.
-
-Whilst we could implement Emma's policy using existing tools it is a clumsy solution---they are not flexible.
+We could implement Emma's policy using existing tools (such as an AV checker, and a taint analysis tool like Flowdroid~\cite{Fritz:2013vi}) but it is a clumsy solution---they are not flexible.
+Each has to be configured separately to implement only part of the policy.
 If Emma changes her policy or Alice changes jobs she needs to recheck her apps and then alter or remove the software on her phone to ensure compliance.
@@
 
-We propose a change to the mobile ecosystem, shown in \autoref{fig:ecosystem}.
+We propose a change to the mobile ecosystem shown in \autoref{fig:ecosystem}.
 People have policies which are enforced by AppPAL on their devices.
@@
   "reputable-policy" isMetBy(App) @\label{lst:corporate_3}@
-      if App isReputable.
+      if App isBuyable.
 
 "emma" says "google-play" can-say
-  App isReputable. @\label{lst:corporate_4}@
+  App isBuyable. @\label{lst:corporate_4}@
     \end{lstlisting}
@@
 \end{minipage}
-\caption{AppPAL policy implementing Emma's security requirements}
+\caption{AppPAL policy implementing Emma's security requirements.}
 \label{lst:corporate}
@@
 If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
-In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isReputable};
+In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
   with \code{"google-play"} specified as the decider of what is buyable or not.
-This time Google is not allowed to delegate the decision further (\code{can-say}).
+Google is not allowed to delegate the decision further (\code{can-say} as oposed to \code{can-say inf}).
 In other words Google is not allowed to specify Amazon as a supplier of apps as well.
@@
 
-In this scenario we have imagined Alice wanting to check the apps as she installs them.
 We could also imagine Emma wanting a personalised app store where all apps sold meet her policy.
@@
 Emma might track Alice using a location tracking app.
-Provided the app only talks to Emma, and it uses SSL correctly (which Mallodroid can check for~\cite{Fahl:2012dj}) she is happy to relax the policy.
+Provided the app only talks to Emma, and it uses SSL correctly (using MalloDroid~\cite{Fahl:2012dj}) she is happy to relax the policy.
 
@@
 AppPAL has the same policy checking rules as SecPAL~\cite{Becker:2006vh}.
-AppPAL uses an assertion context of known facts and rules
-We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based checking algorithm, as no DatalogC library exists for Android.
+AppPAL uses an assertion context of known facts and rules, as well as facts deduced while checking.
+We do not use Becker~\etal's DatalogC based checking algorithm, as no DatalogC library exists for Android.
 We have implemented the rules directly in Java.
@@
 
-Like Becker~\etal~we make use of an assertion context to store known statements and rules, as well as facts that are deduced during checking.
 On a mobile device memory is at a premium.
@@
 These simplify the privacy policies identified as by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
+Using AppPAL we could write more precise rules if we could determine why each permission was requested.
+Lin~\etal~used Androguard~\cite{Desnos:ub} as well as manual analysis to determine the precise reasons for each permission~\cite{Sadeh:2014vq}.
 
@@
     \label{fig:versus_pup}}
-    \subfigure[Advanced and Mallodroid policies.]{%
+    \subfigure[Advanced and MalloDroid policies.]{%
     \includegraphics[width=0.48\linewidth]{./figures/advanced-v-mallodroid_m.eps}
@@
 SSL misconfigurations are dangerous as they can undermine any privacy guarantees that SSL/TLS usage gives.
-We set up AppPAL to use Mallodroid results as a constraint and measured the percentage of apps each user had installed that did not have issues or suspected issues when scanned with Mallodroid.
+MalloDroid can scan for several different misconfigurations, and distinguishes cases where the app is definitely misconfigured from those where there is some doubt.
+We set up AppPAL to use MalloDroid results as a constraint and measured the percentage of apps each user had installed that did not have issues or suspected issues when scanned with MalloDroid.
 Users who were complying with the advanced policy were no better at avoiding apps with SSL errors than any other users (Figure~\autoref{fig:versus_mallodroid}).
+This emphasizes that AppPAL policies are more powerful than manually picking apps as they can incorporate checks privacy aware users may want but which they cannot check for without additional tools.
 
@@
 We have presented AppPAL: an authorization logic for describing app installation policies primarily designed to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways.
-We have shown how the language can be used to describe an app installation policy;
-  and given brief descriptions of how other policies might be described.
 Using app installation data and user privacy policies expressed using AppPAL we showed how the privacy paradox can be seen in the Android ecosystem.
-This motivates the need for tools
+We showed how static analysis tools can be integrated into AppPAL to check for properties beyond the scope of the language and to act as a glue between different policies.
 
-Further work is required to tightly integrate AppPAL into Android.
+AppPAL currently runs either an Android app or a Java program.
+Further work is needed to tightly integrate AppPAL into Android.
 One way to integrate AppPAL on Android would be as a \emph{required checker}: a program that checks all apps before installation.
@@
 We would use AppPAL to check apps meet policies before installation.
-Unfortunately the API is protected and it would require the phone to be rooted.
+The API is protected, however, and it would require the phone to have a custom firmware.
+This is undesirable as it would make AppPAL difficult to install for most users, and negate the other security enhancements (such as timely updates and patches) provided by the standard Android system.
 Alternatively, AppPAL could be integrated as a service to reconfigure app permissions.

### 2015-10-02 paper.tex
further updates and some tweaks to the bibliography
Final version :-S

@@
 \documentclass[]{llncs}
+\usepackage{microtype}
 
@@
 \subtitle{Capturing and Checking Mobile App Policies}
+\titlerunning{AppPAL for Android}
 % \numberofauthors{2}
 \author{Joseph Hallett \and David Aspinall }
-\institute{University of Edinburgh}
+\institute{School of Informatics, University of Edinburgh}
 
@@
   It can be difficult to find mobile apps that respect your security and privacy.
-  Employers rely on employees enforcing company policies about apps at the workplace correctly.
+  Businesses rely on employees enforcing company mobile device policies correctly.
   Users must judge apps by the information shown to them by the store.
-  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and
-  most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
-  We present AppPAL: a policy language for Android to describe app installation  policies.
-  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2008vn}, adding delegation relationships and the connecting to static analysis tools.
-  We use AppPAL to explore what kinds of policies are used when installing apps, demonstrating the privacy paradox when comparing user's privacy views to their actual behaviour.
+  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
+  To address these problems, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
+  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2008vn}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
+  AppPAL also acts as a ``glue'', allowing connection to a variety of local constraint checkers (e.g., static analysis tools, packager manager checks) to combine their results.
+  As well as introducing AppPAL and some examples, we apply it to explore whether real users follow certain intended policies in practice, finding privacy preferences and actual behaviour are not always aligned, in the absence of a rigorous enforcement mechanism.
   \end{abstract}
 
-\section{Introduction \comment{1 page}}
+\section{Introduction}
 \label{sec:introduction}
@@
 Clicking through each one they can find the permissions each requests but not the reasons why each was needed.
-They can see review scores from users but not experts who can run tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
+They can see review scores from users but from tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
 If they want to use the app at work will it break their employers rules for mobile usage?
 
-App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
+App stores give some information about their apps; descriptions, screenshots and review scores.
 Android apps show a list of permissions when they're first installed.
@@
 These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
-This is error-prone.
 An alternative is to write the policy down and make the computer enforce it.
@@
 The language allows us to reason about apps using statements from third parties.
-The implementation allows us to enforce the policies on a device.
+AppPAL allows us to enforce the policies on a device.
 We can express trust relationships amongst these parties; use constraints to do additional checks, such as using static analysis results.
@@
   if "home-policy" isMetBy(App)
-  where at("work") = false.
+  where at("work") = False.
     \end{lstlisting}
@@
   if "work-policy" isMetBy(App)
-  where beforeHourOfDay("17") = true.
+  where beforeHourOfDay("17") = True.
     \end{lstlisting}
@@
 We can write policies specifying which permissions an app must or must not have by its app store categorization.
-For example it would be okay allowing a photography app access to the camera, but not to the location data if the user doesn't want their photos geotagged.
+For example it would be okay allowing a photography app access to the camera, but not to location data if the user doesn't want their photos geotagged.
 \begin{lstlisting}
@@
     category(App, "Photography"),
-    hasPermission(App, "LOCATION") = false,
-    hasPermission(App, "CAMERA") = true.
+    hasPermission(App, "LOCATION") = False,
+    hasPermission(App, "CAMERA") = True.
 \end{lstlisting}
@@
 
-AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owner's are trying to enforce.
+AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owners are trying to enforce.
 This avoids creating tools with hard-coded fixed policies
@@
 \begin{lstlisting}
-"play-store" says App isInstallable
+"play-store" says App isSellable
   if App isAnApp
-  where mallodroidCheck(App) = true,
-         mcafeeAVCheck(App) = true.
+  where mallodroidCheck(App) = True,
+         mcafeeAVCheck(App) = True.
 \end{lstlisting}
 
-% \noindent
-% In this paper we have:
-% \begin{itemize}
-%   \item
-%     Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
-%     and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
-%   \item
-%     Described our implementation of the AppPAL language as a library for Android and the JVM.
-%     We show how the language can describe app properties and Android security policies (\autoref{ssec:language}) as well as how to check policies (\autoref{ssec:eval}).
-%   \item
-%     Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
-% \end{itemize}
-
-\section{Enforcing a policy at work}
+\section{Enforcing A Policy At Work}
 \label{sec:problem}
@@
 
-\section{Expressing policies in AppPAL}
+\section{Expressing Policies In AppPAL}
 \label{sec:idea}
@@
 In \autoref{sec:problem} Alice and Emma had policies they wanted to enforce but no means to do so.
-Instead of using several different tools to enforce Emma's policy disjointedly, we could use an authorization logic.
+Instead of using several tools to enforce Emma's policy disjointedly, we could use an authorization logic.
 In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
@@
   where
-    mcAfeeVirusCheck(App) = false.
+    mcAfeeVirusCheck(App) = False.
 
@@
     where
-      hasPermission(App, "LOCATION")=false.
+      hasPermission(App, "LOCATION")=False.
 \end{lstlisting}
@@
 
-In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
-She allows her to delegate the decision further if she chooses (\code{can-say inf}).
-Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
-If Emma can be convinced all these policies are met then she will say the \code{App isRunnable}.
+In \autoref{lst:corporate_1} Alice less Emma  specify whether an \code{App} (a variable) \code{isRunnable};
+she allows her to delegate the decision further (\code{can-say}\/~\code{inf}).
+Emma specifies her concerns as policies to be met in \autoref{lst:corporate_2}:
+  if Emma is convinced all these are met then she will say the \code{App isRunnable}.
 In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
-  with \code{"google-play"} specified as the decider of what is buyable or not.
-Google is not allowed to delegate the decision further (\code{can-say} as oposed to \code{can-say inf}).
-In other words Google is not allowed to specify Amazon as a supplier of apps as well.
-Google must say what is buyable directly for Emma to accept it.
-Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
-Here we use a constraint.
+  with \code{"google-play"} as the decider of what is buyable or not.
+Google is not allowed to delegate the decision further, i.e.~Google is not allowed to specify Amazon as a supplier of apps as well.
+Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5} using a constraint.
 When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
@@
 
-Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
-To meet Emma's policy the AppPAL checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
-Specifically it needs:
+Alice wants to install a new app (\code{com.facebook.katana}) on her phone
+she needs to collect statements to show the app meets the \code{isRunnable} predicate.
+Namely:
 \begin{itemize}
-  \item\code{"emma" says "com.facebook.katana" isAnApp.}
-    A simple typing statement that can be generated for all apps as they are encountered.
-    This helps keep the number of assertions in the policy low aiding readability.
   \item\code{"google-play" says "com.facebook.katana" isReputable.}
     Required to convince Emma the app came from a reputable source.
-    It should be able to obtain this statement from the Play store as the app is available there.
   \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
@@
 These last two statements require the checker to do some extra checks to satisfy the constraints.
-To get the third statement it must run the \ac{av} program on her app and check the result.
+To get the second statement AppPAL must run the \ac{av} program on her app and check the result.
 The results from the \ac{av} program may change with time as it's signatures are updated;
   so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
-For the forth statement the checker needs to check the permissions of the app.
+For the third statement the AppPAL checker needs to examine the permissions of the app.
 It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if it is running on a device.
@@
         & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
-        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
+        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = True}}^{\bracetext{constraint}}.
       \end{array}
@@
 Some static analysis tools are not quick to run.
-Even taking minutes to run a battery draining analysis can be undesirable.
-If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
+Even taking minutes to run a battery draining analysis can be undesirable:
+if a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
 
 In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
-In Emma's policy we checked this using the app's permissions.
-If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
+In Emma's policy we checked this using the app's permissions;
+if the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
 Some apps may want to access this data, but may not leak it.
@@
   where
-    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
-    hasPermission(App, "ACCESS_COARSE_LOCATION") = false.
+    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
+    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.
 
@@
   where
-    flowDroidCheck(App, "Location", "Internet") = false.
+    flowDroidCheck(App, "Location", "Internet") = False.
 \end{lstlisting}
@@
   where
-    mallodroidSSLCheck(App) = false,
-    connectionsCheck(App, "[https://emma.com]") = true.
+    mallodroidSSLCheck(App) = False,
+    connectionsCheck(App, "[https://emma.com]") = True.
 \end{lstlisting}
@@
 
-\subsection{Policy checking}
+\subsection{Policy Checking}
 \label{ssec:eval}
@@
 AppPAL uses an assertion context of known facts and rules, as well as facts deduced while checking.
-We do not use Becker~\etal's DatalogC based checking algorithm, as no DatalogC library exists for Android.
-We have implemented the rules directly in Java.
-Pseudo-code is given in \autoref{fig:pseudocode}.
+While Becker~\etal used a DatalogC based checking algorithm, we have implemented the rules directly in Java as no DatalogC library is currently available for Android.
+Pseudo-code is shown in \autoref{fig:pseudocode}.
 
@@
   we can fetch the assertion that the subject is an app based on the app in question.
+When delegating we will also be able to request facts from the delegated party dynamically; though implementing this is future work.
 
@@
 
-We envisage AppPAL running on a mobile phone, checking apps when they are installed.
+When AppPAL runs on a mobile phone apps should be checked as they are installed.
 Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
@@
 Each benchmark consisted of a chain of delegations.
-The \emph{1 to 1} benchmark consists of a repeated delegation between all of the principals.
+The \emph{1 to 1} benchmark consists of a repeated delegation between all the principals.
 In the \emph{1 to 2} benchmark each principal delegated to 2 others and in the \emph{1 to 3} benchmark each principal delegated to 3 others.
@@
 
-Throughout we have asserted that user's have policies and that there is a need for policy enforcement tools.
+Throughout we have asserted that users have policies and that there is a need for policy enforcement tools.
 Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
-In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
-Using app installation data from Carat~\cite{Oliner:2013ht} we used AppPAL to find the apps satisfying each policy Lin~\etal~identified and measured the extent each user was following them.
+In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq} demonstrating a refinement of Westin's privacy segmentation index~\cite{Krane:2002jo}.
+Using app installation data from Carat~\cite{Oliner:2013ht,Chia:2012gz} we used AppPAL to find the apps satisfying each policy Lin~\etal~identify and measure the extent that each user was following a policy.
 
@@
 We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
-These simplify the privacy policies identified as by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
+These simplify the privacy policies identified by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
 Using AppPAL we could write more precise rules if we could determine why each permission was requested.
@@
 
-McAfee classify malware into several categories.
+It is also interesting to discover when people install apps classified as malware.
+McAfee classify malware into several categories, and provided us with a dataset of apps classified as malware and \ac{pup}.
 The \emph{malicious} and \emph{trojan} categories describe traditional malware.
@@
 
-We want to test how well policies capture user behaviour.
+We now want to test how closely user behavior follows policies.
 Installation data was taken from a partially anonymized\footnote{Users are replaced with incrementing numbers, app names are replaced with hashes to protect sensitive names.} database of installed apps captured by Carat~\cite{Oliner:2013ht}.
@@
 The initial database has over 90,000 apps and 55,000 users.
-On average each user installed around 90 apps each; 4,300 apps have known names.
+On average each Carat user installed around 90 apps each; 4,300 apps have known names.
 Disregarding system apps (such as \texttt{com.android.vending}) and very common apps (Facebook, Dropbox, Whatsapp, and Twitter) we reduced the set to an average of 20 known apps per user.
@@
 \begin{figure}\centering
-  \subfigure[Use of policies modelling user behaviour.]{%
+  \subfigure[Uptake of Lin~\etal policies.]{%
     \includegraphics[width=0.48\linewidth]{./figures/lin.eps}
     \label{fig:lin}}
-  \subfigure[Percentage of malware in installed apps for users installing some malicious apps.]{%
+  \subfigure[Uptake of malware and PUPs.]{%
     \includegraphics[width=0.48\linewidth]{./figures/malware.eps}
     \label{fig:malware}}
-  \caption{Policy compliance graphs.}
+    \caption{Policy compliance graphs.
+Each histogram shows the number of users who followed a policy to a certain extent.
+Users who installed no malware have been omitted from Figure~\autoref{fig:malware}.}
 \end{figure}
@@
 Figure~\ref{fig:lin} shows that very few users follow Lin~\etal's policies most of the time.
-Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (often referred to as the \emph{privacy paradox}).
+Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (reminiscent of the \emph{privacy paradox}); assuming the user population studied by Lin~\etal behave similarly to data from the Carat study.
 A few users, however, did seem to be installing apps meeting these policies most of the time.
 This suggests that while users may have privacy preferences the majority are not attempting to enforce them.
-Policy enforcement tools, like AppPAL, can help users enforce their own policies which they cannot do easily using the current means available to them.
+Policy enforcement tools, like AppPAL, can help users enforce their own policies which they cannot do easily using the current ad hoc, manual  means available to them.
 
@@
     \label{fig:versus_pup}}
-    \subfigure[Advanced and MalloDroid policies.]{%
+    \subfigure[Advanced and SSL policies.]{%
     \includegraphics[width=0.48\linewidth]{./figures/advanced-v-mallodroid_m.eps}
     \label{fig:versus_mallodroid}}
-  \caption{Compliance with the advanced policy and the non-\ac{pup} and SSL policies.}
+    \caption{Compliance with the advanced policy and the non-\ac{pup} and SSL policies.
+Each data-point represents a user.
+In (a) we see that users who followed the Advanced policy more than 50\% of the time did not install any malware.
+In (b) we see that even users who followed the Advanced policy were no better at avoiding apps with SSL problems than any other users.}
   \label{fig:versus}
@@
 SSL misconfigurations are dangerous as they can undermine any privacy guarantees that SSL/TLS usage gives.
-MalloDroid can scan for several different misconfigurations, and distinguishes cases where the app is definitely misconfigured from those where there is some doubt.
-We set up AppPAL to use MalloDroid results as a constraint and measured the percentage of apps each user had installed that did not have issues or suspected issues when scanned with MalloDroid.
-Users who were complying with the advanced policy were no better at avoiding apps with SSL errors than any other users (Figure~\autoref{fig:versus_mallodroid}).
-This emphasizes that AppPAL policies are more powerful than manually picking apps as they can incorporate checks privacy aware users may want but which they cannot check for without additional tools.
+MalloDroid can scan for several misconfigurations, and distinguishes cases where the app is definitely misconfigured from those where there is some doubt.
+We set up AppPAL to use MalloDroid results as a constraint and measured the percentage of apps each Carat user had installed that did not have issues or suspected issues when scanned with MalloDroid.
+Users who were complying with the advanced policy were no better at avoiding apps with SSL errors than any other users, see Figure~\autoref{fig:versus_mallodroid}.
+This emphasizes that AppPAL can help enforce complex policies that cannot be checked for without additional tools.
 
-%This is significant (P-value $< 0.05$) and suggests that users who pick their apps carefully are less likely to experience malware.
+There are some limitations in this investigation: firstly
+we do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
+So a user may have apps installed that break the policy without us knowing.
+Secondly recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
+Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
+Finally, as mentioned, we have compared a different set of users to the ones Lin~\etal looked at.
+We plan to do a more comprehensive user study in the future that investigates AppPAL in use with different communities.
 
-Some limitations of this investigation include:
-\begin{itemize}
-\item
-  We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
-  So a user may have apps installed that break the policy without us knowing.
-\item
-  Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
-  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
-\item
-  The AppPAL policies we used are a simplification of the identified privacy preferences.
-  User's could be following policies more nuanced than our implementation.
-  We believe our approximation is valid, however, as users will not know the precise reason an app uses a permission, and may decline to install it anyway.
-  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND\_SMS} permission to run).
-\end{itemize}
-
-\section{Related work}
+\section{Related Work}
 
@@
 
-\section{Conclusions and further work}
+\section{Conclusions and Further Work}
 
-We have presented AppPAL: an authorization logic for describing app installation policies primarily designed to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways.
-Using app installation data and user privacy policies expressed using AppPAL we showed how the privacy paradox can be seen in the Android ecosystem.
-We showed how static analysis tools can be integrated into AppPAL to check for properties beyond the scope of the language and to act as a glue between different policies.
+We have presented AppPAL: an authorization logic for describing app installation policies primarily designed to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways, e.g.
+restricting the use of certain apps while at work.
+We showed how static analysis tools can be integrated into AppPAL to check for complex properties and to act as a glue between different policies.
 
@@
 Ad-blocking software works by users subscribing to filter policies written by experts.
-EasyList is a popular choice and they offer many different policies for specific use-cases\footnote{\url{https://easylist.adblockplus.org/en/}}.
+EasyList is a popular choice and they offer many policies for specific use-cases\footnote{\url{https://easylist.adblockplus.org/en/}}.
 We can imagine a similar scheme working well for app installation policies.
@@
 
-We might attempt to learn policies from existing users behavior.
+We might attempt to learn policies from existing user's behavior.
 Given app usage data, from a project like Carat~\cite{Oliner:2013ht}, we could identify security conscious users.
@@
 It gives us a framework for describing and evaluating policies for Android apps.
-The work provides new ways for users to enforce their own rules about how apps should behave.
-Users policies can be enforced more reliably, and with less interaction.
+The work provides new, rigorous, ways for machines to enforce user's and device-owner's rules about how apps should behave.
+These policies can be enforced more reliably, and with less interaction from person operating the device.
+
+\subsection*{Acknowledgements}
+
+Thanks to Igor~Muttik at McAfee, and N.~Asokan at Aalto University and the University of Helsinki for discussions and providing us with data used in \autoref{sec:demonstation}.
 

### 2015-12-21 paper.tex
updates based on feedback

@@
   To address these problems, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
-  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2008vn}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
+  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2009ko}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
   AppPAL also acts as a ``glue'', allowing connection to a variety of local constraint checkers (e.g., static analysis tools, packager manager checks) to combine their results.
@@
 Clicking through each one they can find the permissions each requests but not the reasons why each was needed.
-They can see review scores from users but from tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
+They can see review scores from users but not from tools to check apps for problems and issues like SSL misconfigurations~\cite{Fahl:2012dj}.
 If they want to use the app at work will it break their employers rules for mobile usage?
@@
 Android apps show a list of permissions when they're first installed.
-Soon new apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
+In Android Marshmallow apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
 Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
@@
 AppPAL allows us to enforce the policies on a device.
-We can express trust relationships amongst these parties; use constraints to do additional checks, such as using static analysis results.
+We can express trust relationships amongst these parties and use constraints to do additional checks, such as using static analysis results.
 This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.
@@
 A user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
-Using AppPAL we can write for work and home, and decide which policy to enforce using a user's location, or the time of day:
+Using AppPAL we can write policies for work and home, and decide which policy to enforce using a user's location, or the time of day:
 
@@
     \begin{lstlisting}
-"alice" says App isRunnable
-  if "home-policy" isMetBy(App)
-  where at("work") = False.
+'alice' says App isRunnable
+  if 'home-policy' isMetBy(App)
+  where at('work') = false.
     \end{lstlisting}
@@
     \begin{lstlisting}
-"alice" says App isRunnable
-  if "work-policy" isMetBy(App)
-  where beforeHourOfDay("17") = True.
+'alice' says App isRunnable
+  if 'work-policy' isMetBy(App)
+  where beforeHourOfDay('17') = true.
     \end{lstlisting}
@@
 \begin{lstlisting}
-"alice" says "it-department" can-say "work-policy" isMetBy(App).
-"alice" says "alice" can-act-as "it-department".
+'alice' says 'it-department' can-say 'work-policy' isMetBy(App).
+'alice' says 'alice' can-act-as 'it-department'.
 \end{lstlisting}
@@
 \begin{lstlisting}
-"alice" says App isRunnable
-  if "permissions-policy" isMetBy(App).
-"alice" says "permissions-policy" isMetBy(App)
+'alice' says App isRunnable
+  if 'permissions-policy' isMetBy(App).
+'alice' says 'permissions-policy' isMetBy(App)
   if App isAnApp
   where
-    category(App, "Photography"),
-    hasPermission(App, "LOCATION") = False,
-    hasPermission(App, "CAMERA") = True.
+    category(App, 'Photography'),
+    hasPermission(App, 'LOCATION') = false,
+    hasPermission(App, 'CAMERA') = true.
 \end{lstlisting}
@@
 Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
-TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Fritz:2013vi} can do taint and control flow analysis; sometimes even between app components.
+TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Arzt:2014kf,Li:2015wo} can do taint and control flow analysis; sometimes even between app components.
 Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
@@
 AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owners are trying to enforce.
-This avoids creating tools with hard-coded fixed policies
+This avoids creating tools with hard-coded fixed policies.
 For example a store might not want to sell apps with SSL errors or malicious apps as determined by their anti-virus.
@@
 \begin{lstlisting}
-"play-store" says App isSellable
+'play-store' says App isSellable
   if App isAnApp
-  where mallodroidCheck(App) = True,
-         mcafeeAVCheck(App) = True.
+  where mallodroidCheck(App) = true,
+         mcafeeAVCheck(App) = true.
 \end{lstlisting}
@@
 She might even sign a document promising never to break the rules within the policy.
-This is error-prone---what if she makes a mistake or misses an app that breaks her policy
+This is error-prone---what if she makes a mistake or misses an app that breaks her policy.
 Alternately Emma's policy could be partially enforced using existing tools.
@@
 
-We could implement Emma's policy using existing tools (such as an AV checker, and a taint analysis tool like Flowdroid~\cite{Fritz:2013vi}) but it is a clumsy solution---they are not flexible.
+We could implement Emma's policy using existing tools (such as an AV checker, and a taint analysis tool like Flowdroid~\cite{Arzt:2014kf,Li:2015wo}) but it is a clumsy solution---they are not flexible.
 Each has to be configured separately to implement only part of the policy.
@@
                        keywordstyle=\scriptsize\slshape]
-"alice" says "emma" can-say inf
+'alice' says 'emma' can-say inf
   App isRunnable. @\label{lst:corporate_1}@
 
-"emma" says App isRunnable @\label{lst:corporate_2}@
-  if "no-tracking-policy" isMetBy(App),
-     "reputable-policy" isMetBy(App),
-     "anti-virus-policy" isMetBy(App).
+'emma' says App isRunnable @\label{lst:corporate_2}@
+  if 'no-tracking-policy' isMetBy(App),
+     'reputable-policy' isMetBy(App),
+     'anti-virus-policy' isMetBy(App).
 
-"emma" says
-  "reputable-policy" isMetBy(App) @\label{lst:corporate_3}@
+'emma' says
+  'reputable-policy' isMetBy(App) @\label{lst:corporate_3}@
       if App isBuyable.
 
-"emma" says "google-play" can-say
+'emma' says 'google-play' can-say
   App isBuyable. @\label{lst:corporate_4}@
@@
                        keywordstyle=\scriptsize\slshape]
-"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
+'emma' says 'anti-virus-policy' isMetBy(App) @\label{lst:corporate_5}@
   if App isAnApp
   where
-    mcAfeeVirusCheck(App) = False.
+    mcafeeAVCheck(App) = true.
 
-"emma" says "no-location-permissions"
-  can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
+'emma' says 'no-location-permissions'
+  can-act-as 'no-tracking-policy'. @\label{lst:corporate_6}@
 
-"emma" says
-  "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
+'emma' says
+  'no-location-permissions' isMetBy(App) @\label{lst:corporate_7}@
     if App isAnApp
     where
-      hasPermission(App, "LOCATION")=False.
+      hasPermission(App, 'COARSE_LOCATION')=false,
+      hasPermission(App, 'FINE_LOCATION')=false.
 \end{lstlisting}
@@
 
-Alice wants to install a new app (\code{com.facebook.katana}) on her phone
-she needs to collect statements to show the app meets the \code{isRunnable} predicate.
+Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
+She needs to collect statements to show the app meets the \code{isRunnable} predicate.
 Namely:
@@
 The parser is implemented using ANTLR4.
-The structure of an AppPAL is inherited from SecPAL~\cite{Becker:2006vh} can be seen in \autoref{fig:assertion}.
+AppPAL's syntax is inherited from SecPAL~\cite{Becker:2006vh} (shown in \autoref{fig:assertion}).
 
@@
         & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
-        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = True}}^{\bracetext{constraint}}.
+        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
       \end{array}
@@
 Some apps may want to access this data, but may not leak it.
-We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Fritz:2013vi}).
+We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Arzt:2014kf,Li:2015wo}).
 Our policy now becomes:
@@
 \begin{lstlisting}
-"emma" says "no-locations-permissions"
-  can-act-as "no-tracking-policy".
+'emma' says 'no-locations-permissions'
+  can-act-as 'no-tracking-policy'.
 
-"emma" says "no-locations-permissions" isMetBy(App)
+'emma' says 'no-locations-permissions' isMetBy(App)
   if App isAnApp
   where
-    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
-    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.
+    hasPermission(App, 'ACCESS_FINE_LOCATION') = false,
+    hasPermission(App, 'ACCESS_COARSE_LOCATION') = false.
 
-"emma" says "location-taint-analysis"
-  can-act-as "no-tracking-policy".
+'emma' says 'location-taint-analysis'
+  can-act-as 'no-tracking-policy'.
 
-"emma" says "location-taint-analysis" isMetBy(App)
+'emma' says 'location-taint-analysis' isMetBy(App)
   if App isAnApp
   where
-    flowDroidCheck(App, "Location", "Internet") = False.
+    flowDroidCheck(App, 'Location', 'Internet') = false.
 \end{lstlisting}
@@
 \begin{lstlisting}
-"emma" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
-"emma" says "relaxed-no-tracking-policy" isMetBy(App)
-  if App hasCategory("tracking")
+'emma' says 'relaxed-no-tracking-policy' canActAs 'no-tracking-policy'.
+'emma' says 'relaxed-no-tracking-policy' isMetBy(App)
+  if App hasCategory('tracking')
   where
-    mallodroidSSLCheck(App) = False,
-    connectionsCheck(App, "[https://emma.com]") = True.
+    mallodroidSSLCheck(App) = false,
+    connectionsCheck(App, '[https://emma.com]') = true.
 \end{lstlisting}
@@
 AppPAL uses an assertion context of known facts and rules, as well as facts deduced while checking.
-While Becker~\etal used a DatalogC based checking algorithm, we have implemented the rules directly in Java as no DatalogC library is currently available for Android.
+While Becker~\etal~used a DatalogC based checking algorithm, we have implemented the rules directly in Java as no DatalogC library is currently available for Android.
 Pseudo-code is shown in \autoref{fig:pseudocode}.
@@
 \begin{lstlisting}
-"user" says "mcafee" can-say
-  "malware" isKindOf(App).
-"mcafee" says "trojan" can-act-as "malware".
-"mcafee" says "pup" can-act-as "malware".
+'user' says 'mcafee' can-say
+  'malware' isKindOf(App).
+'mcafee' says 'trojan' can-act-as 'malware'.
+'mcafee' says 'pup' can-act-as 'malware'.
 \end{lstlisting}
@@
 \begin{figure}\centering
-  \subfigure[Uptake of Lin~\etal policies.]{%
+  \subfigure[Uptake of Lin~\etal's policies.]{%
     \includegraphics[width=0.48\linewidth]{./figures/lin.eps}
@@
 Figure~\ref{fig:lin} shows that very few users follow Lin~\etal's policies most of the time.
-Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (reminiscent of the \emph{privacy paradox}); assuming the user population studied by Lin~\etal behave similarly to data from the Carat study.
+Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (reminiscent of the \emph{privacy paradox}); assuming the user population studied by Lin~\etal~behave similarly to data from the Carat study.
 A few users, however, did seem to be installing apps meeting these policies most of the time.
@@
 
-There are some limitations in this investigation: firstly
-we do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
+There are some limitations in this investigation: 
+firstly, we do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
 So a user may have apps installed that break the policy without us knowing.
@@
 Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
-Finally, as mentioned, we have compared a different set of users to the ones Lin~\etal looked at.
+Finally, as mentioned, we have compared a different set of users to the ones Lin~\etal~looked at.
 We plan to do a more comprehensive user study in the future that investigates AppPAL in use with different communities.
@@
 The SecPAL language is designed for access control in distributed systems.
-We picked SecPAL as the basis for AppPAL because it is readable, extensible, and is a good for for the mobile ecosystem setting~\citep{Hallett:2014un}.
+We picked SecPAL as the basis for AppPAL because it is readable, extensible, and is a good fit for the mobile ecosystem setting~\citep{Hallett:2014un}.
 It has also been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
@@
 Policy authors can specify combinations of permissions and broadcast events that should not appear together.
-For example to stop malware sending premium rate text messages, we prevent
- an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions.
+For example to stop malware sending premium rate text messages, we prevent an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions.
 \begin{lstlisting}
@@
 With Kirin we are restricted to permitting or allowing apps.
-AppPAL can describe more scenarios than just permit or allow, and use more app information than just permissions, such as constraints and static analysis results.
+AppPAL can describe more scenarios than just allow or forbid, and use more app information than just permissions, such as constraints and static analysis results.
 By allowing delegation relationships we can understand the provenance and trust relationships in these rules.
@@
 
-Thanks to Igor~Muttik at McAfee, and N.~Asokan at Aalto University and the University of Helsinki for discussions and providing us with data used in \autoref{sec:demonstation}.
+Thanks to Igor~Muttik at McAfee, and N~Asokan at Aalto University and the University of Helsinki for discussions and providing us with data used in \autoref{sec:demonstation}.
 

### 2015-12-21 figures/benchmarks.R

@@ -5,7 +5,7 @@ setEPS()
 postscript("benchmarks.eps", width=6, height=4)
 
 data <- read.csv('benchmarks.csv', header=TRUE)
-qplot(Principals, Check, data=data, group=Delegations, color=Delegations, geom="line") +
+qplot(Principals, Check, data=data, group=Delegations, linetype=Delegations, geom="line") +
   scale_x_log10(breaks=c(10,100,500,1000)) +
   scale_y_log10(breaks=c(1,10,100)) +
   ylab("Check time (s)") +

### 2016-01-08 paper.tex
copyright and hanging package
last minute tweaks

@@
 \usepackage{localstyle}
+\usepackage{hanging}
 
@@
 \begin{abstract}
-  It can be difficult to find mobile apps that respect your security and privacy.
+  It can be difficult to find mobile apps that respect one's security and privacy.
   Businesses rely on employees enforcing company mobile device policies correctly.
   Users must judge apps by the information shown to them by the store.
-  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
-  To address these problems, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
+  Studies have found that most users do not pay attention to an apps permissions during installation~\cite{Felt:2012hm} and most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
+  To address these problems and more, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
   AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2009ko}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
   AppPAL also acts as a ``glue'', allowing connection to a variety of local constraint checkers (e.g., static analysis tools, packager manager checks) to combine their results.
-  As well as introducing AppPAL and some examples, we apply it to explore whether real users follow certain intended policies in practice, finding privacy preferences and actual behaviour are not always aligned, in the absence of a rigorous enforcement mechanism.
+  As well as introducing AppPAL and some examples, we apply it to explore whether users follow certain intended policies in practice, finding privacy preferences and actual behaviour are not always aligned in the absence of a rigorous enforcement mechanism.
   \end{abstract}
@@
 Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
-Ultimately the decision which apps to use and which permissions to grant must be made by the device user.
+Ultimately the decision of which apps to use and which permissions to grant must be made by the device user.
 
@@
 An employer may restrict which apps their employees can use.
-The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
-Some users worry apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
-Such a user avoids apps which can access their location, or address book.
-They may apply their own personal security policies when downloading and running apps.
+The IT department may set a mobile device policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
+Some users worry that apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
+Such a user avoids apps which can access their location, or address book; they may apply their own personal security policies when downloading and running apps.
 
@@
 An alternative is to write the policy down and make the computer enforce it.
-To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
-The policy is written in the logic and enforced by checking the policy is satisfied.
+To implement this we propose using a logic of authorization---a language designed to express rules about permissible actions.
 
-We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool.} and predicates that allow us to decide which apps to run or install.
+We present AppPAL, an instantiation of Becker~\etal's
+SecPAL~\cite{Becker:2006vh} with constraints (statements checkable using information external to the language such as the time of day or static analysis tools) and predicates that allow us to decide which apps to run or install.
 The language allows us to reason about apps using statements from third parties.
 AppPAL allows us to enforce the policies on a device.
-We can express trust relationships amongst these parties and use constraints to do additional checks, such as using static analysis results.
+We can express trust relationships amongst these parties and use constraints to do additional checks, such as using security checks.
 This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.
+Policies can be enforced by the stores selling the apps, on the devices installing apps or by third-parties providing app vetting services.
 
-A user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
+Consider the following example:
+  a user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
 Using AppPAL we can write policies for work and home, and decide which policy to enforce using a user's location, or the time of day:
@@
 \end{minipage}
-
-We can delegate policy specification to third parties or roles, and assign principals to roles:
+\noindent We can delegate policy specification to third parties or roles, and assign principals to roles:
 \begin{lstlisting}
@@
 We can write policies specifying which permissions an app must or must not have by its app store categorization.
-For example it would be okay allowing a photography app access to the camera, but not to location data if the user doesn't want their photos geotagged.
+For example, it would be okay allowing a photography app access to the camera, but not to allow access to location data if the user doesn't want their photos geotagged.
 \begin{lstlisting}
@@
 TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Arzt:2014kf,Li:2015wo} can do taint and control flow analysis; sometimes even between app components.
-Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
+Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between apps.
 ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
@@
 MalloDroid~\citep{Fahl:2012dj} looks for apps configured to use SSL incorrectly (for instance by not verifying hostnames or certificates).
-Many others exist checking and certifying other aspects of app behaviour.
 
@@
 This avoids creating tools with hard-coded fixed policies.
-For example a store might not want to sell apps with SSL errors or malicious apps as determined by their anti-virus.
+ For example a store might not want to sell apps with SSL errors or apps flagged by an anti-virus tool.
 Using AppPAL we can combine tools for checking apps to implement the store's policies.
@@
 \end{lstlisting}
+No additional attempt is made to ensure these static analysis tools are sound.
+The policy designer must be aware of the tool's limitations.
+Black or whitelisting may have to be used to avoid some false positives or negatives.
 
@@
 She might even sign a document promising never to break the rules within the policy.
-This is error-prone---what if she makes a mistake or misses an app that breaks her policy.
-Alternately Emma's policy could be partially enforced using existing tools.
+This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
+Alternatively Emma's policy could be partially enforced using existing tools.
 \emph{Google's Device Policy for Android}~\cite{GoogleAppsDeviceP:tz} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}.
@@
 Each has to be configured separately to implement only part of the policy.
-If Emma changes her policy or Alice changes jobs she needs to recheck her apps and then alter or remove the software on her phone to ensure compliance.
+If Emma changes her policy or Alice changes jobs she must recheck her apps and then alter or remove the software on her phone to ensure compliance.
 It isn't clear what an app must do to be run, or what checks have been done if it is already running on the phone.
@@
 Alice might not be allowed to use location tracking apps at work but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
-Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disable at run time.
+Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disabled at run time.
 Can we enforce different policies at different times or locations?
@@
 
-We propose a change to the mobile ecosystem shown in \autoref{fig:ecosystem}.
+We propose using the mobile ecosystem shown in \autoref{fig:ecosystem}.
 People have policies which are enforced by AppPAL on their devices.
@@
 The device can make use of vetting services which run tools to infer complex properties about apps.
-Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
+Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the store's explicit policies, or ones requested by users.
 Developers could decide which stores to sell their apps in on the basis of policies about stores.
@@
 
-In \autoref{sec:problem} Alice and Emma had policies they wanted to enforce but no means to do so.
+In \autoref{sec:problem}, Alice and Emma had policies they wanted to enforce but no means to do so.
 Instead of using several tools to enforce Emma's policy disjointedly, we could use an authorization logic.
@@
 SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
-We present AppPAL as a modified form of SecPAL, targeting apps on mobile devices.
+We present AppPAL as a modified form of SecPAL, aimed at mobile apps.
+
+Other access control languages, such as XACML~\cite{Oasis:1M71_oFS}, could also have been used as the basis for AppPAL.
+We chose to use SecPAL as the basis, however, as it was more lightweight, designed to be readable, and simpler to extend with additional constraints than XACML.
 
 \begin{figure}
-  \begin{minipage}[t]{0.5\textwidth}
+  \begin{minipage}[t]{0.44\textwidth}
     \begin{lstlisting}[numbers=left,
@@
     \end{lstlisting}
-  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
+  \end{minipage}\begin{minipage}[t]{0.56\textwidth}
     \begin{lstlisting}[numbers=left,
@@
 
-In \autoref{lst:corporate_1} Alice less Emma  specify whether an \code{App} (a variable) \code{isRunnable};
-she allows her to delegate the decision further (\code{can-say}\/~\code{inf}).
+In \autoref{lst:corporate_1} Alice lets Emma specify whether an \code{App} (a variable) \code{isRunnable};
+she allows her to delegate the decision (\code{can-say}\/~\code{inf}).
 Emma specifies her concerns as policies to be met in \autoref{lst:corporate_2}:
-  if Emma is convinced all these are met then she will say the \code{App isRunnable}.
+  if Emma is convinced that these are met then she will say the \code{App isRunnable}.
 In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
-  with \code{"google-play"} as the decider of what is buyable or not.
+  with \code{"google-play"} deciding of what is buyable or not.
 Google is not allowed to delegate the decision further, i.e.~Google is not allowed to specify Amazon as a supplier of apps as well.
@@
 Only if this returns \code{false} will the policy be met.
-To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
+To specify the \code{"no-tracking-policy"} Emma says that the \code{'no-location-permissions'} rules implement the \code{'no-tracking-policy'} (\autoref{lst:corporate_6}).
 Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.
@@
 Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
-She needs to collect statements to show the app meets the \code{isRunnable} predicate.
-Namely:
+She collects statements to show the app meets the \code{isRunnable} predicate.
 \begin{itemize}
   \item\code{"google-play" says "com.facebook.katana" isReputable.}
-    Required to convince Emma the app came from a reputable source.
+    Required to convince Emma that the app came from a reputable source.
   \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
@@
     Needed to show the App meets Emma's no-tracking-policy.
-    Emma will say this if after examining the app the location permissions are missing.
+    Emma will say this if the app has no location permissions.
 \end{itemize}
@@
 To get the second statement AppPAL must run the \ac{av} program on her app and check the result.
-The results from the \ac{av} program may change with time as it's signatures are updated;
+The results from the \ac{av} program may change with time as its signatures are updated;
   so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
@@
 AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
-Additional predicates can be created in the policy files, however constraints are implemented individually.
-For example on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.
+Additional predicates can be created in the policy files, however constraints must be implemented individually.
+For example, on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.
 
@@
 }              \\
-  \bottomrule
+  \bottomrule \\
 \end{tabulary}
@@
 if a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
+In that case, it may be preferable to delegate to an online database.
 
@@
 In Emma's policy we checked this using the app's permissions;
-if the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
+if the app couldn't get access to the GPS sensors (using the permissions) then it meets this policy.
 Some apps may want to access this data, but may not leak it.
 We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Arzt:2014kf,Li:2015wo}).
-Our policy now becomes:
+Our policy becomes:
 
@@
 On a mobile device memory is at a premium.
-We would like to keep the assertion context as small as possible.
+We want to keep the assertion context as small as possible.
 For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
@@
   we can fetch the assertion that the subject is an app based on the app in question.
-When delegating we will also be able to request facts from the delegated party dynamically; though implementing this is future work.
+When delegating we will also be able to request facts from the delegated party dynamically (although this is not yet implemented).
 
@@
 
-When AppPAL runs on a mobile phone apps should be checked as they are installed.
+When AppPAL runs on a mobile phone, apps should be checked as they are installed.
 Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
@@
 In the \emph{1 to 2} benchmark each principal delegated to 2 others and in the \emph{1 to 3} benchmark each principal delegated to 3 others.
-These benchmarks are reasonable as they model the slowest kinds of policies to evaluate---though worse ones could be designed by delegating even more.
+These benchmarks are reasonable as they model the slowest kinds of policies to
+evaluate---though worse ones could be designed by delegating even more or triggering an expensive constraint check.
 
@@
 
-Throughout we have asserted that users have policies and that there is a need for policy enforcement tools.
+Throughout we have asserted that users often have informal policies and that there is a need for policy enforcement tools.
 Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
@@
 The \emph{Unconcerned} (U) users felt okay allowing access to most data for almost any reason.
-The \emph{Advanced} (A) users were comfortable allowing apps access to location data but not if it was for advertising reasons.
-Opinions in the largest cluster, \emph{Fencesitters} (F), varied but were broadly against collection of personal data for advertising purposes.
+The \emph{Advanced} (A) users were comfortable allowing apps access to location data but not if it was for advertising.
+Opinions in the largest cluster, \emph{Fencesitters} (F), varied but were broadly against collection of personal data for advertising.
 We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
-These simplify the privacy policies identified by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
-Using AppPAL we could write more precise rules if we could determine why each permission was requested.
+These simplify the privacy policies identified by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission
+(we could write more precise rules if we could determine why each permission was requested).
 Lin~\etal~used Androguard~\cite{Desnos:ub} as well as manual analysis to determine the precise reasons for each permission~\cite{Sadeh:2014vq}.
@@
     \midrule
-    \lstinline{GET_ACCOUNTS}           & \xmark                 & \xmark             & \xmark                & \xmark                \\
-    \lstinline{ACCESS_FINE_LOCATION}   & \xmark                 & \xmark             & \xmark                &                       \\
-    \lstinline{READ_CONTACT}           & \xmark                 & \xmark             & \xmark                &                       \\
-    \lstinline{READ_PHONE_STATE}       & \xmark                 & \xmark             &                       &                       \\
-    \lstinline{SEND_SMS}               & \xmark                 & \xmark             &                       &                       \\
-    \lstinline{ACCESS_COARSE_LOCATION} & \xmark                 &                    &                       &                       \\
+    \texttt{GET\_ACCOUNTS}            & \xmark                 & \xmark             & \xmark                & \xmark                \\
+    \texttt{ACCESS\_FINE\_LOCATION}   & \xmark                 & \xmark             & \xmark                &                       \\
+    \texttt{READ\_CONTACT}            & \xmark                 & \xmark             & \xmark                &                       \\
+    \texttt{READ\_PHONE\_STATE}       & \xmark                 & \xmark             &                       &                       \\
+    \texttt{SEND\_SMS}                & \xmark                 & \xmark             &                       &                       \\
+    \texttt{ACCESS\_COARSE\_LOCATION} & \xmark                 &                    &                       &                       \\
     \bottomrule
@@
 It is also interesting to discover when people install apps classified as malware.
-McAfee classify malware into several categories, and provided us with a dataset of apps classified as malware and \ac{pup}.
+McAfee classify malware into several categories, and provided us with a dataset of apps classified as malware and \ac{pup}s.
 The \emph{malicious} and \emph{trojan} categories describe traditional malware.
 Other categories classify \ac{pup} such as aggressive adware.
-Using AppPAL we can write policies to differentiate between different kinds of malware, characterising users who allow dangerous apps and those who install poor quality ones.
+Using AppPAL we can write policies to differentiate characterising users who allow dangerous apps and those who install poor quality ones.
 \begin{lstlisting}
@@
 Disregarding system apps (such as \texttt{com.android.vending}) and very common apps (Facebook, Dropbox, Whatsapp, and Twitter) we reduced the set to an average of 20 known apps per user.
-To see some variations in app type, we considered only the 44,000 users who had more than 20 known apps.
+To see some variation in app type, we considered only the 44,000 users who had more than 20 known apps.
 Using this data, and the apps themselves taken from the Google Play Store and Android Observatory~\cite{Barrera:2012iba}, we checked which apps satisfied which policies.
@@
 Figure~\ref{fig:lin} shows that very few users follow Lin~\etal's policies most of the time.
-Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between users privacy preferences and their behaviour (reminiscent of the \emph{privacy paradox}); assuming the user population studied by Lin~\etal~behave similarly to data from the Carat study.
+Whilst the AppPAL policy we used was a simplified version of Lin~\etal's policy, it suggests that there is a disconnect between user's privacy preferences and their behaviour (reminiscent of the \emph{privacy paradox}); assuming the user population studied by Lin~\etal~behave similarly to data from the Carat study.
 A few users, however, did seem to be installing apps meeting these policies most of the time.
@@
 
-We found 1\% of the users had a \ac{pup} or malicious app installed.
+We found that 1\% of the users had a \ac{pup} or malicious app installed.
 Figure~\ref{fig:malware} shows that infection rates for \ac{pup}s and malware is low;
@@
 The \emph{MalloDroid} tool~\citep{Fahl:2012dj} can scan apps for SSL misconfigurations.
-SSL misconfigurations are dangerous as they can undermine any privacy guarantees that SSL/TLS usage gives.
-MalloDroid can scan for several misconfigurations, and distinguishes cases where the app is definitely misconfigured from those where there is some doubt.
+SSL misconfigurations are dangerous as they can undermine any privacy guarantees that SSL/TLS gives.
+MalloDroid distinguishes cases where the app is definitely misconfigured from those where there is some doubt.
 We set up AppPAL to use MalloDroid results as a constraint and measured the percentage of apps each Carat user had installed that did not have issues or suspected issues when scanned with MalloDroid.
-Users who were complying with the advanced policy were no better at avoiding apps with SSL errors than any other users, see Figure~\autoref{fig:versus_mallodroid}.
-This emphasizes that AppPAL can help enforce complex policies that cannot be checked for without additional tools.
+Users who were complied with the advanced policy were no better at avoiding apps with SSL errors than any other users, see Figure~\autoref{fig:versus_mallodroid}.
+This emphasizes that AppPAL can help enforce complex policies that cannot be checked without additional tools.
 
-There are some limitations in this investigation: 
-firstly, we do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
+There are limitations in this study:
+first, we do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
 So a user may have apps installed that break the policy without us knowing.
-Secondly recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
+Second, recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
 Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
@@
 
-Authorization logics have been successfully used to enforce policies in several other domains.
-The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general, if undecidable.
+Authorization logics have been used to enforce policies in several other domains.
+The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general and undecidable.
 Logics that followed like KeyNote~\cite{Blaze:1999fa} and SPKI/SDSI~\cite{Ellison:1999ui} looked at public key infrastructure.
@@
 Google also offer the \emph{Device Policy for Android} app.
-This lets businesses configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
-It cannot be used to describe policies about apps, or describe trust relationships, however.
+This lets businesses configure company-owned devices to be trackable, remote lockable, set passwords and sync with their servers.
+It cannot be used to describe policies about apps, or describe trust relationships.
 
@@
 It has also been used to describe data usage policies~\cite{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
-Other work on SecPAL has added various features such as existential quantification~\cite{Becker:2009vt} and the DKAL family of policy languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
+Other work has added various features such as existential
+quantification~\cite{Becker:2009vt} and extended to the DKAL family of languages~\cite{Gurevich:2008fz,Gurevich:Qo5E3M3}.
 DKAL contains more modalities than \emph{says}, which lets policies describe actions principals carry out rather than just their opinions.
@@
 With DKAL we can guarantee that the action was completed, whereas in AppPAL we do not know if the user actually installed a particular app.
-We chose to use SecPAL as the basis for AppPAL as we did not need the extra features DKAL added to express app installation policies.
-If, in future work, we need additional modalities AppPAL could be extended to support additional DKAL features as SecPAL has been shown to be a subset of the DKAL language.
+We chose to use SecPAL as the basis for AppPAL as we did not need the extra features DKAL added to express app installation policies for our initial applications.
 
@@
 Policy authors can specify combinations of permissions and broadcast events that should not appear together.
-For example to stop malware sending premium rate text messages, we prevent an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions.
-\begin{lstlisting}
-restrict permission [SEND_SMS] and permission [WRITE_SMS]
-\end{lstlisting}
+For example, to stop malware sending premium rate text messages, we prevent an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions:
+
+\lstinline!restrict permission [SEND_SMS] and permission [WRITE_SMS]!
+
 By analyzing apps which broke their policies Enck~\etal~found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
 
-This approach could help identify malware, but it is less suitable for detecting \acp{pup}.
+The Kirin approach has been shown to help identify malware, but it is less suitable for detecting \acp{pup}.
 The behaviours and permissions \ac{pup} displays aren't necessarily malicious.
@@
 
-We have presented AppPAL: an authorization logic for describing app installation policies primarily designed to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways, e.g.
-restricting the use of certain apps while at work.
-We showed how static analysis tools can be integrated into AppPAL to check for complex properties and to act as a glue between different policies.
+We have presented AppPAL: a language for describing app installation policies to help achieve security and privacy objectives but which can also \emph{lock down} devices in other ways, e{.}g{.}~restricting the use of certain apps while at work.
+We showed how static analysis tools can be connected to AppPAL to compose complex properties.
 
-AppPAL currently runs either an Android app or a Java program.
 Further work is needed to tightly integrate AppPAL into Android.
@@
 This is undesirable as it would make AppPAL difficult to install for most users, and negate the other security enhancements (such as timely updates and patches) provided by the standard Android system.
-Alternatively, AppPAL could be integrated as a service to reconfigure app permissions.
+AppPAL could be integrated as a service to reconfigure app permissions.
 Android Marshmallow has an iOS like permissions model where permissions can be granted and revoked at any time.
 These will be manually configurable by the user through the settings app.
-We can imagine AppPAL working to reconfigure these settings (and set their initial grant or deny states) based on a user's policy, as well as the time of day or the user's location.
+We can imagine AppPAL working to reconfigure these settings (and set the device's initial grant or deny states) based on a user's policy, as well as the time of day or the user's location.
 A policy could deny notifications while a user is driving, for example, by checking if they are using Android Auto~\cite{AndroidAuto:uw} (an app to interact with a car's center console) or moving along a road at high speed.
 
-Developing, and testing, policies for users is a key next step.
+Future work includes developing and testing, policies for users.
 Here we described a policy being specified by a user's employer.
-For most end-users writing a policy in a formal language is too much work.
-Ad-blocking software works by users subscribing to filter policies written by experts.
-EasyList is a popular choice and they offer many policies for specific use-cases\footnote{\url{https://easylist.adblockplus.org/en/}}.
+For most end-users writing a policy in a formal language unrealistic.
+With Ad-blocking software users subscribe to filter policies written by experts, such as EasyList\footnote{\url{https://easylist.adblockplus.org/en/}}.
 We can imagine a similar scheme working well for app installation policies.
-Users subscribe to different policies by experts (examples could include no tracking apps, nothing with adult content, no spammy in-app-purchase apps).
-Optionally they can customize them further.
+Users subscribe to different policies by experts (examples could include no tracking apps, nothing with adult content, no in-app-purchase apps).
+Optionally the users could customize the policies further, similar to proof-carrying code~\cite{Necula:1996tr}.
+
+Policy composition raises further questions: what should happen when user's personal and work policies overlap?  
+What should happen when policies contradict? 
+Future work will look at detecting these problems as well as integrating strategies to resolve them.
+
+Another question might be whether we can use evidence to speed re-checking apps against a policy.
+Some static analysis tools, such as Evicheck~\cite{Seghir:2015er}, can create evidence that lets you check an app doesn't have certain behavior faster than it would be to infer the same property in the app without it.
+We can also imagine apps being distributed with evidence that proves the app meets an AppPAL policy but avoids the need to check the against the policy explicitly.
 
@@
 If we can infer these users policies we may be able to describe new policies that the less technical users may want.
-Given a set of apps one user has already installed, we could learn policies about what their personal installation policy is.
+Given a set of apps one user has already installed, we could learn policies about what their personal security relevant installation policy is.
 This may help stores show users apps they're more likely to buy, and users apps that already behave as they want.
 
-AppPAL is a powerful language for describing app installation policies.
-It gives us a framework for describing and evaluating policies for Android apps.
+AppPAL gives us a framework for describing and evaluating policies for Android apps.
 The work provides new, rigorous, ways for machines to enforce user's and device-owner's rules about how apps should behave.
@@
 Thanks to Igor~Muttik at McAfee, and N~Asokan at Aalto University and the University of Helsinki for discussions and providing us with data used in \autoref{sec:demonstation}.
+Thanks also to the App Guarden project and colleagues at the University of Edinburgh for their comments, and the referees for their feedback.
 

### 2016-01-08 paper.tex
final version

@@
 
-There has been a great amount of work on developing app analysis tools for Android.
+There has been much work developing app analysis tools for Android.
 Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
@@
 
-Other access control languages, such as XACML~\cite{Oasis:1M71_oFS}, could also have been used as the basis for AppPAL.
-We chose to use SecPAL as the basis, however, as it was more lightweight, designed to be readable, and simpler to extend with additional constraints than XACML.
+Other access control languages, such as XACML 3.0~\cite{Oasis:1M71_oFS}, could also have been used as the basis for AppPAL.
+SecPAL however can capture distribution and delegation relationships between
+principles and serves as a simplified model of a more complex system like XACML,
+and has a well defined semantics, and decidability.
 
@@
 In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
-  with \code{"google-play"} deciding of what is buyable or not.
+  with \code{'google-play'} deciding of what is buyable or not.
 Google is not allowed to delegate the decision further, i.e.~Google is not allowed to specify Amazon as a supplier of apps as well.
-Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5} using a constraint.
+Emma specifies the \code{'anti-virus-policy'} in \autoref{lst:corporate_5} using a constraint.
 When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
 Only if this returns \code{false} will the policy be met.
-To specify the \code{"no-tracking-policy"} Emma says that the \code{'no-location-permissions'} rules implement the \code{'no-tracking-policy'} (\autoref{lst:corporate_6}).
+To specify the \code{'no-tracking-policy'} Emma says that the \code{'no-location-permissions'} rules implement the \code{'no-tracking-policy'} (\autoref{lst:corporate_6}).
 Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.
@@
 \begin{itemize}
-  \item\code{"google-play" says "com.facebook.katana" isReputable.}
+  \item\code{'google-play' says 'com.facebook.katana' isReputable.}
     Required to convince Emma that the app came from a reputable source.
-  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
+  \item\code{'emma' says 'anti-virus-policy' isMetBy('com.facebook.katana').}
     She can obtain this by running the \ac{av} program on her app.
-  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
+  \item\code{'emma' says 'no-locations-permissions' isMetBy('com.facebook.katana').}
     Needed to show the App meets Emma's no-tracking-policy.
@@
 Policy authors can specify combinations of permissions and broadcast events that should not appear together.
-For example, to stop malware sending premium rate text messages, we prevent an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions:
-
-\lstinline!restrict permission [SEND_SMS] and permission [WRITE_SMS]!
+For example, to stop malware sending premium rate text messages, we prevent an
+app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions one
+could write: \texttt{restrict permission [SEND\_SMS] and permission [WRITE\_SMS]}.
 
@@
 For most end-users writing a policy in a formal language unrealistic.
-With Ad-blocking software users subscribe to filter policies written by experts, such as EasyList\footnote{\url{https://easylist.adblockplus.org/en/}}.
+With Ad-blocking software users subscribe to filter policies written by experts, such as EasyList~\cite{Easylist}.
 We can imagine a similar scheme working well for app installation policies.
 Users subscribe to different policies by experts (examples could include no tracking apps, nothing with adult content, no in-app-purchase apps).
-Optionally the users could customize the policies further, similar to proof-carrying code~\cite{Necula:1996tr}.
+Optionally the users could customize the policies further.
 
-Policy composition raises further questions: what should happen when user's personal and work policies overlap?  
-What should happen when policies contradict? 
+Policy composition raises further questions: what should happen when user's
+personal and work policies overlap or contradict?  
 Future work will look at detecting these problems as well as integrating strategies to resolve them.
@@
 Another question might be whether we can use evidence to speed re-checking apps against a policy.
-Some static analysis tools, such as Evicheck~\cite{Seghir:2015er}, can create evidence that lets you check an app doesn't have certain behavior faster than it would be to infer the same property in the app without it.
+Some static analysis tools, such as Evicheck~\cite{Seghir:2015er}, can create evidence that lets you check an app doesn't have certain behavior faster than it would be to infer the same property in the app without it, similar to proof-carrying code~\cite{Necula:1996tr}.
 We can also imagine apps being distributed with evidence that proves the app meets an AppPAL policy but avoids the need to check the against the policy explicitly.
@@
 The work provides new, rigorous, ways for machines to enforce user's and device-owner's rules about how apps should behave.
-These policies can be enforced more reliably, and with less interaction from person operating the device.
+These policies can be enforced more reliably, and with less interaction from the person operating the device.
 
